### 2017-10-16 manuscript/manuscript-HTML.Rmd
manuscripts - initial commit
manuscript-HTML will produce an HTML preview (easy for editing)

manuscript-Word will produce a word document (to export in a format that co-authors perfer)

Eventually, a latex version will be used, but I won’t add it here, yet.

---
title: "`respR` - An R package for the analysis of respirometry data"
author: "Januar Harianto, Nicholas Carey, Maria Byrne"
output: 
  html_document:
    toc: false
    number_sections: true
    fig_caption: true
    pandoc_args: ["-F","pandoc-crossref"]
    theme: "united"
---

```{r, echo = F}
library(respR)
library(knitr)
library(readr)

# set pander table-layout options
library(pander)
panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

opts_chunk$set(collapse = TRUE, comment = "#>", tidy = TRUE, highlight = TRUE)
```

# Introduction

Metabolic rate is a fundamental trait linked to virtually all other biological functions, and is key to the prediction of patterns in ecology and conservation biology, from populations (Seibel and Drazen, 2007, Barneche et al., 2014) to ecosystems (Brown et al., 2004). It has been widely used to investigate the effects of external stressors on an organism, and has seen increased popularity in studies involving climate change where warming is expected to drive increased metabolism in ectotherms (Pörtner, 2002, McElroy et al., 2012, Carey et al., 2016, Delorme & Sewell, 2016). By far the most common metric used to measure physiological performance in animals, it has been applied from the whole organism to the level of cells and tissues (White & Kearney, 2013), and is typically quantified using respirometry, which measures oxygen uptake over time (Lighton, 2008).

There are at least four broad methodological approaches in respirometry: closed, intermittent flow, flowthrough and open. In closed respirometry, $O_2$ decrease is measured within a hermetically sealed chamber of known volume, sometimes set within a closed loop to allow circulation or mixing of the environment within the chamber. The oxygen recordings may be continuous through use of an oxygen probe, periodic through withdrawing water or gas samples at set intervals, or a two-point measurement consisting of the initial and final concentrations. Metabolic rates are estimated from respirometry by assuming a linear relationship between variables, and estimates of metabolic rate are straightforward in constant volume respirometry using the equation: $$VO_2 = \Delta O_2(V)$$ where $\Delta O_2$ is the slope of the regression that relates O~2~ concentration to time, or in the case of a two-point measurement, the difference in $O_2$ concentration divided by time, and $V$ is the volume of the container (Lighton, 2008).

In intermittent flow respirometry, $O_2$ concentration is measured, but after a set time period the chamber is flushed with new water or air, returning it to initial conditions, sealed again, and the experiment repeated (Svendsen et al., 2016). This technique is essentially closed respirometry, but with the incorporation of repeated measures. Depending on the metabolic rate metric being investigated, final respiration rate is calculated as the mean of the measures (e.g. citation), or the lowest or highest rates recorded in any trial (e.g. Carey et al., 2016).

Flowthrough respirometry involves a closed chamber, but with a regulated flow of air or water through it at a precisely determined rate. After equilibrium has been achieved, the oxygen concentration differential between the incurrent and excurrent channels, along with the flow rate, allows calculation of the oxygen extracted from the flow volume per unit time: $$VO_2 = (C_iO_2 - C_eO_2)FR$$ where $VO_2$ is the rate of $O_2$ consumption in milligrams per minute, $C_iO_2$ and $C_eO_2$ are the incurrent and excurrent $O_2$ concentrations, and $FR$ is the flow rate of water through the system  (Lighton, 2008).

A final method is open respirometry, in which an open tank or semi-enclosed area is used, but the input or mixing rate of oxygen from the surroundings is known or found to be negligible relative to oxygen consumption of the specimens (Leclercq et al., 1999). It is seldom used, but for some applications it is a sufficient and practical methodology (e.g. Gamble et al., 2014). The common equation used for open respirometry is: $$VO_2 = \Delta O_2 + \phi_d$$ where $\Delta O_2$ is the slope of the regression that relates $O_2$ concentration to time, and $\phi_d$ is the air-sea oxygen flux as determined by Fick's Law (Leclercq et al., 1999).

Researchers are increasingly assimilating large, high-resolution respirometry data, often capturing multiple response variables such as maximum metabolic rate (MMR, or $MO_{2,max}$), the rate under high activity or exhaustive exercise, or standard (or resting) metabolic rate (SMR, or $MO_{2,min}$), the minimum metabolic cost of maintaining biological functioning (Rogers et al., 2016), or critical oxygen tension, which represents the lowest level of oxygen at which aerobic metabolism is independent of the ambient partial pressure of oxygen ($P_{c}$; Yeager & Ultsch, 1989, Hochachka and Somero, 2002). Experiments may run for long periods (e.g. 20 h, Norin & Malte, 2012), and in most cases, processing the data involves an *ad hoc* selection of data points and them manually performing calculations in a spreadsheet program (e.g. Microsoft Excel) or an integrated development environment (IDE, e.g. R and Matlab). These approaches can be tedious and time consuming especially when spreadsheet programs (e.g. Excel) struggle with the vast datasets that are generated, while IDEs require a degree of expertise and have substantial learning curves. Dedicated software are available to perform such calculations, but most are proprietary and require licensing requirements or dedicated hardware (e.g. AutoResp by Loligo Systems), complicating or preventing their use on multiple machines.

Recently, a number of free R software packages designed for respirometry have become available (e.g. `respirometry`: https://CRAN.R-project.org/package=respirometry, `rMR`: https://CRAN.R-project.org/package=rMR, `LoLinR`: Olito et al., 2017). Both `respirometry` and `rMR` are useful for aquatic respirometry, but are not focused on the analysis of data, although both packages have functions for the analysis of metabolic rate over intervals, and `rMR` has a specialised function to calculate $P_{c}$. For `LoLinR`, the package provides a statistically robust way of automatically detecting a best-fit regression which are applicable for closed and flow-through respirometry. However, `LoLinR`'s processing time increases exponentially with large datasets, with data containing more than 500 observations taking a few minutes to run (Olito et al., 2017), which makes `LoLinR` intractable with large amounts of data.

Here we describe the R package `respR`, a set of functions desgined to provide a structured workflow for the analysis of respirometry-related data (Fig. 1). The package contains utilities to: (1) analyse closed, intermittent, flow-through and open respirometry data, (2) scale $\Delta O_2$ to volume and/or mass, and (3) automatically and rapidly detect MMR, SMR, $P_{c}$ or best-fit results in large datasets with the help of traditional rolling regression and kernel density estimation techniques. Other smaller, but useful, functions are also avaliable, and are described in more detail in our online html vignette (https://januarharianto.github.io/respR/). While the package has a strong focus on marine and freshwater respirometry, the principals also apply to aerial respirometry. Some oxygen concentration conversions require a salinity value, but in freshwater respirometry this should be set to zero. We demonstrate the utility of the package by analysing three datasets as case studies and proof of concept. These data were collected using closed, intermittent and flow-through respirometry.

# Package overview

`respR` can be automatically installed from the GitHub repository using the `devtools` package:

```{r eval=F}
devtools::install_github("januarharianto/respR")
library(respR)
```

To explore `respR` and its main functions we have provided example data that can be called directly after the package is loaded. The first data object, `urchin2013`, contains measurements of oxygen consumption in 16 individual sea urchins (*Heliocidaris erythrogramma*) and 2 "blank" measurements of background respiration (Harianto, *unpublished*). The `intermittent` data object contains three repeated measures of oxygen consumption in *H. erythrogramma*, collected in a single session with intermittent respirometry (Carey et al. 2016, *JEB*). The `flowthrough` data object contains measurements of oxygen consumption, via flow-through respirometry, in a species of chiton, *Mopalia lignosa* (Carey, *unpublished*). Detailed information about all example data, including their source and methods, can be obtained with the `?` command in the R console (e.g. `?urchin2013`).

## Data import and exploration

Data should be formatted correctly before use in `respR`. The function `check.input()` performs error checks on a data frame of any size, and can extract a two-column data frame for data exploration and subsequent analyses. Data must be numeric, because `respR` has limited support for date/time data (e.g. `POSIXct` and `POSIXlt` classes) and performing calculations on different time formats have many caveats that are beyond the scope of the package. Time data should be sequential, without duplicates, as that might signify instrumental drift. Ideally, time data should also be evenly-spaced, that is, the time period between samples are equal. This has important ramifications on data extraction techniques later on as substantial errors may occur in the event that calculations are based on the number of rows instead of time. A plot of the data is also shown for visual inspection (e.g. Fig. 1).

It should be noted that invoking `check.input` is optional -- the main functions in our package will readily accept any data frame as long as data are numeric. Running `check.input` is an exploratory step that flags potential issues about the data before it is further analysed.

## Estimating rates of change in $O_2$ concentration ($RO_2$)

The functions `calc.bg.rate()`, `calc.rate()`, `auto.rate()` and `pcrit()` can process raw respirometry data and are called depending on the type of analysis required.

### Measuring background changes in $RO_2$

Background $RO_2$ is measured to account for the influence of bacterial respiration in the chamber. Since background rates typically account for a small percentage of experimental rates, these often-called “blank” experiments are routinely conducted alongside main experiments, and the rates are averaged across several datasets to obtain a more accurate estimate of the correction. The function `calc.bg.rate()` can simultaneously process multiple background rate measurements, as long as they share the time data, and produces an output object that can be used in later functions that accept background rate arguments.

### Manual and automatic selection of data for calculations

A typical procedure in the analysis respirometry data is to select the most linear section of the data for calculations of $RO_2$ (e.g. citations). The function `calc.rate()` can manually extract one or more segments from a single data frame to process them simultaneously. Data segments can be selected by time period, row numbers, $O_2$ decrease or proportion, which can accomodate most, if not all, data selection requirements and allow for consistent reporting of methods and results (e.g. estimating $RO_2$ over a 10-minute time window across multiple specimens).

Alternatively, we can take advantage of automated methods in `auto.rate()`, which implements novel rolling regression and kernel density estimation algorithms to detect patterns in the data to automatically obtain the maximum, minimum or most linear sections of the data. There are clear advantages in using `auto.rate()` when estimating the prior mentioned metrics. Unlike manual methods where data is most likely arbitrary selected, `auto.rate()` adheres to well-defined rules to obtain the rate of interest, so its methods are results are transparent and fully reproducible. When called, the function takes the data frame of length $m$, and performs a rolling regression of sample size $n$ iteratively across the entire data frame (Fig. 2b). Thus, a total of $(m−n)+1$ number of iterative regressions are fitted, and we obtain a new dataset of regressions as a function of time (Fig. 3), which is then ranked by size to obtain maximum and minimum values.

To estimate the most linear rate, further computations are performed after rolling regressions. Logically, the most linear section of the data should reflect a consistently stable (i.e. flat) value in the rolling regression data. Thus the value will have a high probability density. The function `auto.rate()` takes advantage of this relationship to identify data that contains the most consistent rate over a rolling window (which equate to peaks in the density plot) and then recalculates the rate across the new data window. This is repeated for all linear sections of the data that are identified by the function, and the results are ranked by size.

### Critical oxygen tension, $P_{c}$

- oxygen tolerance (Hochachka and Somero, 2002)
- broken stick model (Yeager and Ultsch, 1989)
- $RO_2$ vs $P_{O_2}$ data can be generated automatically from raw $O_2$ vs time data.

## Converting units of $RO_2$ and $MO_2$

- talk about calc.mo2 and conversions to volume and weight specific $MO_2$.

# Reproducibility and data accesibility

We have ensured that any computational analysis performed by `respR` are fully reproducible. All main functions produce an output object of class `list`, which contains all the data and variables needed to re-analyse the results, within R or in other software.

# Tables & Figures

**Table 1.** Main functions available in `respR`.
```{r, echo=F, message=F}
# use this later to show table
library(readr)
listfunctions <- read_csv('tables/list-of-functions.csv')
```

```{r, echo=F, results='asis'}
pander(listfunctions)
```

```{r, echo = F, out.width = "900px"}
knitr::include_graphics("figs/fig1.png")
```

**Fig. 1.** Diagram showing a typical workflow in the analysis of respirometry data using `respR`. Data is first checked for errors before the main functions are used to extract and analyse segments of the data. Summarised results and diagnostic plots provide immediate visual feedback on the outcome of the analyses. Once the rate estimates are obtained, they can be converted into volume and/or mass-specific rates.

```{r, echo = F, out.width = "650px"}
knitr::include_graphics("figs/fig2.png")
```

**Fig. 2** Simplified illustration showing the sampling windows used by `auto.rate()` during **(a)** interval regression analysis, where sample windows do not overlap and **(b)** rolling regression analysis, where sample windows overlap and move forward by one sample unit at a time.

```{r, echo = F, , out.width = "800px"}
knitr::include_graphics("figs/fig3.png")
```

**Fig. 3** Selected output plots from the function `calc.rate()` for the analysis of the data frame `urchin2013`.

### 2017-10-16 manuscript/manuscript-Word.Rmd
manuscripts - initial commit
manuscript-HTML will produce an HTML preview (easy for editing)

manuscript-Word will produce a word document (to export in a format that co-authors perfer)

Eventually, a latex version will be used, but I won’t add it here, yet.

--- 2017-10-16 manuscript/manuscript-HTML.Rmd
@@
 ---
-title: "`respR` - An R package for the analysis of respirometry data"
-author: "Januar Harianto, Nicholas Carey, Maria Byrne"
+title: "Manuscript_Word"
+author: "Januar Harianto"
+date: "13/10/2017"
 output: 
-  html_document:
-    toc: false
-    number_sections: true
-    fig_caption: true
-    pandoc_args: ["-F","pandoc-crossref"]
-    theme: "united"
+  word_document:
+    reference_docx: "styles/styles-manuscript-Word.docx"
 ---
 
-```{r, echo = F}
+```{r setup, include=FALSE}
 library(respR)
@@
 
-opts_chunk$set(collapse = TRUE, comment = "#>", tidy = TRUE, highlight = TRUE)
+knitr::opts_chunk$set(echo = TRUE)
 ```
-
-# Introduction
-
-Metabolic rate is a fundamental trait linked to virtually all other biological functions, and is key to the prediction of patterns in ecology and conservation biology, from populations (Seibel and Drazen, 2007, Barneche et al., 2014) to ecosystems (Brown et al., 2004).
-It has been widely used to investigate the effects of external stressors on an organism, and has seen increased popularity in studies involving climate change where warming is expected to drive increased metabolism in ectotherms (Pörtner, 2002, McElroy et al., 2012, Carey et al., 2016, Delorme & Sewell, 2016).
-By far the most common metric used to measure physiological performance in animals, it has been applied from the whole organism to the level of cells and tissues (White & Kearney, 2013), and is typically quantified using respirometry, which measures oxygen uptake over time (Lighton, 2008).
-
-There are at least four broad methodological approaches in respirometry: closed, intermittent flow, flowthrough and open.
-In closed respirometry, $O_2$ decrease is measured within a hermetically sealed chamber of known volume, sometimes set within a closed loop to allow circulation or mixing of the environment within the chamber.
-The oxygen recordings may be continuous through use of an oxygen probe, periodic through withdrawing water or gas samples at set intervals, or a two-point measurement consisting of the initial and final concentrations.
-Metabolic rates are estimated from respirometry by assuming a linear relationship between variables, and estimates of metabolic rate are straightforward in constant volume respirometry using the equation: $$VO_2 = \Delta O_2(V)$$ where $\Delta O_2$ is the slope of the regression that relates O~2~ concentration to time, or in the case of a two-point measurement, the difference in $O_2$ concentration divided by time, and $V$ is the volume of the container (Lighton, 2008).
-
-In intermittent flow respirometry, $O_2$ concentration is measured, but after a set time period the chamber is flushed with new water or air, returning it to initial conditions, sealed again, and the experiment repeated (Svendsen et al., 2016).
-This technique is essentially closed respirometry, but with the incorporation of repeated measures.
-Depending on the metabolic rate metric being investigated, final respiration rate is calculated as the mean of the measures (e.g.
-citation), or the lowest or highest rates recorded in any trial (e.g.
-Carey et al., 2016).
-
-Flowthrough respirometry involves a closed chamber, but with a regulated flow of air or water through it at a precisely determined rate.
-After equilibrium has been achieved, the oxygen concentration differential between the incurrent and excurrent channels, along with the flow rate, allows calculation of the oxygen extracted from the flow volume per unit time: $$VO_2 = (C_iO_2 - C_eO_2)FR$$ where $VO_2$ is the rate of $O_2$ consumption in milligrams per minute, $C_iO_2$ and $C_eO_2$ are the incurrent and excurrent $O_2$ concentrations, and $FR$ is the flow rate of water through the system  (Lighton, 2008).
-
-A final method is open respirometry, in which an open tank or semi-enclosed area is used, but the input or mixing rate of oxygen from the surroundings is known or found to be negligible relative to oxygen consumption of the specimens (Leclercq et al., 1999).
-It is seldom used, but for some applications it is a sufficient and practical methodology (e.g.
-Gamble et al., 2014).
-The common equation used for open respirometry is: $$VO_2 = \Delta O_2 + \phi_d$$ where $\Delta O_2$ is the slope of the regression that relates $O_2$ concentration to time, and $\phi_d$ is the air-sea oxygen flux as determined by Fick's Law (Leclercq et al., 1999).
-
-Researchers are increasingly assimilating large, high-resolution respirometry data, often capturing multiple response variables such as maximum metabolic rate (MMR, or $MO_{2,max}$), the rate under high activity or exhaustive exercise, or standard (or resting) metabolic rate (SMR, or $MO_{2,min}$), the minimum metabolic cost of maintaining biological functioning (Rogers et al., 2016), or critical oxygen tension, which represents the lowest level of oxygen at which aerobic metabolism is independent of the ambient partial pressure of oxygen ($P_{c}$; Yeager & Ultsch, 1989, Hochachka and Somero, 2002).
-Experiments may run for long periods (e.g.
-20 h, Norin & Malte, 2012), and in most cases, processing the data involves an *ad hoc* selection of data points and them manually performing calculations in a spreadsheet program (e.g.
-Microsoft Excel) or an integrated development environment (IDE, e.g.
-R and Matlab).
-These approaches can be tedious and time consuming especially when spreadsheet programs (e.g.
-Excel) struggle with the vast datasets that are generated, while IDEs require a degree of expertise and have substantial learning curves.
-Dedicated software are available to perform such calculations, but most are proprietary and require licensing requirements or dedicated hardware (e.g.
-AutoResp by Loligo Systems), complicating or preventing their use on multiple machines.
-
-Recently, a number of free R software packages designed for respirometry have become available (e.g.
-`respirometry`: https://CRAN.R-project.org/package=respirometry, `rMR`: https://CRAN.R-project.org/package=rMR, `LoLinR`: Olito et al., 2017).
-Both `respirometry` and `rMR` are useful for aquatic respirometry, but are not focused on the analysis of data, although both packages have functions for the analysis of metabolic rate over intervals, and `rMR` has a specialised function to calculate $P_{c}$.
-For `LoLinR`, the package provides a statistically robust way of automatically detecting a best-fit regression which are applicable for closed and flow-through respirometry.
-However, `LoLinR`'s processing time increases exponentially with large datasets, with data containing more than 500 observations taking a few minutes to run (Olito et al., 2017), which makes `LoLinR` intractable with large amounts of data.
-
-Here we describe the R package `respR`, a set of functions desgined to provide a structured workflow for the analysis of respirometry-related data (Fig.
-1).
-The package contains utilities to: (1) analyse closed, intermittent, flow-through and open respirometry data, (2) scale $\Delta O_2$ to volume and/or mass, and (3) automatically and rapidly detect MMR, SMR, $P_{c}$ or best-fit results in large datasets with the help of traditional rolling regression and kernel density estimation techniques.
-Other smaller, but useful, functions are also avaliable, and are described in more detail in our online html vignette (https://januarharianto.github.io/respR/).
-While the package has a strong focus on marine and freshwater respirometry, the principals also apply to aerial respirometry.
-Some oxygen concentration conversions require a salinity value, but in freshwater respirometry this should be set to zero.
-We demonstrate the utility of the package by analysing three datasets as case studies and proof of concept.
-These data were collected using closed, intermittent and flow-through respirometry.
-
-# Package overview
-
-`respR` can be automatically installed from the GitHub repository using the `devtools` package:
-
-```{r eval=F}
-devtools::install_github("januarharianto/respR")
-library(respR)
-```
-
-To explore `respR` and its main functions we have provided example data that can be called directly after the package is loaded.
-The first data object, `urchin2013`, contains measurements of oxygen consumption in 16 individual sea urchins (*Heliocidaris erythrogramma*) and 2 "blank" measurements of background respiration (Harianto, *unpublished*).
-The `intermittent` data object contains three repeated measures of oxygen consumption in *H.
-erythrogramma*, collected in a single session with intermittent respirometry (Carey et al.
-2016, *JEB*).
-The `flowthrough` data object contains measurements of oxygen consumption, via flow-through respirometry, in a species of chiton, *Mopalia lignosa* (Carey, *unpublished*).
-Detailed information about all example data, including their source and methods, can be obtained with the `?` command in the R console (e.g.
-`?urchin2013`).
-
-## Data import and exploration
-
-Data should be formatted correctly before use in `respR`.
-The function `check.input()` performs error checks on a data frame of any size, and can extract a two-column data frame for data exploration and subsequent analyses.
-Data must be numeric, because `respR` has limited support for date/time data (e.g.
-`POSIXct` and `POSIXlt` classes) and performing calculations on different time formats have many caveats that are beyond the scope of the package.
-Time data should be sequential, without duplicates, as that might signify instrumental drift.
-Ideally, time data should also be evenly-spaced, that is, the time period between samples are equal.
-This has important ramifications on data extraction techniques later on as substantial errors may occur in the event that calculations are based on the number of rows instead of time.
-A plot of the data is also shown for visual inspection (e.g.
-Fig.
-1).
-
-It should be noted that invoking `check.input` is optional -- the main functions in our package will readily accept any data frame as long as data are numeric.
-Running `check.input` is an exploratory step that flags potential issues about the data before it is further analysed.
-
-## Estimating rates of change in $O_2$ concentration ($RO_2$)
-
-The functions `calc.bg.rate()`, `calc.rate()`, `auto.rate()` and `pcrit()` can process raw respirometry data and are called depending on the type of analysis required.
-
-### Measuring background changes in $RO_2$
-
-Background $RO_2$ is measured to account for the influence of bacterial respiration in the chamber.
-Since background rates typically account for a small percentage of experimental rates, these often-called “blank” experiments are routinely conducted alongside main experiments, and the rates are averaged across several datasets to obtain a more accurate estimate of the correction.
-The function `calc.bg.rate()` can simultaneously process multiple background rate measurements, as long as they share the time data, and produces an output object that can be used in later functions that accept background rate arguments.
-
-### Manual and automatic selection of data for calculations
-
-A typical procedure in the analysis respirometry data is to select the most linear section of the data for calculations of $RO_2$ (e.g.
-citations).
-The function `calc.rate()` can manually extract one or more segments from a single data frame to process them simultaneously.
-Data segments can be selected by time period, row numbers, $O_2$ decrease or proportion, which can accomodate most, if not all, data selection requirements and allow for consistent reporting of methods and results (e.g.
-estimating $RO_2$ over a 10-minute time window across multiple specimens).
-
-Alternatively, we can take advantage of automated methods in `auto.rate()`, which implements novel rolling regression and kernel density estimation algorithms to detect patterns in the data to automatically obtain the maximum, minimum or most linear sections of the data.
-There are clear advantages in using `auto.rate()` when estimating the prior mentioned metrics.
-Unlike manual methods where data is most likely arbitrary selected, `auto.rate()` adheres to well-defined rules to obtain the rate of interest, so its methods are results are transparent and fully reproducible.
-When called, the function takes the data frame of length $m$, and performs a rolling regression of sample size $n$ iteratively across the entire data frame (Fig.
-2b).
-Thus, a total of $(m−n)+1$ number of iterative regressions are fitted, and we obtain a new dataset of regressions as a function of time (Fig.
-3), which is then ranked by size to obtain maximum and minimum values.
-
-To estimate the most linear rate, further computations are performed after rolling regressions.
-Logically, the most linear section of the data should reflect a consistently stable (i.e.
-flat) value in the rolling regression data.
-Thus the value will have a high probability density.
-The function `auto.rate()` takes advantage of this relationship to identify data that contains the most consistent rate over a rolling window (which equate to peaks in the density plot) and then recalculates the rate across the new data window.
-This is repeated for all linear sections of the data that are identified by the function, and the results are ranked by size.
-
-### Critical oxygen tension, $P_{c}$
-
-- oxygen tolerance (Hochachka and Somero, 2002)
-- broken stick model (Yeager and Ultsch, 1989)
-- $RO_2$ vs $P_{O_2}$ data can be generated automatically from raw $O_2$ vs time data.
-
-## Converting units of $RO_2$ and $MO_2$
-
-- talk about calc.mo2 and conversions to volume and weight specific $MO_2$.
-
-# Reproducibility and data accesibility
-
-We have ensured that any computational analysis performed by `respR` are fully reproducible.
-All main functions produce an output object of class `list`, which contains all the data and variables needed to re-analyse the results, within R or in other software.
-
-# Tables & Figures
-
-**Table 1.** Main functions available in `respR`.
-```{r, echo=F, message=F}
-# use this later to show table
-library(readr)
-listfunctions <- read_csv('tables/list-of-functions.csv')
-```
-
-```{r, echo=F, results='asis'}
-pander(listfunctions)
-```
-
-```{r, echo = F, out.width = "900px"}
-knitr::include_graphics("figs/fig1.png")
-```
-
-**Fig.
-1.** Diagram showing a typical workflow in the analysis of respirometry data using `respR`.
-Data is first checked for errors before the main functions are used to extract and analyse segments of the data.
-Summarised results and diagnostic plots provide immediate visual feedback on the outcome of the analyses.
-Once the rate estimates are obtained, they can be converted into volume and/or mass-specific rates.
-
-```{r, echo = F, out.width = "650px"}
-knitr::include_graphics("figs/fig2.png")
-```
-
-**Fig.
-2** Simplified illustration showing the sampling windows used by `auto.rate()` during **(a)** interval regression analysis, where sample windows do not overlap and **(b)** rolling regression analysis, where sample windows overlap and move forward by one sample unit at a time.
-
-```{r, echo = F, , out.width = "800px"}
-knitr::include_graphics("figs/fig3.png")
-```
-
-**Fig. 3** Selected output plots from the function `calc.rate()` for the analysis of the data frame `urchin2013`.

### 2017-10-16 manuscript/manuscript-HTML.Rmd
manuscript update

@@
 `respirometry`: https://CRAN.R-project.org/package=respirometry, `rMR`: https://CRAN.R-project.org/package=rMR, `LoLinR`: Olito et al., 2017).
-Both `respirometry` and `rMR` are useful for aquatic respirometry, but are not focused on the analysis of data, although both packages have functions for the analysis of metabolic rate over intervals, and `rMR` has a specialised function to calculate $P_{c}$.
-For `LoLinR`, the package provides a statistically robust way of automatically detecting a best-fit regression which are applicable for closed and flow-through respirometry.
+Both `respirometry` and `rMR` are useful for aquatic respirometry, but are not focused on the analysis of data, although both packages have functions for the analysis of metabolic rate over non-overlapping intervals, and `rMR` has a specialised function to calculate $P_{c}$.
+For `LoLinR`, the package provides a statistically robust method of detecting a "best-fit" regression which is applicable for closed and flow-through respirometry.
 However, `LoLinR`'s processing time increases exponentially with large datasets, with data containing more than 500 observations taking a few minutes to run (Olito et al., 2017), which makes `LoLinR` intractable with large amounts of data.
@@
 
-The functions `calc.bg.rate()`, `calc.rate()`, `auto.rate()` and `pcrit()` can process raw respirometry data and are called depending on the type of analysis required.
+The functions `calc.bg.rate()`, `calc.rate()`, `auto.rate()`, `calc.rate.ft()` and `pcrit()` can process raw respirometry data and are called depending on the experimental setup and type of analysis required.
 
@@
 
-Background $RO_2$ is measured to account for the influence of bacterial respiration in the chamber.
-Since background rates typically account for a small percentage of experimental rates, these often-called “blank” experiments are routinely conducted alongside main experiments, and the rates are averaged across several datasets to obtain a more accurate estimate of the correction.
+Background $RO_2$ is measured to account for the influence of bacterial respiration in the chamber, and once measured, is often applied as a correction to the main experiment (Rodgers et al., 2016).
+Since background rates typically account for a small percentage of experimental rates, these “blank” experiments are routinely conducted alongside main experiments, and the rates are averaged across several datasets to obtain a more accurate estimate of the correction.
 The function `calc.bg.rate()` can simultaneously process multiple background rate measurements, as long as they share the time data, and produces an output object that can be used in later functions that accept background rate arguments.
@@
 The function `calc.rate()` can manually extract one or more segments from a single data frame to process them simultaneously.
-Data segments can be selected by time period, row numbers, $O_2$ decrease or proportion, which can accomodate most, if not all, data selection requirements and allow for consistent reporting of methods and results (e.g.
-estimating $RO_2$ over a 10-minute time window across multiple specimens).
+Data segments can be selected by (1) time period, (2) row numbers, (3) $O_2$ decrease or (4) proportion, which should accomodate most, if not all, data selection requirements and allow for consistent reporting of methods and results (e.g.
+estimating $RO_2$ over a regular 10-minute time window across multiple specimens).
 
-Alternatively, we can take advantage of automated methods in `auto.rate()`, which implements novel rolling regression and kernel density estimation algorithms to detect patterns in the data to automatically obtain the maximum, minimum or most linear sections of the data.
-There are clear advantages in using `auto.rate()` when estimating the prior mentioned metrics.
+In addition, we can take advantage of automated methods in `auto.rate()`, which implements novel rolling regression and kernel density estimation algorithms to detect patterns in the data to automatically obtain the maximum or minimum rate of fixed width, or to detect the most linear sections of the data.
 Unlike manual methods where data is most likely arbitrary selected, `auto.rate()` adheres to well-defined rules to obtain the rate of interest, so its methods are results are transparent and fully reproducible.
@@
 To estimate the most linear rate, further computations are performed after rolling regressions.
-Logically, the most linear section of the data should reflect a consistently stable (i.e.
-flat) value in the rolling regression data.
-Thus the value will have a high probability density.
-The function `auto.rate()` takes advantage of this relationship to identify data that contains the most consistent rate over a rolling window (which equate to peaks in the density plot) and then recalculates the rate across the new data window.
-This is repeated for all linear sections of the data that are identified by the function, and the results are ranked by size.
+Logically, linear sections of the data should reflect a consistently stable (i.e.
+flat) value in the rolling regression data (Fig.
+3b).
+The function `auto.rate()` takes advantage of this relationship to identify data windows that contain consistent rates over a rolling window, implementing the methods of Sheather & Jones (1991), avaliable in the `stats` package in base R, to select the bandwidth using pilot estimation of derivatives and calculate "peaks" in the density that correspond to stable rate values (Fig.
+3c).
+The function then selects the rolling data windows that fall within the kernel density binwidth, merges all overlapping windows into one, and performs a linear regression on each new window.
 

### 2017-10-17 manuscript/manuscript-HTML.Rmd
manuscript update
tried to fix equations but not sure if they are right, yet

@@
 It has been widely used to investigate the effects of external stressors on an organism, and has seen increased popularity in studies involving climate change where warming is expected to drive increased metabolism in ectotherms (Pörtner, 2002, McElroy et al., 2012, Carey et al., 2016, Delorme & Sewell, 2016).
-By far the most common metric used to measure physiological performance in animals, it has been applied from the whole organism to the level of cells and tissues (White & Kearney, 2013), and is typically quantified using respirometry, which measures oxygen uptake over time (Lighton, 2008).
+By far the most common metric used to measure physiological performance in animals, it can be applied from the whole organism to the level of cells and tissues (White & Kearney, 2013), and is typically quantified using respirometry, which measures oxygen uptake over time (Lighton, 2008).
 
@@
 The oxygen recordings may be continuous through use of an oxygen probe, periodic through withdrawing water or gas samples at set intervals, or a two-point measurement consisting of the initial and final concentrations.
-Metabolic rates are estimated from respirometry by assuming a linear relationship between variables, and estimates of metabolic rate are straightforward in constant volume respirometry using the equation: $$VO_2 = \Delta O_2(V)$$ where $\Delta O_2$ is the slope of the regression that relates O~2~ concentration to time, or in the case of a two-point measurement, the difference in $O_2$ concentration divided by time, and $V$ is the volume of the container (Lighton, 2008).
+Metabolic rates are estimated from respirometry by assuming a linear relationship between variables, and estimates of metabolic rate are straightforward in constant volume respirometry using the equation: $$RO_2 = \Delta O_2(V)$$ where $\Delta O_2$ is the slope of the regression that relates O~2~ concentration to time, or in the case of a two-point measurement, the difference in $O_2$ concentration divided by time, and $V$ is the volume of the container (Lighton, 2008).
 
@@
 Flowthrough respirometry involves a closed chamber, but with a regulated flow of air or water through it at a precisely determined rate.
-After equilibrium has been achieved, the oxygen concentration differential between the incurrent and excurrent channels, along with the flow rate, allows calculation of the oxygen extracted from the flow volume per unit time: $$VO_2 = (C_iO_2 - C_eO_2)FR$$ where $VO_2$ is the rate of $O_2$ consumption in milligrams per minute, $C_iO_2$ and $C_eO_2$ are the incurrent and excurrent $O_2$ concentrations, and $FR$ is the flow rate of water through the system  (Lighton, 2008).
+After equilibrium has been achieved, the oxygen concentration differential between the incurrent and excurrent channels, along with the flow rate, allows calculation of the oxygen extracted from the flow volume per unit time: $$VO_2 = (C_iO_2 - C_eO_2)FR$$ where $VO_2$ is the rate of $O_2$ consumption over time, $C_iO_2$ and $C_eO_2$ are the incurrent and excurrent $O_2$ concentrations, and $FR$ is the flow rate of water through the system  (Lighton, 2008).
 
@@
 Gamble et al., 2014).
-The common equation used for open respirometry is: $$VO_2 = \Delta O_2 + \phi_d$$ where $\Delta O_2$ is the slope of the regression that relates $O_2$ concentration to time, and $\phi_d$ is the air-sea oxygen flux as determined by Fick's Law (Leclercq et al., 1999).
+The common equation used for open respirometry is: $$RO_2 = \Delta O_2(V) + \phi_d$$ where $\Delta O_2$ is the slope of the regression that relates $O_2$ concentration to time, $V$ is the volume of the container and $\phi_d$ is the air-sea oxygen flux as determined by Fick's Law (Leclercq et al., 1999).
 
@@
 `POSIXct` and `POSIXlt` classes) and performing calculations on different time formats have many caveats that are beyond the scope of the package.
-Time data should be sequential, without duplicates, as that might signify instrumental drift.
+Time data should be sequential, without duplicates, as such errors strongly indicate underlying sampling issues.
 Ideally, time data should also be evenly-spaced, that is, the time period between samples are equal.
-This has important ramifications on data extraction techniques later on as substantial errors may occur in the event that calculations are based on the number of rows instead of time.
+This has important ramifications on data extraction techniques later on, as substantial errors may occur in the event that calculations are based on the number of rows instead of time.
 A plot of the data is also shown for visual inspection (e.g.
@@
 Background $RO_2$ is measured to account for the influence of bacterial respiration in the chamber, and once measured, is often applied as a correction to the main experiment (Rodgers et al., 2016).
-Since background rates typically account for a small percentage of experimental rates, these “blank” experiments are routinely conducted alongside main experiments, and the rates are averaged across several datasets to obtain a more accurate estimate of the correction.
-The function `calc.bg.rate()` can simultaneously process multiple background rate measurements, as long as they share the time data, and produces an output object that can be used in later functions that accept background rate arguments.
+Since background rates typically account for a small percentage of experimental rates, multiple “blank” experiments are routinely conducted alongside main experiments, and the rates are averaged across several datasets to obtain a more accurate estimate of the correction.
+The function `calc.bg.rate()` uses simple linear regression to simultaneously process multiple background rate measurements, as long as they share the time data, and produces an output object that can be used in later functions that accept background rate arguments.
 
@@
 
-A typical procedure in the analysis respirometry data is to select the most linear section of the data for calculations of $RO_2$ (e.g.
-citations).
+A typical procedure in the analysis respirometry data is to select one or more linear sections of the data for calculations of $RO_2$ (e.g.
+Gordon et al., 1989, Dorey et al., 2014, Chu & Gal, 2017).
 The function `calc.rate()` can manually extract one or more segments from a single data frame to process them simultaneously.
@@
 estimating $RO_2$ over a regular 10-minute time window across multiple specimens).
+Rates are determined using linear regression, however a simple delta approach is also included alongside all regression results, similar to a two-point analysis which is simply the difference in $O_2$ concentration divided by time.
 
-In addition, we can take advantage of automated methods in `auto.rate()`, which implements novel rolling regression and kernel density estimation algorithms to detect patterns in the data to automatically obtain the maximum or minimum rate of fixed width, or to detect the most linear sections of the data.
-Unlike manual methods where data is most likely arbitrary selected, `auto.rate()` adheres to well-defined rules to obtain the rate of interest, so its methods are results are transparent and fully reproducible.
-When called, the function takes the data frame of length $m$, and performs a rolling regression of sample size $n$ iteratively across the entire data frame (Fig.
+In addition, we can take advantage of automated methods in `auto.rate()`, which implements rolling regression and kernel density estimation algorithms to detect patterns in the data.
+From the output we can automatically obtain the maximum or minimum rate of a specified sample size, extract regressions at non-overlapping intervals (Fig.
+2a), or detect and rank linear sections of the data by sample size.
+Unlike manual methods, where data is, likely, arbitrary selected, `auto.rate()` adheres to well-defined rules to obtain the rate of interest, so its methods are results are transparent and fully reproducible.
+When called, the function takes the data frame of length $m$, and performs a rolling regression of sample size $n$ iteratively across the length of the data (Fig.
 2b).
-Thus, a total of $(m−n)+1$ number of iterative regressions are fitted, and we obtain a new dataset of regressions as a function of time (Fig.
+Thus, a total of $(m−n)+1$ number of overlapping regressions are fitted, and we obtain a new dataset of regressions as a function of time (rolling regression, Fig.
 3), which is then ranked by size to obtain maximum and minimum values.
@@
 Logically, linear sections of the data should reflect a consistently stable (i.e.
-flat) value in the rolling regression data (Fig.
-3b).
-The function `auto.rate()` takes advantage of this relationship to identify data windows that contain consistent rates over a rolling window, implementing the methods of Sheather & Jones (1991), avaliable in the `stats` package in base R, to select the bandwidth using pilot estimation of derivatives and calculate "peaks" in the density that correspond to stable rate values (Fig.
-3c).
-The function then selects the rolling data windows that fall within the kernel density binwidth, merges all overlapping windows into one, and performs a linear regression on each new window.
+flat) value in the rolling regression data.
+The function `auto.rate()` takes advantage of this relationship to identify consistent rates by performing a kernel density estimate of the data, implementing the methods of Sheather & Jones (1991) -- available in the `stats` package in R -- to derive the bandwidth used to construct the "peaks" in the density that correspond to stable rate values.
+The bandwidth is then used to select the range of overlapping regressions around the peak, and the single section of the data is resampled once using linear regression.
+
+## Flowthrough respirometry
+
+Calculating oxygen consumption using flowthrough respirometry requires different inputs of incurrent and excurrent oxygen concentration and a measure of the rate of flow through the respirometer.
+The function `calc.rate.ft`
 

### 2017-10-18 manuscript/manuscript-HTML.Rmd
manuscript update

@@
     fig_caption: true
-    pandoc_args: ["-F","pandoc-crossref"]
-    theme: "united"
+    theme: "default"
 ---
@@
 The oxygen recordings may be continuous through use of an oxygen probe, periodic through withdrawing water or gas samples at set intervals, or a two-point measurement consisting of the initial and final concentrations.
-Metabolic rates are estimated from respirometry by assuming a linear relationship between variables, and estimates of metabolic rate are straightforward in constant volume respirometry using the equation: $$RO_2 = \Delta O_2(V)$$ where $\Delta O_2$ is the slope of the regression that relates O~2~ concentration to time, or in the case of a two-point measurement, the difference in $O_2$ concentration divided by time, and $V$ is the volume of the container (Lighton, 2008).
+Metabolic rates are estimated from respirometry by assuming a linear relationship between variables, and estimates of metabolic rate are straightforward in constant volume respirometry using the equation: $$VO_2 = \Delta O_2V$$ where $\Delta O_2V$ is the slope of the regression that relates to the rate of change in $O_2$ concentration over time, or in the case of a two-point measurement, the difference in $O_2$ concentration divided by time elapsed, and $V$ is the volume of the container (Lighton, 2008).
+Note, we use here and in the package the term $RO_2$ to denote rate of $O_2$ depletion or uptake within an experiment.
+We use this term used to distinguish it from other common terms, for example $VO_2$ which is can be used to denote both volumetric O2 uptake and whole animal uptake, or $MO_2$ which typically denotes mass-specific uptake.
 
@@
 Gamble et al., 2014).
-The common equation used for open respirometry is: $$RO_2 = \Delta O_2(V) + \phi_d$$ where $\Delta O_2$ is the slope of the regression that relates $O_2$ concentration to time, $V$ is the volume of the container and $\phi_d$ is the air-sea oxygen flux as determined by Fick's Law (Leclercq et al., 1999).
+The common equation used for open respirometry is: $$VO_2 = \Delta O_2V + \phi_d$$ where $\Delta O_2V$ is the slope of the regression that relates $O_2$ concentration to time, $V$ is the volume of the container and $\phi_d$ is the air-sea oxygen flux as determined by Fick's Law (Leclercq et al., 1999).
 
@@
 1).
-The package contains utilities to: (1) analyse closed, intermittent, flow-through and open respirometry data, (2) scale $\Delta O_2$ to volume and/or mass, and (3) automatically and rapidly detect MMR, SMR, $P_{c}$ or best-fit results in large datasets with the help of traditional rolling regression and kernel density estimation techniques.
+The package contains utilities to: (1) analyse closed, intermittent, flow-through and open respirometry data, (2) scale $\Delta O_2$ to volume and/or mass, and (3) automatically detect MMR, SMR, $P_{c}$ or most linear results in large datasets with the help of traditional rolling regression and kernel density estimation techniques.
 Other smaller, but useful, functions are also avaliable, and are described in more detail in our online html vignette (https://januarharianto.github.io/respR/).
@@
 To explore `respR` and its main functions we have provided example data that can be called directly after the package is loaded.
-The first data object, `urchin2013`, contains measurements of oxygen consumption in 16 individual sea urchins (*Heliocidaris erythrogramma*) and 2 "blank" measurements of background respiration (Harianto, *unpublished*).
-The `intermittent` data object contains three repeated measures of oxygen consumption in *H.
-erythrogramma*, collected in a single session with intermittent respirometry (Carey et al.
-2016, *JEB*).
-The `flowthrough` data object contains measurements of oxygen consumption, via flow-through respirometry, in a species of chiton, *Mopalia lignosa* (Carey, *unpublished*).
 Detailed information about all example data, including their source and methods, can be obtained with the `?` command in the R console (e.g.
 `?urchin2013`).
+All analyses described in this paper use the example data provided and are fully reproducible (see supplementary information).
 
@@
 `POSIXct` and `POSIXlt` classes) and performing calculations on different time formats have many caveats that are beyond the scope of the package.
-Time data should be sequential, without duplicates, as such errors strongly indicate underlying sampling issues.
-Ideally, time data should also be evenly-spaced, that is, the time period between samples are equal.
-This has important ramifications on data extraction techniques later on, as substantial errors may occur in the event that calculations are based on the number of rows instead of time.
-A plot of the data is also shown for visual inspection (e.g.
+Time data, being a continuous variable, should be sequential, without duplicates.
+Ideally, time data should also be evenly-spaced, that is, the time period between samples are equal, as this has important ramifications on data extraction methods later on and substantial errors may occur if data is analysed by rows instead of time.
+The function summarises error checks in the R console (see supplementary material) and plot of the data is shown for visual inspection (e.g.
 Fig.
@@
 
-It should be noted that invoking `check.input` is optional -- the main functions in our package will readily accept any data frame as long as data are numeric.
+It should be noted that invoking `check.input()` is optional -- the main functions in our package will readily accept any data frame as long as data are numeric.
 Running `check.input` is an exploratory step that flags potential issues about the data before it is further analysed.
 
-## Estimating rates of change in $O_2$ concentration ($RO_2$)
+## Estimating $RO_2$
 
-The functions `calc.bg.rate()`, `calc.rate()`, `auto.rate()`, `calc.rate.ft()` and `pcrit()` can process raw respirometry data and are called depending on the experimental setup and type of analysis required.
+The functions `calc.bg.rate()`, `calc.rate()`, `auto.rate()`, `calc.rate.ft()` and `pcrit()` are used to process raw respirometry data.
+For delibrate reasons, units of measurements are not specified in any of these functions.
+Instead, they are specified only during unit conversions in `convert.do()`, or when units are scaled to volume and mass in `calc.mo2()`.
+Separating the
 
-### Measuring background changes in $RO_2$
+### Analysing background changes in $RO_2$
 
-Background $RO_2$ is measured to account for the influence of bacterial respiration in the chamber, and once measured, is often applied as a correction to the main experiment (Rodgers et al., 2016).
-Since background rates typically account for a small percentage of experimental rates, multiple “blank” experiments are routinely conducted alongside main experiments, and the rates are averaged across several datasets to obtain a more accurate estimate of the correction.
-The function `calc.bg.rate()` uses simple linear regression to simultaneously process multiple background rate measurements, as long as they share the time data, and produces an output object that can be used in later functions that accept background rate arguments.
+Background $RO_2$ is often measured to account for the influence of bacterial respiration in the chamber, and applied as a correction to the main experiment (Rodgers et al., 2016).
+Since background rates typically account for a small percentage of experimental rates, multiple “blank” experiments are routinely conducted and the rates are averaged across several datasets to obtain a more accurate estimate of the correction (e.g.
+Carey and Sigwart, 2014; Daoud et al., 2007).
+The function `calc.bg.rate()` uses simple linear regression to simultaneously process multiple background rate measurements and produces an output object that can be used in later functions that accept background rate arguments.
+Because the oxygen consumption units are non-volume and non-mass specific at the point of correction, the functions in `respR` can adjust $RO_2$ directly without bias: $$RO_{2(correction)} = RO_2 - RO_{2(bg)}$$
 
-### Manual and automatic selection of data for calculations
+### Manual selection of data
 
-A typical procedure in the analysis respirometry data is to select one or more linear sections of the data for calculations of $RO_2$ (e.g.
+Typically, the analysis respirometry data requires the selection one or more linear sections of the raw data for more accurate calculations of $RO_2$ (e.g.
 Gordon et al., 1989, Dorey et al., 2014, Chu & Gal, 2017).
-The function `calc.rate()` can manually extract one or more segments from a single data frame to process them simultaneously.
+The function `calc.rate()` can manually extract and process data segments from a `data.frame` object.
 Data segments can be selected by (1) time period, (2) row numbers, (3) $O_2$ decrease or (4) proportion, which should accomodate most, if not all, data selection requirements and allow for consistent reporting of methods and results (e.g.
 estimating $RO_2$ over a regular 10-minute time window across multiple specimens).
-Rates are determined using linear regression, however a simple delta approach is also included alongside all regression results, similar to a two-point analysis which is simply the difference in $O_2$ concentration divided by time.
+Rates are determined using linear regression, however a simple two-point analysis (equation 1) is also included alongside all regression results.
 
-In addition, we can take advantage of automated methods in `auto.rate()`, which implements rolling regression and kernel density estimation algorithms to detect patterns in the data.
-From the output we can automatically obtain the maximum or minimum rate of a specified sample size, extract regressions at non-overlapping intervals (Fig.
-2a), or detect and rank linear sections of the data by sample size.
-Unlike manual methods, where data is, likely, arbitrary selected, `auto.rate()` adheres to well-defined rules to obtain the rate of interest, so its methods are results are transparent and fully reproducible.
+### Automatic selection of data
+
+The function `auto.rate()` uses rolling regression and kernel density estimate algorithms to detect patterns in respirometry data.
+It can calculate maximum or minimum rates, extract rates at non-overlapping intervals, and has an exploratory method to detect the most linear sections of the data.
+
+There are clear advantages to using `auto.rate()` to estimate MMR and SMR.
+Unlike manual methods, where data is often arbitrary selected and inconvenient to replicate, `auto.rate()` adheres to well-defined rules to obtain the rate of interest, so its methods are results are transparent and fully reproducible.
 When called, the function takes the data frame of length $m$, and performs a rolling regression of sample size $n$ iteratively across the length of the data (Fig.
@@
 Thus, a total of $(m−n)+1$ number of overlapping regressions are fitted, and we obtain a new dataset of regressions as a function of time (rolling regression, Fig.
-3), which is then ranked by size to obtain maximum and minimum values.
+3b), which is then ranked by size to obtain maximum and minimum values.
 
-To estimate the most linear rate, further computations are performed after rolling regressions.
-Logically, linear sections of the data should reflect a consistently stable (i.e.
-flat) value in the rolling regression data.
-The function `auto.rate()` takes advantage of this relationship to identify consistent rates by performing a kernel density estimate of the data, implementing the methods of Sheather & Jones (1991) -- available in the `stats` package in R -- to derive the bandwidth used to construct the "peaks" in the density that correspond to stable rate values.
-The bandwidth is then used to select the range of overlapping regressions around the peak, and the single section of the data is resampled once using linear regression.
+While using rolling regressions to estimate MMR and SMR
 
-## Flowthrough respirometry
+### Analysing flowthrough respirometry data
 
-Calculating oxygen consumption using flowthrough respirometry requires different inputs of incurrent and excurrent oxygen concentration and a measure of the rate of flow through the respirometer.
-The function `calc.rate.ft`
+The function `calc.rate.ft()` is used to analyse flowthrough respirometry data, and accepts inputs of incurrent and excurrent oxygen concentration, and flow rate.
+It is crucial that all the input units match as the function does not perform unit conversions (but see `calc.mo2()`), and the output gives $RO_2$.
+Like `calc.rate()`, a segment of the data can be selected before processing and the average, or median $VO_2$ are provided.
+
+## Converting units of $RO_2$ to $VO_2$ and $MO_2$
+
+Once
 
@@
 
-## Converting units of $RO_2$ and $MO_2$
-
-- talk about calc.mo2 and conversions to volume and weight specific $MO_2$.
+The function `pcrit()` provides a way to consistently report the parameters associated with estimating $P_c$.
+$P_c$ is often estimated using the "broken-stick" regression (BSR) approach (Yeager and Ultsch, 1989), although there are some cases where it is estimated without indication of how thie value was determined numerically (e.g.
+need examples).
+While in many cases the PCrit and oxyregulatory ability can be clear and obvious (e.g./refs), many studies still class species or individuals as oxyregulators or oxyconformers without recognising that there may be intermediate cases.
 

### 2017-10-18 manuscript/manuscript-HTML.Rmd
manuscript update

@@
 Time data, being a continuous variable, should be sequential, without duplicates.
-Ideally, time data should also be evenly-spaced, that is, the time period between samples are equal, as this has important ramifications on data extraction methods later on and substantial errors may occur if data is analysed by rows instead of time.
+Ideally, time data should also be evenly-spaced, that is, the time period between samples are equal, as this has important ramifications on data extraction methods later on, and substantial errors may occur if data is analysed by rows instead of time.
 The function summarises error checks in the R console (see supplementary material) and plot of the data is shown for visual inspection (e.g.
@@
 
-It should be noted that invoking `check.input()` is optional -- the main functions in our package will readily accept any data frame as long as data are numeric.
+It should be noted that invoking `check.input` is optional -- the main functions in our package will readily accept any data frame as long as data are numeric.
 Running `check.input` is an exploratory step that flags potential issues about the data before it is further analysed.
@@
 The functions `calc.bg.rate()`, `calc.rate()`, `auto.rate()`, `calc.rate.ft()` and `pcrit()` are used to process raw respirometry data.
-For delibrate reasons, units of measurements are not specified in any of these functions.
-Instead, they are specified only during unit conversions in `convert.do()`, or when units are scaled to volume and mass in `calc.mo2()`.
-Separating the
+For delibrate reasons, units of measurements are not required in any of these functions.
 

### 2017-10-18 manuscript/manuscript-HTML.Rmd
NC edits and comments

@@
 
-Metabolic rate is a fundamental trait linked to virtually all other biological functions, and is key to the prediction of patterns in ecology and conservation biology, from populations (Seibel and Drazen, 2007, Barneche et al., 2014) to ecosystems (Brown et al., 2004).
-It has been widely used to investigate the effects of external stressors on an organism, and has seen increased popularity in studies involving climate change where warming is expected to drive increased metabolism in ectotherms (Pörtner, 2002, McElroy et al., 2012, Carey et al., 2016, Delorme & Sewell, 2016).
-By far the most common metric used to measure physiological performance in animals, it can be applied from the whole organism to the level of cells and tissues (White & Kearney, 2013), and is typically quantified using respirometry, which measures oxygen uptake over time (Lighton, 2008).
+Metabolic rate is a fundamental trait associated with virtually all biological functions, and is key in predicting patterns in ecology and conservation biology, from populations (Seibel and Drazen, 2007, Barneche et al., 2014) to ecosystems (Brown et al., 2004).
+It has been widely used to investigate the effects of external stressors on organisms, and has seen increased interest in studies of climate change where warming is expected to drive increased metabolism in ectotherms (Pörtner, 2002, McElroy et al., 2012, Carey et al., 2016, Delorme & Sewell, 2016).
+By far the most common metric used to assess physiological performance in animals, metabolic rates are determined from whole organisms to the level of cells and tissues (White & Kearney, 2013).
+They are typically quantified using respirometry, which measures oxygen uptake over time as a proxy for metabolic rate, since oxygen is rarely stored and anaerobic metabolism is typically sustained over only brief periods (Lighton, 2008).
 
-There are at least four broad methodological approaches in respirometry: closed, intermittent flow, flowthrough and open.
-In closed respirometry, $O_2$ decrease is measured within a hermetically sealed chamber of known volume, sometimes set within a closed loop to allow circulation or mixing of the environment within the chamber.
-The oxygen recordings may be continuous through use of an oxygen probe, periodic through withdrawing water or gas samples at set intervals, or a two-point measurement consisting of the initial and final concentrations.
-Metabolic rates are estimated from respirometry by assuming a linear relationship between variables, and estimates of metabolic rate are straightforward in constant volume respirometry using the equation: $$VO_2 = \Delta O_2V$$ where $\Delta O_2V$ is the slope of the regression that relates to the rate of change in $O_2$ concentration over time, or in the case of a two-point measurement, the difference in $O_2$ concentration divided by time elapsed, and $V$ is the volume of the container (Lighton, 2008).
-Note, we use here and in the package the term $RO_2$ to denote rate of $O_2$ depletion or uptake within an experiment.
-We use this term used to distinguish it from other common terms, for example $VO_2$ which is can be used to denote both volumetric O2 uptake and whole animal uptake, or $MO_2$ which typically denotes mass-specific uptake.
+There are four broad methodological approaches to respirometry: closed, intermittent flow, flowthrough and open.
+In closed respirometry, $O_2$ decrease is measured within a hermetically sealed chamber of known volume, sometimes set within a closed loop to allow mixing of the environment within the chamber.
+Oxygen recordings may be continuous through use of an oxygen probe, periodic through withdrawing water or gas samples at set intervals, or a two-point measurement consisting of the initial and final concentrations.
+Metabolic rates are estimated from respirometry data by assuming a linear relationship between variables, and estimates of metabolic rate are straightforward in constant volume respirometry using the equation: $$VO_2 = \Delta O_2V$$ where $\Delta O_2V$ is the slope of the regression that describes the rate of change in $O_2$ concentration over time, or in the case of a two-point measurement, the difference in $O_2$ concentration divided by time elapsed, and $V$ is the volume of the container (Lighton, 2008).
+~~Note, we use here and in the package the term $RO_2$ to denote rate of $O_2$ depletion or uptake within an experiment.
+We use this term used to distinguish it from other common terms, for example $VO_2$ which is can be used to denote both volumetric O2 uptake and whole animal uptake, or $MO_2$ which typically denotes mass-specific uptake~~ [nc - don't need this anymore.
+I had it when we used that term in my functions].
 
-In intermittent flow respirometry, $O_2$ concentration is measured, but after a set time period the chamber is flushed with new water or air, returning it to initial conditions, sealed again, and the experiment repeated (Svendsen et al., 2016).
+In intermittent flow respirometry, $O_2$ concentration is measured as described above, but periodically the chamber is flushed with new water or air, returning it to initial conditions, resealed, and the experiment repeated (Svendsen et al., 2016).
 This technique is essentially closed respirometry, but with the incorporation of repeated measures.
-Depending on the metabolic rate metric being investigated, final respiration rate is calculated as the mean of the measures (e.g.
-citation), or the lowest or highest rates recorded in any trial (e.g.
-Carey et al., 2016).
+Depending on the metabolic rate metric being investigated, final respiration rate can be calculated as the mean of the measures (e.g.
+Carey et al., 2016), or the lowest or highest rates recorded in any trial (e.g.
+Stoffels, 2015).
+[nc-will try and find a better ref for that]
+
+Stoffels, R.
+J.
+(2015).
+Physiological trade-offs along a fast-slow lifestyle continuum in fishes: What do they tell us about resistance and resilience to hypoxia?
+Plos One, 10(6), e0130303.
+https://doi.org/10.1371/journal.pone.0130303
 
@@
 Gamble et al., 2014).
-The common equation used for open respirometry is: $$VO_2 = \Delta O_2V + \phi_d$$ where $\Delta O_2V$ is the slope of the regression that relates $O_2$ concentration to time, $V$ is the volume of the container and $\phi_d$ is the air-sea oxygen flux as determined by Fick's Law (Leclercq et al., 1999).
+The common equation used for open respirometry is: $$VO_2 = \Delta O_2V + \phi_d$$ where $\Delta O_2V$ is the slope of the regression that relates $O_2$ concentration to time, $V$ is the volume of the container and $\phi_d$ is the oxygen flux as determined by Fick's Law (Leclercq et al., 1999).
 
-Researchers are increasingly assimilating large, high-resolution respirometry data, often capturing multiple response variables such as maximum metabolic rate (MMR, or $MO_{2,max}$), the rate under high activity or exhaustive exercise, or standard (or resting) metabolic rate (SMR, or $MO_{2,min}$), the minimum metabolic cost of maintaining biological functioning (Rogers et al., 2016), or critical oxygen tension, which represents the lowest level of oxygen at which aerobic metabolism is independent of the ambient partial pressure of oxygen ($P_{c}$; Yeager & Ultsch, 1989, Hochachka and Somero, 2002).
+[nc - so the following needs to be spot on.
+Physiologists get very worked up (if not confused themselves) about the differences between these, esp.
+SMR v Routine v Resting v Basal!
+Especially those who work on ecto vs endo therms]
+
+Depending on the experiment, different metabolic rate metrics may be of interest to researchers.
+These include maximum metabolic rates (MMR, or $MO_{2,max}$), rates under high activity or exhaustive exercise, and minimal metabolic rates $MO_{2,min}$, which depending on the organism may be termed standard (SMR), basal (BMR), or resting metabolic rates, and typically represent the minimum metabolic cost of maintaining biological functioning (White and Kearney, 2014) [much better ref for all this].
+Routine metabolic rate (RMR), is similar but takes into account that in some organisms energy is expended on small, sponateous movements to maintain posture or position (Rogers et al.
+2016).
+Another metric of great interest is the critical oxygen tension, which represents the lowest level of oxygen at which aerobic metabolism is independent of the ambient partial pressure of oxygen ($P_{crit}$; Yeager & Ultsch, 1989, Hochachka and Somero, 2002).
+
+White, C.
+R., & Kearney, M.
+R.
+(2014).
+Metabolic scaling in animals: methods, empirical results, and theoretical explanations.
+Comprehensive Physiology, 4(1), 231–256.
+https://doi.org/10.1002/cphy.c110049
+
+Researchers are increasingly assimilating large, high-resolution respirometry datasets.
 Experiments may run for long periods (e.g.
-20 h, Norin & Malte, 2012), and in most cases, processing the data involves an *ad hoc* selection of data points and them manually performing calculations in a spreadsheet program (e.g.
+20 h, Norin & Malte, 2012), and in most cases, processing the data involves an *ad hoc* selection of data points under vague criteria, with subsequent manual processing in a spreadsheet program (e.g.
 Microsoft Excel) or an integrated development environment (IDE, e.g.
 R and Matlab).
-These approaches can be tedious and time consuming especially when spreadsheet programs (e.g.
-Excel) struggle with the vast datasets that are generated, while IDEs require a degree of expertise and have substantial learning curves.
-Dedicated software are available to perform such calculations, but most are proprietary and require licensing requirements or dedicated hardware (e.g.
-AutoResp by Loligo Systems), complicating or preventing their use on multiple machines.
+These approaches can be tedious and time consuming especially when spreadsheet programs (in particular Excel) struggle with the vast datasets that are generated, while IDEs require a degree of expertise and have substantial learning curves.
+Dedicated software are available to perform such analyses, but many have expensive licensing requirements (e.g.
+AutoResp by Loligo Systems), complicating or preventing their use on multiple machines, and are proprietary and closed-source hindering scientific reproducability and transparency.
 
-Recently, a number of free R software packages designed for respirometry have become available (e.g.
-`respirometry`: https://CRAN.R-project.org/package=respirometry, `rMR`: https://CRAN.R-project.org/package=rMR, `LoLinR`: Olito et al., 2017).
-Both `respirometry` and `rMR` are useful for aquatic respirometry, but are not focused on the analysis of data, although both packages have functions for the analysis of metabolic rate over non-overlapping intervals, and `rMR` has a specialised function to calculate $P_{c}$.
-For `LoLinR`, the package provides a statistically robust method of detecting a "best-fit" regression which is applicable for closed and flow-through respirometry.
-However, `LoLinR`'s processing time increases exponentially with large datasets, with data containing more than 500 observations taking a few minutes to run (Olito et al., 2017), which makes `LoLinR` intractable with large amounts of data.
+[r.e. below para. LoLinR is not strictly for respirometry - just monotonic, something-changing-over-time data]
 
-Here we describe the R package `respR`, a set of functions desgined to provide a structured workflow for the analysis of respirometry-related data (Fig.
+Recently, a number of open-source R software packages have become available (e.g.
+`respirometry`: https://CRAN.R-project.org/package=respirometry, `rMR`: https://CRAN.R-project.org/package=rMR, `LoLinR`: Olito et al., 2017) which are suitable for or designed to analyse respirometry data.
+~~ Both `respirometry` and `rMR` are useful for aquatic respirometry, but are not focused on the analysis of data, although both packages have functions for the analysis of metabolic rate over non-overlapping intervals, and `rMR` has a specialised function to calculate $P_{c}$.
+~~ [nc - need a better way of basically saying, they're ok, but very limited, and not user-friendly, or adaptable, or].
+`LoLinR` provides a very statistically robust method of detecting a "best-fit" regression which is applicable for closed and flow-through respirometry.
+[nc - try to add something else positive about it!
+We need to be nice to them].
+However, `LoLinR`'s processing time increases exponentially with large datasets, with data containing more than 500 observations taking several minutes to run (Olito et al., 2017), which makes `LoLinR` somewhat intractable with large amounts of data.
+In addition, its focus is on data which is evenly spaced with regard to time, and where the response variable changes in a monotonic fashion, which is far from typical of respirometry data.
+
+Here we describe the R package `respR`, a set of functions desgined to provide a structured, reproduceable workflow for the analysis of respirometry-related data (Fig.
 1).
-The package contains utilities to: (1) analyse closed, intermittent, flow-through and open respirometry data, (2) scale $\Delta O_2$ to volume and/or mass, and (3) automatically detect MMR, SMR, $P_{c}$ or most linear results in large datasets with the help of traditional rolling regression and kernel density estimation techniques.
-Other smaller, but useful, functions are also avaliable, and are described in more detail in our online html vignette (https://januarharianto.github.io/respR/).
-While the package has a strong focus on marine and freshwater respirometry, the principals also apply to aerial respirometry.
-Some oxygen concentration conversions require a salinity value, but in freshwater respirometry this should be set to zero.
+The package contains utilities to: (1) analyse closed, intermittent, flow-through and open respirometry data, (2) scale $\Delta O_2$ to volume and/or mass, and surface area [nc - soon yes?] (3) automatically detect maximal and minimal rates, $P_{c}$, or the most linear results in large datasets with the help of traditional rolling regression and kernel density estimation techniques.
+Other functions useful in processing respirometry data are avaliable, and are described in more detail in our online html vignette (https://januarharianto.github.io/respR/).
+While the package has a strong focus on aquatic respirometry, similar principals apply to aerial respirometry.
+~~Some oxygen concentration conversions require a salinity value, but in freshwater respirometry this should be set to zero.~~ [nc - don't need this, can go in the help docs].
 We demonstrate the utility of the package by analysing three datasets as case studies and proof of concept.
 These data were collected using closed, intermittent and flow-through respirometry.
+
+[nc - i would drop closed, focus on intermittent (which is same as closed essentially) and flowthrough, and include a PCrit case.
+It's a major selling point!
+Several people i have told about it have asked - does it do PCrit?]
 

### 2017-10-18 manuscript/manuscript-HTML.Rmd
NC - a few additional edits

@@
 `respirometry`: https://CRAN.R-project.org/package=respirometry, `rMR`: https://CRAN.R-project.org/package=rMR, `LoLinR`: Olito et al., 2017) which are suitable for or designed to analyse respirometry data.
-~~ Both `respirometry` and `rMR` are useful for aquatic respirometry, but are not focused on the analysis of data, although both packages have functions for the analysis of metabolic rate over non-overlapping intervals, and `rMR` has a specialised function to calculate $P_{c}$.
-~~ [nc - need a better way of basically saying, they're ok, but very limited, and not user-friendly, or adaptable, or].
+~~Both `respirometry` and `rMR` are useful for aquatic respirometry, but are not focused on the analysis of data, although both packages have functions for the analysis of metabolic rate over non-overlapping intervals, and `rMR` has a specialised function to calculate $P_{c}$.~~ [nc - need a better way of basically saying, they're ok, but very limited, and not user-friendly, or adaptable, or].
 `LoLinR` provides a very statistically robust method of detecting a "best-fit" regression which is applicable for closed and flow-through respirometry.
@@
 
-Here we describe the R package `respR`, a set of functions desgined to provide a structured, reproduceable workflow for the analysis of respirometry-related data (Fig.
+Here we describe the R package `respR`, a set of functions desgined to provide a fast [or better word], structured, reproduceable workflow for the analysis of respirometry-related data (Fig.
 1).
-The package contains utilities to: (1) analyse closed, intermittent, flow-through and open respirometry data, (2) scale $\Delta O_2$ to volume and/or mass, and surface area [nc - soon yes?] (3) automatically detect maximal and minimal rates, $P_{c}$, or the most linear results in large datasets with the help of traditional rolling regression and kernel density estimation techniques.
+The package contains utilities to: (1) analyse closed, intermittent, flow-through and open respirometry data, (2) determine volume-, mass-, and surface area- [nc - soon yes?] specific oxygen uptake rates, (3) automatically detect maximal and minimal rates, $P_{c}$, or the most linear results in large datasets with the help of traditional rolling regression and kernel density estimation techniques.
 Other functions useful in processing respirometry data are avaliable, and are described in more detail in our online html vignette (https://januarharianto.github.io/respR/).
-While the package has a strong focus on aquatic respirometry, similar principals apply to aerial respirometry.
+The package has a strong focus on aquatic respirometry, however similar principals apply to aerial respirometry, so the analysis workflow should also be suitable for these data.
 ~~Some oxygen concentration conversions require a salinity value, but in freshwater respirometry this should be set to zero.~~ [nc - don't need this, can go in the help docs].
@@
 It's a major selling point!
-Several people i have told about it have asked - does it do PCrit?]
+Several people i have told about it have asked - does it do PCrit?
+Or i guess, make the closed e.g.
+the PCrit one...
+]
 

### 2017-10-19 manuscript/manuscript-HTML.Rmd
manuscript update - with edits included

@@
 Metabolic rates are estimated from respirometry data by assuming a linear relationship between variables, and estimates of metabolic rate are straightforward in constant volume respirometry using the equation: $$VO_2 = \Delta O_2V$$ where $\Delta O_2V$ is the slope of the regression that describes the rate of change in $O_2$ concentration over time, or in the case of a two-point measurement, the difference in $O_2$ concentration divided by time elapsed, and $V$ is the volume of the container (Lighton, 2008).
-~~Note, we use here and in the package the term $RO_2$ to denote rate of $O_2$ depletion or uptake within an experiment.
-We use this term used to distinguish it from other common terms, for example $VO_2$ which is can be used to denote both volumetric O2 uptake and whole animal uptake, or $MO_2$ which typically denotes mass-specific uptake~~ [nc - don't need this anymore.
-I had it when we used that term in my functions].
+<!-- Note, we use here and in the package the term $RO_2$ to denote rate of $O_2$ depletion or uptake within an experiment.
+We use this term used to distinguish it from other common terms, for example $VO_2$ which is can be used to denote both volumetric O2 uptake and whole animal uptake, or $MO_2$ which typically denotes mass-specific uptake.
+-->
 
@@
 Stoffels, 2015).
-[nc-will try and find a better ref for that]
+<!--[nc-will try and find a better ref for that]
 
@@
 https://doi.org/10.1371/journal.pone.0130303
+-->
 
@@
 
-[nc - so the following needs to be spot on.
+<!--[nc - so the following needs to be spot on.
 Physiologists get very worked up (if not confused themselves) about the differences between these, esp.
 SMR v Routine v Resting v Basal!
-Especially those who work on ecto vs endo therms]
+Especially those who work on ecto vs endo therms]-->
 
-Depending on the experiment, different metabolic rate metrics may be of interest to researchers.
-These include maximum metabolic rates (MMR, or $MO_{2,max}$), rates under high activity or exhaustive exercise, and minimal metabolic rates $MO_{2,min}$, which depending on the organism may be termed standard (SMR), basal (BMR), or resting metabolic rates, and typically represent the minimum metabolic cost of maintaining biological functioning (White and Kearney, 2014) [much better ref for all this].
-Routine metabolic rate (RMR), is similar but takes into account that in some organisms energy is expended on small, sponateous movements to maintain posture or position (Rogers et al.
+Depending on the experiment, different metabolic rate parameters may be of interest to researchers.
+These include maximum metabolic rates (MMR, or $MO_{2,max}$), rates under high activity or exhaustive exercise, and minimal metabolic rates ($MO_{2,min}$), which, depending on the organism, may be termed standard (SMR), basal (BMR), or resting metabolic rates, and typically represent the minimum metabolic cost of maintaining biological functioning (White and Kearney, 2014).
+Routine metabolic rate (RMR), is similar to $MO_{2,min}$, but takes into account that in some organisms energy is expended on small, sponateous movements to maintain posture or position (Rogers et al.
 2016).
@@
 
-White, C.
+<!--White, C.
 R., & Kearney, M.
@@
 Comprehensive Physiology, 4(1), 231–256.
-https://doi.org/10.1002/cphy.c110049
+https://doi.org/10.1002/cphy.c110049-->
 
-Researchers are increasingly assimilating large, high-resolution respirometry datasets.
-Experiments may run for long periods (e.g.
-20 h, Norin & Malte, 2012), and in most cases, processing the data involves an *ad hoc* selection of data points under vague criteria, with subsequent manual processing in a spreadsheet program (e.g.
+Respirometry experiments are increasingly assimilating large, high-resolution datasets and are run for longer periods (e.g.
+20 h, Norin & Malte, 2012).
+In most cases, processing the data involves an *ad hoc* selection of data points with undefined criteria, and subsequent manual processing of the data subset(s) using a spreadsheet program (e.g.
 Microsoft Excel) or an integrated development environment (IDE, e.g.
-R and Matlab).
-These approaches can be tedious and time consuming especially when spreadsheet programs (in particular Excel) struggle with the vast datasets that are generated, while IDEs require a degree of expertise and have substantial learning curves.
-Dedicated software are available to perform such analyses, but many have expensive licensing requirements (e.g.
-AutoResp by Loligo Systems), complicating or preventing their use on multiple machines, and are proprietary and closed-source hindering scientific reproducability and transparency.
+R or Matlab).
+These approaches can be tedious and time consuming especially when spreadsheet programs struggle with the vast datasets that are generated, while IDEs require a degree of expertise to use and have substantial learning curves.
+Dedicated software are also available to perform metabolic rate analyses, but many have costly licensing requirements (e.g.
+AutoResp by Loligo Systems), complicating or preventing their use on multiple machines, and are proprietary and closed-source, hindering scientific reproducability and transparency.
 
-[r.e. below para. LoLinR is not strictly for respirometry - just monotonic, something-changing-over-time data]
+<!-- r.e. below para. LoLinR is not strictly for respirometry - just monotonic, something-changing-over-time data -->
 
-Recently, a number of open-source R software packages have become available (e.g.
-`respirometry`: https://CRAN.R-project.org/package=respirometry, `rMR`: https://CRAN.R-project.org/package=rMR, `LoLinR`: Olito et al., 2017) which are suitable for or designed to analyse respirometry data.
-~~Both `respirometry` and `rMR` are useful for aquatic respirometry, but are not focused on the analysis of data, although both packages have functions for the analysis of metabolic rate over non-overlapping intervals, and `rMR` has a specialised function to calculate $P_{c}$.~~ [nc - need a better way of basically saying, they're ok, but very limited, and not user-friendly, or adaptable, or].
-`LoLinR` provides a very statistically robust method of detecting a "best-fit" regression which is applicable for closed and flow-through respirometry.
-[nc - try to add something else positive about it!
-We need to be nice to them].
-However, `LoLinR`'s processing time increases exponentially with large datasets, with data containing more than 500 observations taking several minutes to run (Olito et al., 2017), which makes `LoLinR` somewhat intractable with large amounts of data.
-In addition, its focus is on data which is evenly spaced with regard to time, and where the response variable changes in a monotonic fashion, which is far from typical of respirometry data.
+Recently, a number of open-source R software packages have become available which are designed for, or are suitable to analyse respirometry data.
+The `respirometry` package (https://CRAN.R-project.org/package=respirometry) contains a comprehensive collection of tools to explore and evaluate experimental parameters in aquatic respirometry and is useful for the design and diagnosis of experimental setups.
+The `rMR` package (https://CRAN.R-project.org/package=rMR) performs interval-based metabolic rate calculations and has a method to automatically detect $P_{crit}$ using the "broken-stick" regression method (Yeager & Ultsch, 1989).
+`LoLinR` (Olito et al., 2017), while not strictly coded with respirometry in mind, provides a statistically robust method of detecting a "best-fit" regression that applies to the calculation of metabolic rate, and performs very well at identifying truly linear subsets of a data series.
+Importantly, the packages above all assume that data are monotonic with evenly-spaced sampling periods with respect to time and do not have the tools to analyse complicated datasets that require extraction of multiple respirometry parameters.
 
-Here we describe the R package `respR`, a set of functions desgined to provide a fast [or better word], structured, reproduceable workflow for the analysis of respirometry-related data (Fig.
+Here we describe the R package `respR`, a set of functions desgined to provide an efficient, structured, and reproduceable workflow for the analysis of respirometry-related data (Fig.
 1).
-The package contains utilities to: (1) analyse closed, intermittent, flow-through and open respirometry data, (2) determine volume-, mass-, and surface area- [nc - soon yes?] specific oxygen uptake rates, (3) automatically detect maximal and minimal rates, $P_{c}$, or the most linear results in large datasets with the help of traditional rolling regression and kernel density estimation techniques.
+The package contains utilities to: (1) analyse closed, intermittent, flow-through and open respirometry data, (2) determine volume, mass, and surface area specific oxygen uptake rates, (3) automatically detect maximal and minimal rates, $P_{crit}$, or the most linear results in large datasets with the help of traditional rolling regression and kernel density estimation techniques.
 Other functions useful in processing respirometry data are avaliable, and are described in more detail in our online html vignette (https://januarharianto.github.io/respR/).
 The package has a strong focus on aquatic respirometry, however similar principals apply to aerial respirometry, so the analysis workflow should also be suitable for these data.
-~~Some oxygen concentration conversions require a salinity value, but in freshwater respirometry this should be set to zero.~~ [nc - don't need this, can go in the help docs].
 We demonstrate the utility of the package by analysing three datasets as case studies and proof of concept.
-These data were collected using closed, intermittent and flow-through respirometry.
-
-[nc - i would drop closed, focus on intermittent (which is same as closed essentially) and flowthrough, and include a PCrit case.
-It's a major selling point!
-Several people i have told about it have asked - does it do PCrit?
-Or i guess, make the closed e.g.
-the PCrit one...
-]
+These data were collected using closed, intermittent and flow-through respirometry, and we show how `respR` might be used to determine $MO_{2,max}$, $MO_{2,min}$, and $P_{crit}$.
 
@@
 Time data, being a continuous variable, should be sequential, without duplicates.
-Ideally, time data should also be evenly-spaced, that is, the time period between samples are equal, as this has important ramifications on data extraction methods later on, and substantial errors may occur if data is analysed by rows instead of time.
-The function summarises error checks in the R console (see supplementary material) and plot of the data is shown for visual inspection (e.g.
+Ideally, time data should also be evenly-spaced, that is, the time period between samples are equal, as this will affect data-extraction methods later on if data is analysed by sample length instead of time.
+Error check results are summarised in the R console (see supplementary material) and plot of the data is shown for visual inspection (e.g.
 Fig.
@@
 
-## Estimating $RO_2$
+## Estimating the rate of change in $O_2$ concentration ($RO_2$)
 
-The functions `calc.bg.rate()`, `calc.rate()`, `auto.rate()`, `calc.rate.ft()` and `pcrit()` are used to process raw respirometry data.
-For delibrate reasons, units of measurements are not required in any of these functions.
-
-### Analysing background changes in $RO_2$
+### Calculating background $RO_2$
 
@@
 
-### Manual selection of data
+### Extracting data for $RO_2$
 
@@
 Gordon et al., 1989, Dorey et al., 2014, Chu & Gal, 2017).
-The function `calc.rate()` can manually extract and process data segments from a `data.frame` object.
+The function `calc.rate()` can manually extract and process data segments from respiromety data.
 Data segments can be selected by (1) time period, (2) row numbers, (3) $O_2$ decrease or (4) proportion, which should accomodate most, if not all, data selection requirements and allow for consistent reporting of methods and results (e.g.
 estimating $RO_2$ over a regular 10-minute time window across multiple specimens).
-Rates are determined using linear regression, however a simple two-point analysis (equation 1) is also included alongside all regression results.
+Rates are determined using linear regression, however a two-point analysis is also included alongside all regression results.
 
-### Automatic selection of data
-
-The function `auto.rate()` uses rolling regression and kernel density estimate algorithms to detect patterns in respirometry data.
+For more complex respirometry parameters, the function `auto.rate()` uses rolling regression and kernel density estimate algorithms to detect patterns in respirometry data.
 It can calculate maximum or minimum rates, extract rates at non-overlapping intervals, and has an exploratory method to detect the most linear sections of the data.
-
-There are clear advantages to using `auto.rate()` to estimate MMR and SMR.
-Unlike manual methods, where data is often arbitrary selected and inconvenient to replicate, `auto.rate()` adheres to well-defined rules to obtain the rate of interest, so its methods are results are transparent and fully reproducible.
 When called, the function takes the data frame of length $m$, and performs a rolling regression of sample size $n$ iteratively across the length of the data (Fig.
@@
 3b), which is then ranked by size to obtain maximum and minimum values.
+There are clear advantages to using `auto.rate()` when estimating parameters for $MO_{2,max}$ or $MO_{2,min}$.
+Unlike manual methods, where data may be arbitrary selected, `auto.rate()` adheres to well-defined rules to obtain the rate of interest, so its methods are results are transparent and fully reproducible.
+To estimate the most linear sections of the data, further computations are performed after rolling regressions.
+Logically, linear sections of the data should reflect a consistently stable (i.e.
+flat) rate.
+The function takes advantage of this relationship and performs a Gaussian kernel density estimate of all regressions (see Silverman, 1986).
+New datasets around the kernel bandwidth are then resampled around peak values of the generated model, re-analysed, and then ranked by size to determine the most linear sections.
 
-While using rolling regressions to estimate MMR and SMR
+The function `calc.rate.ft()` is used to specifically analyse flowthrough respirometry data, and is similar to `calc.rate()` in functionality.
+However it accepts additional inputs of incurrent and excurrent oxygen concentration, and flow rate, that are not available in `calc.rate()`.
+Results in `calc.rate.ft()` are processed to determine the mean or modal $RO_2$.
 
-### Analysing flowthrough respirometry data
+## Estimating critical oxygen tension, $P_{crit}$
 
-The function `calc.rate.ft()` is used to analyse flowthrough respirometry data, and accepts inputs of incurrent and excurrent oxygen concentration, and flow rate.
-It is crucial that all the input units match as the function does not perform unit conversions (but see `calc.mo2()`), and the output gives $RO_2$.
-Like `calc.rate()`, a segment of the data can be selected before processing and the average, or median $VO_2$ are provided.
+We provide two methods of estimating $P_{crit}$.
+The first is a "broken-stick" regression (BSR) approach, adopted from Yeager and Ultsch (1989) in which two segments of the data are iteratively fitted and the intersection with the smallest sum of the residual sum of squares between the two linear models is the estimated critical point.
+The second method is the segmented, or "broken-line" regression approach, presented by Muggeo (2003) which estimates the critical point by iteratively fitting two intersecting models on the data, and picking the point that minimises the "gap" between the two fitted lines.
+To determine $P_{crit}$, $RO_2$ (or $MO_2$) and ambient oxygen concentration ($PO_2$) must be provided.
+If $RO_2$ has not been calculated, `pcrit()` can automatically perform a rolling regression directly from raw respirometry data to generate the data and arrange it against $PO_2$, and then use one of the two methods to estimate the value (Fig.
+4).
 
@@
 
-Once
-
-### Critical oxygen tension, $P_{c}$
-
-- oxygen tolerance (Hochachka and Somero, 2002)
-- broken stick model (Yeager and Ultsch, 1989)
-- $RO_2$ vs $P_{O_2}$ data can be generated automatically from raw $O_2$ vs time data.
-
-The function `pcrit()` provides a way to consistently report the parameters associated with estimating $P_c$.
-$P_c$ is often estimated using the "broken-stick" regression (BSR) approach (Yeager and Ultsch, 1989), although there are some cases where it is estimated without indication of how thie value was determined numerically (e.g.
-need examples).
-While in many cases the PCrit and oxyregulatory ability can be clear and obvious (e.g./refs), many studies still class species or individuals as oxyregulators or oxyconformers without recognising that there may be intermediate cases.
+For delibrate reasons, units of measurements are not used in any of the functions `calc.bg.rate()`, `calc.rate()`, `auto.rate()`, `calc.rate.ft()` and `pcrit()`. This is because
 

### 2017-10-19 manuscript/manuscript-HTML.Rmd
manuscript update

@@
 
-For delibrate reasons, units of measurements are not used in any of the functions `calc.bg.rate()`, `calc.rate()`, `auto.rate()`, `calc.rate.ft()` and `pcrit()`. This is because
+For delibrate reasons, units of measurements are not used in any of the functions `calc.bg.rate()`, `calc.rate()`, `auto.rate()`, `calc.rate.ft()` and `pcrit()`.
+Instead, two functions, `convert.do()` and `calc.mo2()`, need to be called separately to convert units and to perform volumetric or mass-specific scaling.
+This unique approach has the advantage of allowing us to perform "dimensionless" rate calculations for data that may not be fully supported by the entire `respR` workflow.
+For example, measurements of respiration can be scaled to surface area (diatoms, Iversen & Ploug, 2013, Naumann et al., 2014), volume (copepod eggs, Hammervold et al., 2015), or density (sea urchin larvae, Stumpp et al., 2011).
 
-# Reproducibility and data accesibility
+The function `convert.do()` is used to convert units of oxygen concentration (e.g.
+mg/L or %), while `calc.mo2()` directly accepts object outputs from prior rate-calculating functions and performs unit conversions, volumetric and mass-specific conversions at the same time.
+
+# Reproducibility and data accessibility
 
@@
 All main functions produce an output object of class `list`, which contains all the data and variables needed to re-analyse the results, within R or in other software.
+
+# Concluding remarks and future improvements
 
@@
 **Fig. 3** Selected output plots from the function `calc.rate()` for the analysis of the data frame `urchin2013`.
+
+# Supplementary Material

### 2017-10-19 manuscript/manuscript-Word.Rmd
manuscript update

@@
 ---
-title: "Manuscript_Word"
-author: "Januar Harianto"
-date: "13/10/2017"
+title: "`respR` - An R package for the analysis of respirometry data"
+author: "Januar Harianto, Nicholas Carey, Maria Byrne"
 output: 
@@
 
-```{r setup, include=FALSE}
+```{r, echo = F}
 library(respR)
@@
 
-knitr::opts_chunk$set(echo = TRUE)
+opts_chunk$set(collapse = TRUE, comment = "#>", tidy = TRUE, highlight = TRUE)
 ```
+
+# Introduction
+
+Metabolic rate is a fundamental trait associated with virtually all biological functions, and is key in predicting patterns in ecology and conservation biology, from populations (Seibel and Drazen, 2007, Barneche et al., 2014) to ecosystems (Brown et al., 2004).
+It has been widely used to investigate the effects of external stressors on organisms, and has seen increased interest in studies of climate change where warming is expected to drive increased metabolism in ectotherms (Pörtner, 2002, McElroy et al., 2012, Carey et al., 2016, Delorme & Sewell, 2016).
+By far the most common metric used to assess physiological performance in animals, metabolic rates are determined from whole organisms to the level of cells and tissues (White & Kearney, 2013).
+They are typically quantified using respirometry, which measures oxygen uptake over time as a proxy for metabolic rate, since oxygen is rarely stored and anaerobic metabolism is typically sustained over only brief periods (Lighton, 2008).
+
+There are four broad methodological approaches to respirometry: closed, intermittent flow, flowthrough and open.
+In closed respirometry, $O_2$ decrease is measured within a hermetically sealed chamber of known volume, sometimes set within a closed loop to allow mixing of the environment within the chamber.
+Oxygen recordings may be continuous through use of an oxygen probe, periodic through withdrawing water or gas samples at set intervals, or a two-point measurement consisting of the initial and final concentrations.
+Metabolic rates are estimated from respirometry data by assuming a linear relationship between variables, and estimates of metabolic rate are straightforward in constant volume respirometry using the equation: $$VO_2 = \Delta O_2V$$ where $\Delta O_2$ is the slope of the regression that describes the rate of change in $O_2$ concentration over time, or in the case of a two-point measurement, the difference in $O_2$ concentration divided by time elapsed, and $V$ is the volume of the container (Lighton, 2008).
+<!-- Note, we use here and in the package the term $RO_2$ to denote rate of $O_2$ depletion or uptake within an experiment.
+We use this term used to distinguish it from other common terms, for example $VO_2$ which is can be used to denote both volumetric O2 uptake and whole animal uptake, or $MO_2$ which typically denotes mass-specific uptake.
+-->
+
+In intermittent flow respirometry, $O_2$ concentration is measured as described above, but periodically the chamber is flushed with new water or air, returning it to initial conditions, resealed, and the experiment repeated (Svendsen et al., 2016).
+This technique is essentially closed respirometry, but with the incorporation of repeated measures.
+Depending on the metabolic rate metric being investigated, final respiration rate can be calculated as the mean of the measures (e.g.
+Carey et al., 2016), or the lowest or highest rates recorded in any trial (e.g.
+Stoffels, 2015).
+<!--[nc-will try and find a better ref for that]
+
+Stoffels, R.
+J.
+(2015).
+Physiological trade-offs along a fast-slow lifestyle continuum in fishes: What do they tell us about resistance and resilience to hypoxia?
+Plos One, 10(6), e0130303.
+https://doi.org/10.1371/journal.pone.0130303
+-->
+
+Flowthrough respirometry involves a closed chamber, but with a regulated flow of air or water through it at a precisely determined rate.
+After equilibrium has been achieved, the oxygen concentration differential between the incurrent and excurrent channels, along with the flow rate, allows calculation of the oxygen extracted from the flow volume per unit time: $$VO_2 = (C_iO_2 - C_eO_2)FR$$ where $VO_2$ is the rate of $O_2$ consumption over time, $C_iO_2$ and $C_eO_2$ are the incurrent and excurrent $O_2$ concentrations, and $FR$ is the flow rate of water through the system  (Lighton, 2008).
+
+A final method is open respirometry, in which an open tank or semi-enclosed area is used, but the input or mixing rate of oxygen from the surroundings is known or found to be negligible relative to oxygen consumption of the specimens (Leclercq et al., 1999).
+It is seldom used, but for some applications it is a sufficient and practical methodology (e.g.
+Gamble et al., 2014).
+The common equation used for open respirometry is: $$VO_2 = \Delta O_2V + \phi_d$$ where $\Delta O_2V$ is the slope of the regression that relates $O_2$ concentration to time, $V$ is the volume of the container and $\phi_d$ is the oxygen flux as determined by Fick's Law (Leclercq et al., 1999).
+
+<!--[nc - so the following needs to be spot on.
+Physiologists get very worked up (if not confused themselves) about the differences between these, esp.
+SMR v Routine v Resting v Basal!
+Especially those who work on ecto vs endo therms]-->
+
+Depending on the experiment, different metabolic rate parameters may be of interest to researchers.
+These include maximum metabolic rates (MMR, or $MO_{2,max}$), rates under high activity or exhaustive exercise, and minimal metabolic rates ($MO_{2,min}$), which, depending on the organism, may be termed standard (SMR), basal (BMR), or resting metabolic rates, and typically represent the minimum metabolic cost of maintaining biological functioning (White and Kearney, 2014).
+Routine metabolic rate (RMR), is similar to $MO_{2,min}$, but takes into account that in some organisms energy is expended on small, sponateous movements to maintain posture or position (Rogers et al.
+2016).
+Another metric of great interest is the critical oxygen tension, which represents the lowest level of oxygen at which aerobic metabolism is independent of the ambient partial pressure of oxygen ($P_{crit}$; Yeager & Ultsch, 1989, Hochachka and Somero, 2002).
+
+<!--White, C.
+R., & Kearney, M.
+R.
+(2014).
+Metabolic scaling in animals: methods, empirical results, and theoretical explanations.
+Comprehensive Physiology, 4(1), 231–256.
+https://doi.org/10.1002/cphy.c110049-->
+
+Respirometry experiments are increasingly assimilating large, high-resolution datasets and are run for longer periods (e.g.
+20 h, Norin & Malte, 2012).
+In most cases, processing the data involves an *ad hoc* selection of data points with undefined criteria, and subsequent manual processing of the data subset(s) using a spreadsheet program (e.g.
+Microsoft Excel) or an integrated development environment (IDE, e.g.
+R or Matlab).
+These approaches can be tedious and time consuming especially when spreadsheet programs struggle with the vast datasets that are generated, while IDEs require a degree of expertise to use and have substantial learning curves.
+Dedicated software are also available to perform metabolic rate analyses, but many have costly licensing requirements (e.g.
+AutoResp by Loligo Systems), complicating or preventing their use on multiple machines, and are proprietary and closed-source, hindering scientific reproducability and transparency.
+
+<!-- r.e. below para. LoLinR is not strictly for respirometry - just monotonic, something-changing-over-time data -->
+
+Recently, a number of open-source R software packages have become available which are designed for, or are suitable to analyse respirometry data.
+The `respirometry` package (https://CRAN.R-project.org/package=respirometry) contains a comprehensive collection of tools to explore and evaluate experimental parameters in aquatic respirometry and is useful for the design and diagnosis of experimental setups.
+The `rMR` package (https://CRAN.R-project.org/package=rMR) performs interval-based metabolic rate calculations and has a method to automatically detect $P_{crit}$ using the "broken-stick" regression method (Yeager & Ultsch, 1989).
+`LoLinR` (Olito et al., 2017), while not strictly coded with respirometry in mind, provides a statistically robust method of detecting a "best-fit" regression that applies to the calculation of metabolic rate, and performs very well at identifying truly linear subsets of a data series.
+Importantly, the packages above all assume that data are monotonic with evenly-spaced sampling periods with respect to time and do not have the tools to analyse complicated datasets that require extraction of multiple respirometry parameters.
+
+Here we describe the R package `respR`, a set of functions desgined to provide an efficient, structured, and reproduceable workflow for the analysis of respirometry-related data (Fig.
+1).
+The package contains utilities to: (1) analyse closed, intermittent, flow-through and open respirometry data, (2) determine volume, mass, and surface area specific oxygen uptake rates, (3) automatically detect maximal and minimal rates, $P_{crit}$, or the most linear results in large datasets with the help of traditional rolling regression and kernel density estimation techniques.
+Other functions useful in processing respirometry data are avaliable, and are described in more detail in our online html vignette (https://januarharianto.github.io/respR/).
+The package has a strong focus on aquatic respirometry, however similar principals apply to aerial respirometry, so the analysis workflow should also be suitable for these data.
+We demonstrate the utility of the package by analysing three datasets as case studies and proof of concept.
+These data were collected using closed, intermittent and flow-through respirometry, and we show how `respR` might be used to determine $MO_{2,max}$, $MO_{2,min}$, and $P_{crit}$.
+
+# Package overview
+
+`respR` can be automatically installed from the GitHub repository using the `devtools` package:
+
+```{r eval=F}
+devtools::install_github("januarharianto/respR")
+library(respR)
+```
+
+To explore `respR` and its main functions we have provided example data that can be called directly after the package is loaded.
+Detailed information about all example data, including their source and methods, can be obtained with the `?` command in the R console (e.g.
+`?urchin2013`).
+All analyses described in this paper use the example data provided and are fully reproducible (see supplementary information).
+
+## Data import and exploration
+
+Data should be formatted correctly before use in `respR`.
+The function `check.input()` performs error checks on a data frame of any size, and can extract a two-column data frame for data exploration and subsequent analyses.
+Data must be numeric, because `respR` has limited support for date/time data (e.g.
+`POSIXct` and `POSIXlt` classes) and performing calculations on different time formats have many caveats that are beyond the scope of the package.
+Time data, being a continuous variable, should be sequential, without duplicates.
+Ideally, time data should also be evenly-spaced, that is, the time period between samples are equal, as this will affect data-extraction methods later on if data is analysed by sample length instead of time.
+Error check results are summarised in the R console (see supplementary material) and plot of the data is shown for visual inspection (e.g.
+Fig.
+1).
+
+It should be noted that invoking `check.input` is optional -- the main functions in our package will readily accept any data frame as long as data are numeric.
+Running `check.input` is an exploratory step that flags potential issues about the data before it is further analysed.
+
+## Estimating the rate of change in $O_2$ concentration ($RO_2$)
+
+### Calculating background $RO_2$
+
+Background $RO_2$ is often measured to account for the influence of bacterial respiration in the chamber, and applied as a correction to the main experiment (Rodgers et al., 2016).
+Since background rates typically account for a small percentage of experimental rates, multiple “blank” experiments are routinely conducted and the rates are averaged across several datasets to obtain a more accurate estimate of the correction (e.g.
+Carey and Sigwart, 2014; Daoud et al., 2007).
+The function `calc.bg.rate()` uses simple linear regression to simultaneously process multiple background rate measurements and produces an output object that can be used in later functions that accept background rate arguments.
+Because the oxygen consumption units are non-volume and non-mass specific at the point of correction, the functions in `respR` can adjust $RO_2$ directly without bias: $$RO_{2(correction)} = RO_2 - RO_{2(bg)}$$
+
+### Extracting data for $RO_2$
+
+Typically, the analysis respirometry data requires the selection one or more linear sections of the raw data for more accurate calculations of $RO_2$ (e.g.
+Gordon et al., 1989, Dorey et al., 2014, Chu & Gal, 2017).
+The function `calc.rate()` can manually extract and process data segments from respiromety data.
+Data segments can be selected by (1) time period, (2) row numbers, (3) $O_2$ decrease or (4) proportion, which should accomodate most, if not all, data selection requirements and allow for consistent reporting of methods and results (e.g.
+estimating $RO_2$ over a regular 10-minute time window across multiple specimens).
+Rates are determined using linear regression, however a two-point analysis is also included alongside all regression results.
+
+For more complex respirometry parameters, the function `auto.rate()` uses rolling regression and kernel density estimate algorithms to detect patterns in respirometry data.
+It can calculate maximum or minimum rates, extract rates at non-overlapping intervals, and has an exploratory method to detect the most linear sections of the data.
+When called, the function takes the data frame of length $m$, and performs a rolling regression of sample size $n$ iteratively across the length of the data (Fig.
+2b).
+Thus, a total of $(m−n)+1$ number of overlapping regressions are fitted, and we obtain a new dataset of regressions as a function of time (rolling regression, Fig.
+3b), which is then ranked by size to obtain maximum and minimum values.
+There are clear advantages to using `auto.rate()` when estimating parameters for $MO_{2,max}$ or $MO_{2,min}$.
+Unlike manual methods, where data may be arbitrary selected, `auto.rate()` adheres to well-defined rules to obtain the rate of interest, so its methods are results are transparent and fully reproducible.
+To estimate the most linear sections of the data, further computations are performed after rolling regressions.
+Logically, linear sections of the data should reflect a consistently stable (i.e.
+flat) rate.
+The function takes advantage of this relationship and performs a Gaussian kernel density estimate of all regressions (see Silverman, 1986).
+New datasets around the kernel bandwidth are then resampled around peak values of the generated model, re-analysed, and then ranked by size to determine the most linear sections.
+
+The function `calc.rate.ft()` is used to specifically analyse flowthrough respirometry data, and is similar to `calc.rate()` in functionality.
+However it accepts additional inputs of incurrent and excurrent oxygen concentration, and flow rate, that are not available in `calc.rate()`.
+Results in `calc.rate.ft()` are processed to determine the mean or modal $RO_2$.
+
+## Estimating critical oxygen tension, $P_{crit}$
+
+We provide two methods of estimating $P_{crit}$.
+The first is a "broken-stick" regression (BSR) approach, adopted from Yeager and Ultsch (1989) in which two segments of the data are iteratively fitted and the intersection with the smallest sum of the residual sum of squares between the two linear models is the estimated critical point.
+The second method is the segmented, or "broken-line" regression approach, presented by Muggeo (2003) which estimates the critical point by iteratively fitting two intersecting models on the data, and picking the point that minimises the "gap" between the two fitted lines.
+To determine $P_{crit}$, $RO_2$ (or $MO_2$) and ambient oxygen concentration ($PO_2$) must be provided.
+If $RO_2$ has not been calculated, `pcrit()` can automatically perform a rolling regression directly from raw respirometry data to generate the data and arrange it against $PO_2$, and then use one of the two methods to estimate the value (Fig.
+4).
+
+## Converting units of $RO_2$ to $VO_2$ and $MO_2$
+
+For delibrate reasons, units of measurements are not used in any of the functions `calc.bg.rate()`, `calc.rate()`, `auto.rate()`, `calc.rate.ft()` and `pcrit()`.
+Instead, two functions, `convert.do()` and `calc.mo2()`, need to be called separately to convert units and to perform volumetric or mass-specific scaling.
+This unique approach has the advantage of allowing us to perform "dimensionless" rate calculations for data that may not be fully supported by the entire `respR` workflow.
+For example, measurements of respiration can be scaled to surface area (diatoms, Iversen & Ploug, 2013, Naumann et al., 2014), volume (copepod eggs, Hammervold et al., 2015), or density (sea urchin larvae, Stumpp et al., 2011).
+
+The function `convert.do()` is used to convert units of oxygen concentration (e.g.
+mg/L or %), while `calc.mo2()` directly accepts object outputs from prior rate-calculating functions and performs unit conversions, volumetric and mass-specific conversions at the same time.
+
+# Reproducibility and data accessibility
+
+We have ensured that any computational analysis performed by `respR` are fully reproducible.
+All main functions produce an output object of class `list`, which contains all the data and variables needed to re-analyse the results, within R or in other software.
+
+# Concluding remarks and future improvements
+
+# Tables & Figures
+
+**Table 1.** Main functions available in `respR`.
+```{r, echo=F, message=F}
+# use this later to show table
+library(readr)
+listfunctions <- read_csv('tables/list-of-functions.csv')
+```
+
+```{r, echo=F, results='asis'}
+pander(listfunctions)
+```
+
+```{r, echo = F, out.width = "900px"}
+knitr::include_graphics("figs/fig1.png")
+```
+
+**Fig.
+1.** Diagram showing a typical workflow in the analysis of respirometry data using `respR`.
+Data is first checked for errors before the main functions are used to extract and analyse segments of the data.
+Summarised results and diagnostic plots provide immediate visual feedback on the outcome of the analyses.
+Once the rate estimates are obtained, they can be converted into volume and/or mass-specific rates.
+
+```{r, echo = F, out.width = "650px"}
+knitr::include_graphics("figs/fig2.png")
+```
+
+**Fig.
+2** Simplified illustration showing the sampling windows used by `auto.rate()` during **(a)** interval regression analysis, where sample windows do not overlap and **(b)** rolling regression analysis, where sample windows overlap and move forward by one sample unit at a time.
+
+```{r, echo = F, , out.width = "800px"}
+knitr::include_graphics("figs/fig3.png")
+```
+
+**Fig. 3** Selected output plots from the function `calc.rate()` for the analysis of the data frame `urchin2013`.
+
+# Supplementary Material

### 2017-11-01 manuscript/manuscript-HTML.Rmd
manuscript update

@@
 Depending on the experiment, different metabolic rate parameters may be of interest to researchers.
-These include maximum metabolic rates (MMR, or $MO_{2,max}$), rates under high activity or exhaustive exercise, and minimal metabolic rates ($MO_{2,min}$), which, depending on the organism, may be termed standard (SMR), basal (BMR), or resting metabolic rates, and typically represent the minimum metabolic cost of maintaining biological functioning (White and Kearney, 2014).
+These include maximum metabolic rates (MMR, or $MO_{2,max}$), which define rates under high activity or exhaustive exercise, and minimal metabolic rates ($MO_{2,min}$), which may be termed standard (SMR), basal (BMR), or resting metabolic rates, and typically represent the minimum metabolic cost of maintaining biological functioning (White and Kearney, 2014).
 Routine metabolic rate (RMR), is similar to $MO_{2,min}$, but takes into account that in some organisms energy is expended on small, sponateous movements to maintain posture or position (Rogers et al.
@@
 The `rMR` package (https://CRAN.R-project.org/package=rMR) performs interval-based metabolic rate calculations and has a method to automatically detect $P_{crit}$ using the "broken-stick" regression method (Yeager & Ultsch, 1989).
-`LoLinR` (Olito et al., 2017), while not strictly coded with respirometry in mind, provides a statistically robust method of detecting a "best-fit" regression that applies to the calculation of metabolic rate, and performs very well at identifying truly linear subsets of a data series.
-Importantly, the packages above all assume that data are monotonic with evenly-spaced sampling periods with respect to time and do not have the tools to analyse complicated datasets that require extraction of multiple respirometry parameters.
+`LoLinR` (Olito et al., 2017), while not strictly coded for respirometry, provides a statistically robust method of detecting a "best-fit" regression that applies to the calculation of metabolic rate, and performs very well at identifying truly linear subsets of a data series.
+Importantly, all of these packages assume that data are monotonic with evenly-spaced sampling periods with respect to time, and do not have the tools to analyse complicated datasets to automatically determine useful metric such as $MO_{2,max}$ and $MO_{2,min}$.
 
-Here we describe the R package `respR`, a set of functions desgined to provide an efficient, structured, and reproduceable workflow for the analysis of respirometry-related data (Fig.
-1).
-The package contains utilities to: (1) analyse closed, intermittent, flow-through and open respirometry data, (2) determine volume, mass, and surface area specific oxygen uptake rates, (3) automatically detect maximal and minimal rates, $P_{crit}$, or the most linear results in large datasets with the help of traditional rolling regression and kernel density estimation techniques.
-Other functions useful in processing respirometry data are avaliable, and are described in more detail in our online html vignette (https://januarharianto.github.io/respR/).
+<!-- FIGURE 1 -->
+Here we describe the R package `respR`, a set of functions desgined to provide an efficient, structured, and reproduceable workflow for the analysis of respirometry-related data (**Table 1**).
+The package contains utilities to: (1) analyse closed, intermittent, flow-through and open respirometry data, (2) determine volume, mass, and surface area specific oxygen uptake rates, (3) automatically detect maximal and minimal rates, $P_{crit}$, or (4) detect the most linear sections of large datasets, with the help of traditional rolling regression and kernel density estimation techniques.
+Other functions useful in processing respirometry data are avaliable and are described in more detail in our online html vignette (https://januarharianto.github.io/respR/).
 The package has a strong focus on aquatic respirometry, however similar principals apply to aerial respirometry, so the analysis workflow should also be suitable for these data.
-We demonstrate the utility of the package by analysing three datasets as case studies and proof of concept.
-These data were collected using closed, intermittent and flow-through respirometry, and we show how `respR` might be used to determine $MO_{2,max}$, $MO_{2,min}$, and $P_{crit}$.
+We demonstrate the utility of the package by analysing real data that were collected using closed, intermittent and flow-through respirometry as proof of concept, and show how `respR` might be used to determine $MO_{2,max}$, $MO_{2,min}$, and $P_{crit}$.
 
@@
 Data should be formatted correctly before use in `respR`.
-The function `check.input()` performs error checks on a data frame of any size, and can extract a two-column data frame for data exploration and subsequent analyses.
-Data must be numeric, because `respR` has limited support for date/time data (e.g.
-`POSIXct` and `POSIXlt` classes) and performing calculations on different time formats have many caveats that are beyond the scope of the package.
+The function `inspect.data()` performs error checks on a data frame of any size, and can extract a two-column data frame for data exploration and subsequent analyses.
+Ideally all data are numeric, because `respR` has limited support for date/time data (e.g.
+`POSIXct` and `POSIXlt` classes), and performing calculations on different time formats have many caveats that are beyond the scope of the package.
 Time data, being a continuous variable, should be sequential, without duplicates.
-Ideally, time data should also be evenly-spaced, that is, the time period between samples are equal, as this will affect data-extraction methods later on if data is analysed by sample length instead of time.
-Error check results are summarised in the R console (see supplementary material) and plot of the data is shown for visual inspection (e.g.
-Fig.
-1).
+Ideally, sampling frequency should be monotonic, but in some cases this may be difficult as machine drift and/or data dropouts may occur.
+However, irregular time series data can still be analysed by `respR` as it supports time-based data extraction methods.
 
-It should be noted that invoking `check.input` is optional -- the main functions in our package will readily accept any data frame as long as data are numeric.
-Running `check.input` is an exploratory step that flags potential issues about the data before it is further analysed.
+It should be noted that using `inspect.data()` is optional -- the main functions in our package will readily work with data frames that were not checked.
+Running `inspect.data()` is an exploratory step that flags potential issues about the data before they are analysed.
+Error check results are summarised in the R console, and all error locations (i.e.
+row numbers) are saved in the output for inspection.
+A plot of the data is also shown for visual inspection (**Fig.
+1**).
 
@@
 Carey and Sigwart, 2014; Daoud et al., 2007).
-The function `calc.bg.rate()` uses simple linear regression to simultaneously process multiple background rate measurements and produces an output object that can be used in later functions that accept background rate arguments.
-Because the oxygen consumption units are non-volume and non-mass specific at the point of correction, the functions in `respR` can adjust $RO_2$ directly without bias: $$RO_{2(correction)} = RO_2 - RO_{2(bg)}$$
+The function `calc.rate.bg()` uses simple linear regression to simultaneously process multiple background rate measurements and produces an output object that can be used directly in later functions that accept background rate arguments.
+Data segments can be truncated before analysis, by either time period or row numbers.
+As the oxygen consumption units are not mass-specific at the point of correction, the functions in `respR` can adjust $RO_2$ directly without bias using the equation: $$RO_{2(correction)} = RO_2 - RO_{2(bg)}$$
 
-### Extracting data for $RO_2$
+### Extracting data and calculating $RO_2$
 
@@
 Data segments can be selected by (1) time period, (2) row numbers, (3) $O_2$ decrease or (4) proportion, which should accomodate most, if not all, data selection requirements and allow for consistent reporting of methods and results (e.g.
-estimating $RO_2$ over a regular 10-minute time window across multiple specimens).
+estimating $RO_2$ over a fixed 10-minute time window across multiple specimens).
 Rates are determined using linear regression, however a two-point analysis is also included alongside all regression results.
@@
 When called, the function takes the data frame of length $m$, and performs a rolling regression of sample size $n$ iteratively across the length of the data (Fig.
-2b).
+2a).
 Thus, a total of $(m−n)+1$ number of overlapping regressions are fitted, and we obtain a new dataset of regressions as a function of time (rolling regression, Fig.
 3b), which is then ranked by size to obtain maximum and minimum values.
-There are clear advantages to using `auto.rate()` when estimating parameters for $MO_{2,max}$ or $MO_{2,min}$.
-Unlike manual methods, where data may be arbitrary selected, `auto.rate()` adheres to well-defined rules to obtain the rate of interest, so its methods are results are transparent and fully reproducible.
+If an interval-based regression is required, the function selects non-overlapping sections of the data for regressions (Fig.
+2b).
+
 To estimate the most linear sections of the data, further computations are performed after rolling regressions.
@@
 flat) rate.
-The function takes advantage of this relationship and performs a Gaussian kernel density estimate of all regressions (see Silverman, 1986).
-New datasets around the kernel bandwidth are then resampled around peak values of the generated model, re-analysed, and then ranked by size to determine the most linear sections.
+If the rate changes at some point during the sampling period, thent the rolling estimate should detect the variability.
+The function takes advantage of this relationship and performs a Gaussian kernel density estimate (KDE) of all regressions (see Silverman, 1986).
+Peaks in the density thus reflect stable rates, and new segments of the data within the kernel bandwidth of the peak values are resampled and re-analysed to determine the most linear sections.
+
+There are clear advantages to using `auto.rate()` when estimating parameters for $MO_{2,max}$ or $MO_{2,min}$.
+Unlike manual methods, where data may be arbitrary selected, `auto.rate()` adheres to well-defined rules to obtain the rate of interest, so its methods are results are transparent and fully reproducible.
 
 The function `calc.rate.ft()` is used to specifically analyse flowthrough respirometry data, and is similar to `calc.rate()` in functionality.
-However it accepts additional inputs of incurrent and excurrent oxygen concentration, and flow rate, that are not available in `calc.rate()`.
-Results in `calc.rate.ft()` are processed to determine the mean or modal $RO_2$.
+However it accepts additional inputs of incurrent and excurrent oxygen concentration, and flow rate, that are not available in `calc.rate()`.`calc.rate.ft()` can process data to determine the mean or modal $RO_2$ over a period.
 
@@
 The second method is the segmented, or "broken-line" regression approach, presented by Muggeo (2003) which estimates the critical point by iteratively fitting two intersecting models on the data, and picking the point that minimises the "gap" between the two fitted lines.
+
 To determine $P_{crit}$, $RO_2$ (or $MO_2$) and ambient oxygen concentration ($PO_2$) must be provided.
-If $RO_2$ has not been calculated, `pcrit()` can automatically perform a rolling regression directly from raw respirometry data to generate the data and arrange it against $PO_2$, and then use one of the two methods to estimate the value (Fig.
-4).
+If $RO_2$ is not available, but the respirometry data exists to calculate it, `pcrit()` can automatically perform a rolling regression directly on the data to generate $RO_2$ and arrange it against the rolling mean of $PO_2$.
+This method is unique to `respR` -- in the past, a user may generate the same by cutting the data into equal sections and performing a linear regression on every section.
+A rolling regression
 
@@
 
-For delibrate reasons, units of measurements are not used in any of the functions `calc.bg.rate()`, `calc.rate()`, `auto.rate()`, `calc.rate.ft()` and `pcrit()`.
-Instead, two functions, `convert.do()` and `calc.mo2()`, need to be called separately to convert units and to perform volumetric or mass-specific scaling.
+For delibrate reasons, units of measurements are not used in any of the functions `calc.rate.bg()`, `calc.rate()`, `calc.rate.ft()`, `auto.rate()` and `pcrit()`.
+Instead, two functions, `convert.rate()` and `scale.rate()`, need to be called separately to convert units and to perform volumetric or mass-specific scaling.
 This unique approach has the advantage of allowing us to perform "dimensionless" rate calculations for data that may not be fully supported by the entire `respR` workflow.
 For example, measurements of respiration can be scaled to surface area (diatoms, Iversen & Ploug, 2013, Naumann et al., 2014), volume (copepod eggs, Hammervold et al., 2015), or density (sea urchin larvae, Stumpp et al., 2011).
+Users may thus first perform rate calculations using `respR`, and then scale the results manually after if there is no support for their specific scaling method.
 
-The function `convert.do()` is used to convert units of oxygen concentration (e.g.
-mg/L or %), while `calc.mo2()` directly accepts object outputs from prior rate-calculating functions and performs unit conversions, volumetric and mass-specific conversions at the same time.
+The function `convert.rate()` is used to convert units of oxygen concentration (e.g.
+mg/L or %) only and has support for over 132 unit combinations.
+As unit conventions differ between research disciplines and we do not expect
+
+`scale.rate()` directly accepts object outputs from prior rate-calculating functions and performs volumetric and mass-specific conversions. We have implemented a "semi-fuzzy" approach to
 
@@
 # Concluding remarks and future improvements
+While performing the required linear regressions to estimate the parameter of interest is a relatively simple task, a considerable amount of time is spent in data formatting and manipulation.
 
@@
 ```{r, echo = F, out.width = "900px"}
-knitr::include_graphics("figs/fig1.png")
+knitr::include_graphics("figs/Figure 1.png")
 ```
@@
 ```{r, echo = F, out.width = "650px"}
-knitr::include_graphics("figs/fig2.png")
+knitr::include_graphics("figs/Figure 2.png")
 ```
@@
 **Fig.
-2** Simplified illustration showing the sampling windows used by `auto.rate()` during **(a)** interval regression analysis, where sample windows do not overlap and **(b)** rolling regression analysis, where sample windows overlap and move forward by one sample unit at a time.
+2** Illustration showing the sampling method used by `auto.rate()` during **(a)** interval regression analysis, where sample windows do not overlap and **(b)** rolling regression analysis, where sample windows overlap and move forward by one sample unit at a time.
 
 ```{r, echo = F, , out.width = "800px"}
-knitr::include_graphics("figs/fig3.png")
+knitr::include_graphics("figs/Figure 3.png")
 ```
 
-**Fig. 3** Selected output plots from the function `calc.rate()` for the analysis of the data frame `urchin2013`.
+**Fig.
+3** Typical graphical outputs of the function `auto.rate()`.
+(a) Depending on the `method` input, the function is able to show: (clockwise) the maximum rate detected, the minimum rate detected, calculation of rate by fixed-width intervals, and the most linear region as determined by kernel density estimates.
+(b) In addition, the outputs will also print diagnostic plots of: (clockwise) the rolling regression of rate as a function of time with dotted lines indicating the region used to estimate the parameter, a kernel density plot that shows only when the `linear` method is called, a residual plot of the regression, and a normal Q-Q plot of the residuals of the regression.
+All data is obtained from the example dataset `sardine` which is included in the package.
 

### 2017-11-03 manuscript/manuscript-HTML.Rmd
manuscript update and new figures

@@
 library(readr)
-
+Sys.setenv(TZ="Australia/Sydney")
 # set pander table-layout options
@@
 
+# Abstract
+
+1.
+Estimating metabolic rate from respirometry is a conceptually simple, but fundamentally difficult technique where researchers often arbitrarily extract one or more segments of a $O_2$ timeseries to estimate the parameter of interest.
+As respirometry data becomes increasingly large and analytical approaches become more complex, manipulating it remains a challenge for current tools, with important limitations that prevent objectively reproducible research.
+
+2.
+Here we describe the `respR` R package, a collection of functions that implement a workflow-based approach to automate the analysis and visualisation of respirometry data.
+The package can be used for closed, intermittent flow, flowthrough and open respirometry and uses well-defined sets of rules to realiably and rapidly facilitate reproducible results.
+
+3.
+We demonstrate how `respR` uses novel and parallelised computing methods such as rolling regressions and kernel density estimates to sample the data iteratively to realiably detect maximum, minimum & most linear sections of the data, and critical oxygen tension, $P_{crit}$.
+
+4.
+Although designed specifically with aquatic respirometry in mind, the object-oriented approach of the package the "dimensionless" nature of its analytical functions mean that parts of the package could easily be used to estimate any other linear relationships in other research disciplines.
+
 # Introduction
 
-Metabolic rate is a fundamental trait associated with virtually all biological functions, and is key in predicting patterns in ecology and conservation biology, from populations (Seibel and Drazen, 2007, Barneche et al., 2014) to ecosystems (Brown et al., 2004).
-It has been widely used to investigate the effects of external stressors on organisms, and has seen increased interest in studies of climate change where warming is expected to drive increased metabolism in ectotherms (Pörtner, 2002, McElroy et al., 2012, Carey et al., 2016, Delorme & Sewell, 2016).
-By far the most common metric used to assess physiological performance in animals, metabolic rates are determined from whole organisms to the level of cells and tissues (White & Kearney, 2013).
+Metabolic rate is a fundamental trait associated with virtually all biological functions and is key in predicting patterns in ecology and conservation biology, from populations (Seibel and Drazen, 2007, Barneche et al., 2014) to ecosystems (Brown et al., 2004).
+It has been widely investigated in studies of the effects of external stressors on organisms, and is increasingly used to study resilience in climate change-related studies where warming is expected to drive increased metabolism in ectotherms (e.g.
+Pörtner, 2002, McElroy et al., 2012, Carey et al., 2016, Delorme & Sewell, 2016).
+By far the most common metric used to assess physiological performance in animals, metabolic rate has been determined from whole organisms, to the level of cells and tissues (White & Kearney, 2013).
 They are typically quantified using respirometry, which measures oxygen uptake over time as a proxy for metabolic rate, since oxygen is rarely stored and anaerobic metabolism is typically sustained over only brief periods (Lighton, 2008).
@@
 
-Recently, a number of open-source R software packages have become available which are designed for, or are suitable to analyse respirometry data.
+Further, a number of open-source R software packages have recently become available which are designed for, or are suitable to analyse respirometry data.
 The `respirometry` package (https://CRAN.R-project.org/package=respirometry) contains a comprehensive collection of tools to explore and evaluate experimental parameters in aquatic respirometry and is useful for the design and diagnosis of experimental setups.
@@
 <!-- FIGURE 1 -->
-Here we describe the R package `respR`, a set of functions desgined to provide an efficient, structured, and reproduceable workflow for the analysis of respirometry-related data (**Table 1**).
-The package contains utilities to: (1) analyse closed, intermittent, flow-through and open respirometry data, (2) determine volume, mass, and surface area specific oxygen uptake rates, (3) automatically detect maximal and minimal rates, $P_{crit}$, or (4) detect the most linear sections of large datasets, with the help of traditional rolling regression and kernel density estimation techniques.
+Here we describe `respR`, an open-source R package desgined to provide an efficient and reproduceable workflow for the analysis of respirometry data.
+The package contains utilities to: (1) analyse closed, intermittent, flow-through and open respirometry data, (2) determine volume, mass, and surface area specific oxygen uptake rates, (3) automatically detect maximal and minimal rates, $P_{crit}$, and (4) detect the most linear sections of data using traditional rolling regression and kernel density estimation techniques.
 Other functions useful in processing respirometry data are avaliable and are described in more detail in our online html vignette (https://januarharianto.github.io/respR/).
-The package has a strong focus on aquatic respirometry, however similar principals apply to aerial respirometry, so the analysis workflow should also be suitable for these data.
-We demonstrate the utility of the package by analysing real data that were collected using closed, intermittent and flow-through respirometry as proof of concept, and show how `respR` might be used to determine $MO_{2,max}$, $MO_{2,min}$, and $P_{crit}$.
+We demonstrate the utility of the package by analysing past data that were collected using closed, intermittent and flow-through respirometry as proof of concept, and show how `respR` might be used to determine $MO_{2,max}$, $MO_{2,min}$, and $P_{crit}$.
 
@@
 
-`respR` can be automatically installed from the GitHub repository using the `devtools` package:
+The `respR` package streamlines the analysis of respirometry data by partitioning the data processing workflow into several independent actions: (1) initially, data is imported, checked for errors, automatically plotted for visual inspection and then exported as an object; (2) the object is then analysed to estimate the desired rate parameter; (3) diagnostic plots are automatically generated and the data can be further explored using the generic S3 `print()`, `plot()`, and `summary()` commands to verify the results; (4) adjustments to the data, either from background respiration and/or oxygen flux in open respirometry, are applied; (5) volume or mass-specific rate conversions, specific to the respiratory chamber and specimen(s), are calculated where necessary (**Fig.
+1**).
+By design, each step of the workflow requires minimal manual manipulation of the data, and the user can pass the output object of a function to the next.
+The object-oriented approach makes `respR` easy to use, even for novice users of R, and allows us (or anyone interested in the open-source project) to easily extend the functionality of the package, either by adding more intermediate actions or substituting current actions with other code without breaking the rest of the package.
+
+Installation of `respR` can be performed directly in the R console using `devtools`:
 
 ```{r eval=F}
+# install.packages("devtools") 
 devtools::install_github("januarharianto/respR")
@@
 
-To explore `respR` and its main functions we have provided example data that can be called directly after the package is loaded.
-Detailed information about all example data, including their source and methods, can be obtained with the `?` command in the R console (e.g.
-`?urchin2013`).
-All analyses described in this paper use the example data provided and are fully reproducible (see supplementary information).
+To illustrate the main features of the package, 4 example datasets, `sardine.rd`, `urchin.rd`, `intermittent.rd` and `squid.rd` are distributed with `respR`.
+Detailed information about the example data, including their source and methods, can be obtained with the `?` command in the R console (e.g.
+`?sardine.rd`).
 
-## Data import and exploration
+# Data import and exploration
 
 Data should be formatted correctly before use in `respR`.
-The function `inspect.data()` performs error checks on a data frame of any size, and can extract a two-column data frame for data exploration and subsequent analyses.
+The function `inspect.data()` extracts a two-column data frame from a multi-column dataset and performs error checks on a data frame of any size, while plotting the data for quick data visualisation.
 Ideally all data are numeric, because `respR` has limited support for date/time data (e.g.
@@
 Time data, being a continuous variable, should be sequential, without duplicates.
-Ideally, sampling frequency should be monotonic, but in some cases this may be difficult as machine drift and/or data dropouts may occur.
-However, irregular time series data can still be analysed by `respR` as it supports time-based data extraction methods.
+Ideally, sampling frequency should also be monotonic, but in some cases this may be difficult as machine drift and/or data dropouts can occur during regular sampling.
+In any case, `respR` uses time-based data extraction methods by default in most of its analytical functions (except `pcrit()`), and correctly handle irregular timeseries data by matching the input times to the closes matching values within the data, rounded down.
 
-It should be noted that using `inspect.data()` is optional -- the main functions in our package will readily work with data frames that were not checked.
+<!-- ...Identifying irregularities in the sampling time is therefore important as it as variations in data intervals may cause systemic bias in the analysis, especially if data loss or machine drift during sampling is persistent through time and time-based comparisons are made between experiments.
+-->
+
+It should be noted that using `inspect.data()` is optional -- the main functions in our package will readily work with data frames that were not processed by the function.
 Running `inspect.data()` is an exploratory step that flags potential issues about the data before they are analysed.
-Error check results are summarised in the R console, and all error locations (i.e.
-row numbers) are saved in the output for inspection.
-A plot of the data is also shown for visual inspection (**Fig.
-1**).
+Error check results are summarised in the R console and all tests are saved in the output for us to view the locations (i.e.
+row numbers) of the errors:
 
-## Estimating the rate of change in $O_2$ concentration ($RO_2$)
+```{r}
+inspect.data(sardine, plot = FALSE)
+```
+
+## Data analysis
 
@@
 Carey and Sigwart, 2014; Daoud et al., 2007).
-The function `calc.rate.bg()` uses simple linear regression to simultaneously process multiple background rate measurements and produces an output object that can be used directly in later functions that accept background rate arguments.
+The function `calc.rate.bg()` uses simple linear regression to simultaneously process multiple background rate measurements.
 Data segments can be truncated before analysis, by either time period or row numbers.
-As the oxygen consumption units are not mass-specific at the point of correction, the functions in `respR` can adjust $RO_2$ directly without bias using the equation: $$RO_{2(correction)} = RO_2 - RO_{2(bg)}$$
+The results are stored as an object for later use.
 
@@
 The function `calc.rate()` can manually extract and process data segments from respiromety data.
-Data segments can be selected by (1) time period, (2) row numbers, (3) $O_2$ decrease or (4) proportion, which should accomodate most, if not all, data selection requirements and allow for consistent reporting of methods and results (e.g.
-estimating $RO_2$ over a fixed 10-minute time window across multiple specimens).
+Data segments can be selected by (1) time period, (2) row numbers, (3) $O_2$ decrease or (4) proportion, which should accomodate most, if not all, data selection requirements and allow for consistent reporting of methods and results.
 Rates are determined using linear regression, however a two-point analysis is also included alongside all regression results.
 
-For more complex respirometry parameters, the function `auto.rate()` uses rolling regression and kernel density estimate algorithms to detect patterns in respirometry data.
+For more complex respirometry parameters, the function `auto.rate()` uses novel rolling regression and kernel density estimate algorithms to detect patterns in respirometry data.
 It can calculate maximum or minimum rates, extract rates at non-overlapping intervals, and has an exploratory method to detect the most linear sections of the data.
-When called, the function takes the data frame of length $m$, and performs a rolling regression of sample size $n$ iteratively across the length of the data (Fig.
-2a).
-Thus, a total of $(m−n)+1$ number of overlapping regressions are fitted, and we obtain a new dataset of regressions as a function of time (rolling regression, Fig.
-3b), which is then ranked by size to obtain maximum and minimum values.
-If an interval-based regression is required, the function selects non-overlapping sections of the data for regressions (Fig.
-2b).
+When called, the function takes the data frame of length $m$, and performs regressions of sample size $n$ iteratively across the length of the data (i.e.
+rolling regressions), one sample at a time (**Fig.
+2a**).
+Thus, a total of $(m−n)+1$ number of overlapping regressions are fitted, which are then ranked by size to obtain maximum and minimum values.
+If an interval-based regression is required, we selecth the `"interval"` method when running the function and it automatically selects non-overlapping sections of the data for regressions (**Fig.
+2b**).
 
@@
 Peaks in the density thus reflect stable rates, and new segments of the data within the kernel bandwidth of the peak values are resampled and re-analysed to determine the most linear sections.
-
-There are clear advantages to using `auto.rate()` when estimating parameters for $MO_{2,max}$ or $MO_{2,min}$.
-Unlike manual methods, where data may be arbitrary selected, `auto.rate()` adheres to well-defined rules to obtain the rate of interest, so its methods are results are transparent and fully reproducible.
+Summary diagnostics and visual plots provide supporting information on the technique and the validity of the results (**Fig.
+3**).
 
@@
 
-## Estimating critical oxygen tension, $P_{crit}$
+### Estimating critical oxygen tension, $P_{crit}$
 
@@
 The first is a "broken-stick" regression (BSR) approach, adopted from Yeager and Ultsch (1989) in which two segments of the data are iteratively fitted and the intersection with the smallest sum of the residual sum of squares between the two linear models is the estimated critical point.
-The second method is the segmented, or "broken-line" regression approach, presented by Muggeo (2003) which estimates the critical point by iteratively fitting two intersecting models on the data, and picking the point that minimises the "gap" between the two fitted lines.
+The second method is a wrapper for the segmented, or "broken-line" regression approach, presented by Muggeo (2003) and avaliable as an R package (Muggeo 2008), which estimates the critical point by iteratively fitting two intersecting models on the data and picks the point that minimises the "gap" between the two fitted lines.
 
 To determine $P_{crit}$, $RO_2$ (or $MO_2$) and ambient oxygen concentration ($PO_2$) must be provided.
-If $RO_2$ is not available, but the respirometry data exists to calculate it, `pcrit()` can automatically perform a rolling regression directly on the data to generate $RO_2$ and arrange it against the rolling mean of $PO_2$.
-This method is unique to `respR` -- in the past, a user may generate the same by cutting the data into equal sections and performing a linear regression on every section.
-A rolling regression
+If $RO_2$ is not available, but the respirometry data exists to calculate it, `pcrit()` can automatically perform a rolling regression directly on the data to generate $RO_2$ and arrange it against the rolling mean of $PO_2$ (**Fig.
+4**), before calculating $P_{crit}$.
 
-## Converting units of $RO_2$ to $VO_2$ and $MO_2$
+## Correcting, converting and scaling data
 
-For delibrate reasons, units of measurements are not used in any of the functions `calc.rate.bg()`, `calc.rate()`, `calc.rate.ft()`, `auto.rate()` and `pcrit()`.
+### Corrections for oxygen flux and background respiration
+
+The function `adjust.rate()` can be used to correct $RO_2$ estimates to oxygen flux and/or background respirometry data.
+As the oxygen consumption units are not mass-specific at the point of correction, the function adjusts $RO_2$ directly without bias using the equation: $$RO_{2(correction)} = RO_2 - (\phi_d + RO_{2(bg)})$$ where $\phi_d$ is the oxygen flux and $RO_{2(bg)}$ is background respiration.
+
+### Rate conversions
+
+For delibrate reasons, units of measurements are not used when estimating rates in R.
 Instead, two functions, `convert.rate()` and `scale.rate()`, need to be called separately to convert units and to perform volumetric or mass-specific scaling.
 This unique approach has the advantage of allowing us to perform "dimensionless" rate calculations for data that may not be fully supported by the entire `respR` workflow.
-For example, measurements of respiration can be scaled to surface area (diatoms, Iversen & Ploug, 2013, Naumann et al., 2014), volume (copepod eggs, Hammervold et al., 2015), or density (sea urchin larvae, Stumpp et al., 2011).
-Users may thus first perform rate calculations using `respR`, and then scale the results manually after if there is no support for their specific scaling method.
+For example, measurements of respiration can be scaled to surface area (e.g.
+diatoms, Iversen & Ploug, 2013, Naumann et al., 2014), volume (e.g.
+copepod eggs, Hammervold et al., 2015), or density (e.g.
+sea urchin larvae, Stumpp et al., 2011).
+Users may also be interested in using `respR`'s analytical functions to investigate other linear relationships or perform breakpoint analyses on non-respirometry-related data.
 
-The function `convert.rate()` is used to convert units of oxygen concentration (e.g.
-mg/L or %) only and has support for over 132 unit combinations.
-As unit conventions differ between research disciplines and we do not expect
-
-`scale.rate()` directly accepts object outputs from prior rate-calculating functions and performs volumetric and mass-specific conversions. We have implemented a "semi-fuzzy" approach to
-
-# Reproducibility and data accessibility
-
-We have ensured that any computational analysis performed by `respR` are fully reproducible.
-All main functions produce an output object of class `list`, which contains all the data and variables needed to re-analyse the results, within R or in other software.
+The function `convert.rate()` is used to convert units of oxygen concentration only (e.g.
+$mg/L$ or %) and has support for various unit combinations (Table 2).
+For ease of use, we have implemented a simple fuzzy string-matching algorithm to match relevant units even when they have been provided in different formats (e.g.
+"$mg/L$", "$mg/l$", "$mg L^{-1}$" and "$mg l^{-1}$" are identified as the same units).
+The function `scale.rate()` is conceptually similar to `convert.rate()`, but it focuses volumetric and mass-specific conversions (e.g.
+$mg h^{-1} kg^{-1}$).
 
 # Concluding remarks and future improvements
-While performing the required linear regressions to estimate the parameter of interest is a relatively simple task, a considerable amount of time is spent in data formatting and manipulation.
+The package `respR` implements a number of methods not available in any other R package or software.
+These include the use of rolling regression techniques to estimate $MO_{2,max}$, $MO_{2,min}$, or $P_{crit}$, the use of kernel density estimation techniques to rank linear segments of the data, and the ability to extract and analyse data precicely by time, row, or $O_2$ concentration.
+In particular, there are clear advantages to using `respR` when estimating parameters for $MO_{2,max}$ or $MO_{2,min}$.
+Unlike manual methods where data may be arbitrary selected, `respR` adheres to well-defined rules to obtain rates of interest, and its methods are results are transparent and fully reproducible.
 
@@
 
+<!-- TAB 1 ----------------------------------------------------->
 **Table 1.** Main functions available in `respR`.
+
 ```{r, echo=F, message=F}
@@
 library(readr)
-listfunctions <- read_csv('tables/list-of-functions.csv')
+listfunctions <- readxl::read_excel('tables/list-of-functions.xlsx')
+conversions <- readxl::read_excel("tables/conversion-units.xlsx")
 ```
@@
 
-```{r, echo = F, out.width = "900px"}
+<!-- TAB 2 ----------------------------------------------------->
+**Table 2.** Units available for conversions in `calc.rate()` and `scale.rate()`.
+
+```{r, echo = F, results='asis'}
+pander(conversions)
+
+```
+
+<!-- FIG 1 ----------------------------------------------------->
+
+```{r, echo = F, out.width = "800px"}
 knitr::include_graphics("figs/Figure 1.png")
@@
 
+<!-- FIG 2 ----------------------------------------------------->
+
 ```{r, echo = F, out.width = "650px"}
@@
 2** Illustration showing the sampling method used by `auto.rate()` during **(a)** interval regression analysis, where sample windows do not overlap and **(b)** rolling regression analysis, where sample windows overlap and move forward by one sample unit at a time.
+<!-- FIG 3 ----------------------------------------------------->
 
@@
 3** Typical graphical outputs of the function `auto.rate()`.
-(a) Depending on the `method` input, the function is able to show: (clockwise) the maximum rate detected, the minimum rate detected, calculation of rate by fixed-width intervals, and the most linear region as determined by kernel density estimates.
-(b) In addition, the outputs will also print diagnostic plots of: (clockwise) the rolling regression of rate as a function of time with dotted lines indicating the region used to estimate the parameter, a kernel density plot that shows only when the `linear` method is called, a residual plot of the regression, and a normal Q-Q plot of the residuals of the regression.
-All data is obtained from the example dataset `sardine` which is included in the package.
+**(a)** Depending on the `method` input, the function is able to show: (clockwise form top left) the maximum rate detected, the minimum rate detected, calculation of rate by fixed-width intervals, and the most linear region as determined by kernel density estimates.
+**(b)** In addition, the outputs will also print diagnostic plots of: (clockwise from top left) the rolling regression of rate as a function of time with dotted lines indicating the region used to estimate the parameter, a kernel density plot that shows only when the `linear` method is called, a residual plot of the regression, and a normal Q-Q plot of the residuals of the regression.
+Analysis is performed using the example dataset `sardine`, which is included in the package.
+Code to reproduce the plots is available in the Supplementary Material.
 
-# Supplementary Material
+<!-- FIG 44 ----------------------------------------------------->
+
+```{r, echo = F, , out.width = "800px"}
+knitr::include_graphics("figs/Figure 4.png")
+```
+
+**Fig.
+4** Output plots of the function `pcrit()`, used in the calculation of critical oxygen tension, $P_{crit}$.
+**(a)** To determine $P_{crit}$ in a regular $O_2$ timeseries, the function performs a rolling regression againt the rolling mean of the data.
+This produces **(b)** data that represents the rate of $O_2$ consumption as a function of $O_2$ concentration.
+Regression lines are then fitted iteratively, based on the methods of Yeager and Ultsch (1989) and Muggeo (2003), until convergence is achieved.
+The output then highlights the critical points by colour.
+
+# Supplementary material
+```{r, eval = F}
+# Measuring intermittent data
+# Load data:
+data("intermittent") 
+int <- inspect.data(intermittent) # check for errors
+# Calculate rate in 3 locations:
+int.rate <- calc.rate(int$df, c(200,2300,4100), c(1800,3200,4600), by = 'time')
+plot(int.rate, rep = 1)
+
+# Estimating max, min, interval, most linear (best-fit) rates
+data("sardine")
+sar <- inspect.data(sardine) # check for errors
+sar.max <- auto_rate(sar$df, width = 1500, method = "max") # maximum rate
+sar.min <- auto_rate(sar$df, width = 1500, method = "min") # minimum rate
+
+sdd.int <- auto_rate(sar$df, width = 1500, method = "interval") # interval sampling
+print(sar.int)
+summary(sar.int)
+plot(sar.int, pos = 3) # view the plots for the third interval of the data
+
+sar.lin <- auto_rate(sar$df, width = 1500, method = "linear") # most linear rate
+sar.lin # same as print(sdd.max)
+print(sar.lin, pos = 2) # view rank 2 results
+plot(sar.lin, pos = 2) # plot rank 2 results
+summary(sar.lin) # summary call
+
+# NOTE: using using sample number as roll width is possible and exponentially faster.
+# We can switch to `by = "row"` if time is evenly-spaced.
+sar.max.row <- auto_rate(sar$df, width = 1500, by = "row", method = "max") # maximum rate
+
+# Pcrit analysis
+data(squid)
+sqd <- inspect.data(squid, 2, 3)
+# Perfomr pcrit using defaut
+sqd.pcrit <- pcrit2(sqd$df)
+sqd.pcrit # same as print(sqd.pcrit)
+
+```

### 2017-11-03 manuscript/manuscript-Word.Rmd
manuscript update and new figures

--- 2017-10-16 manuscript/manuscript-Word.Rmd
@@
 library(readr)
+Sys.setenv(TZ="Australia/Sydney")
 
@@
 
-knitr::opts_chunk$set(echo = TRUE)
+opts_chunk$set(collapse = TRUE, comment = "#>", tidy = TRUE, highlight = TRUE)
 ```
+
+# Abstract
+
+1.
+Estimating metabolic rate from respirometry is a conceptually simple, but fundamentally difficult technique where researchers often arbitrarily extract one or more segments of a $O_2$ timeseries to estimate the parameter of interest.
+As respirometry data becomes increasingly large and analytical approaches become more complex, manipulating it remains a challenge for current tools, with important limitations that prevent objectively reproducible research.
+
+2.
+Here we describe the `respR` R package, a collection of functions that implement a workflow-based approach to automate the analysis and visualisation of respirometry data.
+The package can be used for closed, intermittent flow, flowthrough and open respirometry and uses well-defined sets of rules to realiably and rapidly facilitate reproducible results.
+
+3.
+We demonstrate how `respR` uses novel and parallelised computing methods such as rolling regressions and kernel density estimates to sample the data iteratively to realiably detect maximum, minimum & most linear sections of the data, and critical oxygen tension, $P_{crit}$.
+
+4.
+Although designed specifically with aquatic respirometry in mind, the object-oriented approach of the package the "dimensionless" nature of its analytical functions mean that parts of the package could easily be used to estimate any other linear relationships in other research disciplines.
+
+# Introduction
+
+Metabolic rate is a fundamental trait associated with virtually all biological functions and is key in predicting patterns in ecology and conservation biology, from populations (Seibel and Drazen, 2007, Barneche et al., 2014) to ecosystems (Brown et al., 2004).
+It has been widely investigated in studies of the effects of external stressors on organisms, and is increasingly used to study resilience in climate change-related studies where warming is expected to drive increased metabolism in ectotherms (e.g.
+Pörtner, 2002, McElroy et al., 2012, Carey et al., 2016, Delorme & Sewell, 2016).
+By far the most common metric used to assess physiological performance in animals, metabolic rate has been determined from whole organisms, to the level of cells and tissues (White & Kearney, 2013).
+They are typically quantified using respirometry, which measures oxygen uptake over time as a proxy for metabolic rate, since oxygen is rarely stored and anaerobic metabolism is typically sustained over only brief periods (Lighton, 2008).
+
+There are four broad methodological approaches to respirometry: closed, intermittent flow, flowthrough and open.
+In closed respirometry, $O_2$ decrease is measured within a hermetically sealed chamber of known volume, sometimes set within a closed loop to allow mixing of the environment within the chamber.
+Oxygen recordings may be continuous through use of an oxygen probe, periodic through withdrawing water or gas samples at set intervals, or a two-point measurement consisting of the initial and final concentrations.
+Metabolic rates are estimated from respirometry data by assuming a linear relationship between variables, and estimates of metabolic rate are straightforward in constant volume respirometry using the equation: $$VO_2 = \Delta O_2V$$ where $\Delta O_2V$ is the slope of the regression that describes the rate of change in $O_2$ concentration over time, or in the case of a two-point measurement, the difference in $O_2$ concentration divided by time elapsed, and $V$ is the volume of the container (Lighton, 2008).
+<!-- Note, we use here and in the package the term $RO_2$ to denote rate of $O_2$ depletion or uptake within an experiment.
+We use this term used to distinguish it from other common terms, for example $VO_2$ which is can be used to denote both volumetric O2 uptake and whole animal uptake, or $MO_2$ which typically denotes mass-specific uptake.
+-->
+
+In intermittent flow respirometry, $O_2$ concentration is measured as described above, but periodically the chamber is flushed with new water or air, returning it to initial conditions, resealed, and the experiment repeated (Svendsen et al., 2016).
+This technique is essentially closed respirometry, but with the incorporation of repeated measures.
+Depending on the metabolic rate metric being investigated, final respiration rate can be calculated as the mean of the measures (e.g.
+Carey et al., 2016), or the lowest or highest rates recorded in any trial (e.g.
+Stoffels, 2015).
+<!--[nc-will try and find a better ref for that]
+
+Stoffels, R.
+J.
+(2015).
+Physiological trade-offs along a fast-slow lifestyle continuum in fishes: What do they tell us about resistance and resilience to hypoxia?
+Plos One, 10(6), e0130303.
+https://doi.org/10.1371/journal.pone.0130303
+-->
+
+Flowthrough respirometry involves a closed chamber, but with a regulated flow of air or water through it at a precisely determined rate.
+After equilibrium has been achieved, the oxygen concentration differential between the incurrent and excurrent channels, along with the flow rate, allows calculation of the oxygen extracted from the flow volume per unit time: $$VO_2 = (C_iO_2 - C_eO_2)FR$$ where $VO_2$ is the rate of $O_2$ consumption over time, $C_iO_2$ and $C_eO_2$ are the incurrent and excurrent $O_2$ concentrations, and $FR$ is the flow rate of water through the system  (Lighton, 2008).
+
+A final method is open respirometry, in which an open tank or semi-enclosed area is used, but the input or mixing rate of oxygen from the surroundings is known or found to be negligible relative to oxygen consumption of the specimens (Leclercq et al., 1999).
+It is seldom used, but for some applications it is a sufficient and practical methodology (e.g.
+Gamble et al., 2014).
+The common equation used for open respirometry is: $$VO_2 = \Delta O_2V + \phi_d$$ where $\Delta O_2V$ is the slope of the regression that relates $O_2$ concentration to time, $V$ is the volume of the container and $\phi_d$ is the oxygen flux as determined by Fick's Law (Leclercq et al., 1999).
+
+<!--[nc - so the following needs to be spot on.
+Physiologists get very worked up (if not confused themselves) about the differences between these, esp.
+SMR v Routine v Resting v Basal!
+Especially those who work on ecto vs endo therms]-->
+
+Depending on the experiment, different metabolic rate parameters may be of interest to researchers.
+These include maximum metabolic rates (MMR, or $MO_{2,max}$), which define rates under high activity or exhaustive exercise, and minimal metabolic rates ($MO_{2,min}$), which may be termed standard (SMR), basal (BMR), or resting metabolic rates, and typically represent the minimum metabolic cost of maintaining biological functioning (White and Kearney, 2014).
+Routine metabolic rate (RMR), is similar to $MO_{2,min}$, but takes into account that in some organisms energy is expended on small, sponateous movements to maintain posture or position (Rogers et al.
+2016).
+Another metric of great interest is the critical oxygen tension, which represents the lowest level of oxygen at which aerobic metabolism is independent of the ambient partial pressure of oxygen ($P_{crit}$; Yeager & Ultsch, 1989, Hochachka and Somero, 2002).
+
+<!--White, C.
+R., & Kearney, M.
+R.
+(2014).
+Metabolic scaling in animals: methods, empirical results, and theoretical explanations.
+Comprehensive Physiology, 4(1), 231–256.
+https://doi.org/10.1002/cphy.c110049-->
+
+Respirometry experiments are increasingly assimilating large, high-resolution datasets and are run for longer periods (e.g.
+20 h, Norin & Malte, 2012).
+In most cases, processing the data involves an *ad hoc* selection of data points with undefined criteria, and subsequent manual processing of the data subset(s) using a spreadsheet program (e.g.
+Microsoft Excel) or an integrated development environment (IDE, e.g.
+R or Matlab).
+These approaches can be tedious and time consuming especially when spreadsheet programs struggle with the vast datasets that are generated, while IDEs require a degree of expertise to use and have substantial learning curves.
+Dedicated software are also available to perform metabolic rate analyses, but many have costly licensing requirements (e.g.
+AutoResp by Loligo Systems), complicating or preventing their use on multiple machines, and are proprietary and closed-source, hindering scientific reproducability and transparency.
+
+<!-- r.e. below para. LoLinR is not strictly for respirometry - just monotonic, something-changing-over-time data -->
+
+Further, a number of open-source R software packages have recently become available which are designed for, or are suitable to analyse respirometry data.
+The `respirometry` package (https://CRAN.R-project.org/package=respirometry) contains a comprehensive collection of tools to explore and evaluate experimental parameters in aquatic respirometry and is useful for the design and diagnosis of experimental setups.
+The `rMR` package (https://CRAN.R-project.org/package=rMR) performs interval-based metabolic rate calculations and has a method to automatically detect $P_{crit}$ using the "broken-stick" regression method (Yeager & Ultsch, 1989).
+`LoLinR` (Olito et al., 2017), while not strictly coded for respirometry, provides a statistically robust method of detecting a "best-fit" regression that applies to the calculation of metabolic rate, and performs very well at identifying truly linear subsets of a data series.
+Importantly, all of these packages assume that data are monotonic with evenly-spaced sampling periods with respect to time, and do not have the tools to analyse complicated datasets to automatically determine useful metric such as $MO_{2,max}$ and $MO_{2,min}$.
+
+<!-- FIGURE 1 -->
+Here we describe `respR`, an open-source R package desgined to provide an efficient and reproduceable workflow for the analysis of respirometry data.
+The package contains utilities to: (1) analyse closed, intermittent, flow-through and open respirometry data, (2) determine volume, mass, and surface area specific oxygen uptake rates, (3) automatically detect maximal and minimal rates, $P_{crit}$, and (4) detect the most linear sections of data using traditional rolling regression and kernel density estimation techniques.
+Other functions useful in processing respirometry data are avaliable and are described in more detail in our online html vignette (https://januarharianto.github.io/respR/).
+We demonstrate the utility of the package by analysing past data that were collected using closed, intermittent and flow-through respirometry as proof of concept, and show how `respR` might be used to determine $MO_{2,max}$, $MO_{2,min}$, and $P_{crit}$.
+
+# Package overview
+
+The `respR` package streamlines the analysis of respirometry data by partitioning the data processing workflow into several independent actions: (1) initially, data is imported, checked for errors, automatically plotted for visual inspection and then exported as an object; (2) the object is then analysed to estimate the desired rate parameter; (3) diagnostic plots are automatically generated and the data can be further explored using the generic S3 `print()`, `plot()`, and `summary()` commands to verify the results; (4) adjustments to the data, either from background respiration and/or oxygen flux in open respirometry, are applied; (5) volume or mass-specific rate conversions, specific to the respiratory chamber and specimen(s), are calculated where necessary (**Fig.
+1**).
+By design, each step of the workflow requires minimal manual manipulation of the data, and the user can pass the output object of a function to the next.
+The object-oriented approach makes `respR` easy to use, even for novice users of R, and allows us (or anyone interested in the open-source project) to easily extend the functionality of the package, either by adding more intermediate actions or substituting current actions with other code without breaking the rest of the package.
+
+Installation of `respR` can be performed directly in the R console using `devtools`:
+
+```{r eval=F}
+# install.packages("devtools") 
+devtools::install_github("januarharianto/respR")
+library(respR)
+```
+
+To illustrate the main features of the package, 4 example datasets, `sardine.rd`, `urchin.rd`, `intermittent.rd` and `squid.rd` are distributed with `respR`.
+Detailed information about the example data, including their source and methods, can be obtained with the `?` command in the R console (e.g.
+`?sardine.rd`).
+
+# Data import and exploration
+
+Data should be formatted correctly before use in `respR`.
+The function `inspect.data()` extracts a two-column data frame from a multi-column dataset and performs error checks on a data frame of any size, while plotting the data for quick data visualisation.
+Ideally all data are numeric, because `respR` has limited support for date/time data (e.g.
+`POSIXct` and `POSIXlt` classes), and performing calculations on different time formats have many caveats that are beyond the scope of the package.
+Time data, being a continuous variable, should be sequential, without duplicates.
+Ideally, sampling frequency should also be monotonic, but in some cases this may be difficult as machine drift and/or data dropouts can occur during regular sampling.
+In any case, `respR` uses time-based data extraction methods by default in most of its analytical functions (except `pcrit()`), and correctly handle irregular timeseries data by matching the input times to the closes matching values within the data, rounded down.
+
+<!-- ...Identifying irregularities in the sampling time is therefore important as it as variations in data intervals may cause systemic bias in the analysis, especially if data loss or machine drift during sampling is persistent through time and time-based comparisons are made between experiments.
+-->
+
+It should be noted that using `inspect.data()` is optional -- the main functions in our package will readily work with data frames that were not processed by the function.
+Running `inspect.data()` is an exploratory step that flags potential issues about the data before they are analysed.
+Error check results are summarised in the R console and all tests are saved in the output for us to view the locations (i.e.
+row numbers) of the errors:
+
+```{r}
+inspect.data(sardine, plot = FALSE)
+```
+
+## Data analysis
+
+### Calculating background $RO_2$
+
+Background $RO_2$ is often measured to account for the influence of bacterial respiration in the chamber, and applied as a correction to the main experiment (Rodgers et al., 2016).
+Since background rates typically account for a small percentage of experimental rates, multiple “blank” experiments are routinely conducted and the rates are averaged across several datasets to obtain a more accurate estimate of the correction (e.g.
+Carey and Sigwart, 2014; Daoud et al., 2007).
+The function `calc.rate.bg()` uses simple linear regression to simultaneously process multiple background rate measurements.
+Data segments can be truncated before analysis, by either time period or row numbers.
+The results are stored as an object for later use.
+
+### Extracting data and calculating $RO_2$
+
+Typically, the analysis respirometry data requires the selection one or more linear sections of the raw data for more accurate calculations of $RO_2$ (e.g.
+Gordon et al., 1989, Dorey et al., 2014, Chu & Gal, 2017).
+The function `calc.rate()` can manually extract and process data segments from respiromety data.
+Data segments can be selected by (1) time period, (2) row numbers, (3) $O_2$ decrease or (4) proportion, which should accomodate most, if not all, data selection requirements and allow for consistent reporting of methods and results.
+Rates are determined using linear regression, however a two-point analysis is also included alongside all regression results.
+
+For more complex respirometry parameters, the function `auto.rate()` uses novel rolling regression and kernel density estimate algorithms to detect patterns in respirometry data.
+It can calculate maximum or minimum rates, extract rates at non-overlapping intervals, and has an exploratory method to detect the most linear sections of the data.
+When called, the function takes the data frame of length $m$, and performs regressions of sample size $n$ iteratively across the length of the data (i.e.
+rolling regressions), one sample at a time (**Fig.
+2a**).
+Thus, a total of $(m−n)+1$ number of overlapping regressions are fitted, which are then ranked by size to obtain maximum and minimum values.
+If an interval-based regression is required, we selecth the `"interval"` method when running the function and it automatically selects non-overlapping sections of the data for regressions (**Fig.
+2b**).
+
+To estimate the most linear sections of the data, further computations are performed after rolling regressions.
+Logically, linear sections of the data should reflect a consistently stable (i.e.
+flat) rate.
+If the rate changes at some point during the sampling period, thent the rolling estimate should detect the variability.
+The function takes advantage of this relationship and performs a Gaussian kernel density estimate (KDE) of all regressions (see Silverman, 1986).
+Peaks in the density thus reflect stable rates, and new segments of the data within the kernel bandwidth of the peak values are resampled and re-analysed to determine the most linear sections.
+Summary diagnostics and visual plots provide supporting information on the technique and the validity of the results (**Fig.
+3**).
+
+The function `calc.rate.ft()` is used to specifically analyse flowthrough respirometry data, and is similar to `calc.rate()` in functionality.
+However it accepts additional inputs of incurrent and excurrent oxygen concentration, and flow rate, that are not available in `calc.rate()`.`calc.rate.ft()` can process data to determine the mean or modal $RO_2$ over a period.
+
+### Estimating critical oxygen tension, $P_{crit}$
+
+We provide two methods of estimating $P_{crit}$.
+The first is a "broken-stick" regression (BSR) approach, adopted from Yeager and Ultsch (1989) in which two segments of the data are iteratively fitted and the intersection with the smallest sum of the residual sum of squares between the two linear models is the estimated critical point.
+The second method is a wrapper for the segmented, or "broken-line" regression approach, presented by Muggeo (2003) and avaliable as an R package (Muggeo 2008), which estimates the critical point by iteratively fitting two intersecting models on the data and picks the point that minimises the "gap" between the two fitted lines.
+
+To determine $P_{crit}$, $RO_2$ (or $MO_2$) and ambient oxygen concentration ($PO_2$) must be provided.
+If $RO_2$ is not available, but the respirometry data exists to calculate it, `pcrit()` can automatically perform a rolling regression directly on the data to generate $RO_2$ and arrange it against the rolling mean of $PO_2$ (**Fig.
+4**), before calculating $P_{crit}$.
+
+## Correcting, converting and scaling data
+
+### Corrections for oxygen flux and background respiration
+
+The function `adjust.rate()` can be used to correct $RO_2$ estimates to oxygen flux and/or background respirometry data.
+As the oxygen consumption units are not mass-specific at the point of correction, the function adjusts $RO_2$ directly without bias using the equation: $$RO_{2(correction)} = RO_2 - (\phi_d + RO_{2(bg)})$$ where $\phi_d$ is the oxygen flux and $RO_{2(bg)}$ is background respiration.
+
+### Rate conversions
+
+For delibrate reasons, units of measurements are not used when estimating rates in R.
+Instead, two functions, `convert.rate()` and `scale.rate()`, need to be called separately to convert units and to perform volumetric or mass-specific scaling.
+This unique approach has the advantage of allowing us to perform "dimensionless" rate calculations for data that may not be fully supported by the entire `respR` workflow.
+For example, measurements of respiration can be scaled to surface area (e.g.
+diatoms, Iversen & Ploug, 2013, Naumann et al., 2014), volume (e.g.
+copepod eggs, Hammervold et al., 2015), or density (e.g.
+sea urchin larvae, Stumpp et al., 2011).
+Users may also be interested in using `respR`'s analytical functions to investigate other linear relationships or perform breakpoint analyses on non-respirometry-related data.
+
+The function `convert.rate()` is used to convert units of oxygen concentration only (e.g.
+$mg/L$ or %) and has support for various unit combinations (Table 2).
+For ease of use, we have implemented a simple fuzzy string-matching algorithm to match relevant units even when they have been provided in different formats (e.g.
+"$mg/L$", "$mg/l$", "$mg L^{-1}$" and "$mg l^{-1}$" are identified as the same units).
+The function `scale.rate()` is conceptually similar to `convert.rate()`, but it focuses volumetric and mass-specific conversions (e.g.
+$mg h^{-1} kg^{-1}$).
+
+# Concluding remarks and future improvements
+The package `respR` implements a number of methods not available in any other R package or software.
+These include the use of rolling regression techniques to estimate $MO_{2,max}$, $MO_{2,min}$, or $P_{crit}$, the use of kernel density estimation techniques to rank linear segments of the data, and the ability to extract and analyse data precicely by time, row, or $O_2$ concentration.
+In particular, there are clear advantages to using `respR` when estimating parameters for $MO_{2,max}$ or $MO_{2,min}$.
+Unlike manual methods where data may be arbitrary selected, `respR` adheres to well-defined rules to obtain rates of interest, and its methods are results are transparent and fully reproducible.
+
+# Tables & Figures
+
+<!-- TAB 1 ----------------------------------------------------->
+**Table 1.** Main functions available in `respR`.
+
+```{r, echo=F, message=F}
+# use this later to show table
+library(readr)
+listfunctions <- readxl::read_excel('tables/list-of-functions.xlsx')
+conversions <- readxl::read_excel("tables/conversion-units.xlsx")
+```
+
+```{r, echo=F, results='asis'}
+pander(listfunctions)
+```
+
+<!-- TAB 2 ----------------------------------------------------->
+**Table 2.** Units available for conversions in `calc.rate()` and `scale.rate()`.
+
+```{r, echo = F, results='asis'}
+pander(conversions)
+
+```
+
+<!-- FIG 1 ----------------------------------------------------->
+
+```{r, echo = F, out.width = "800px"}
+knitr::include_graphics("figs/Figure 1.png")
+```
+
+**Fig.
+1.** Diagram showing a typical workflow in the analysis of respirometry data using `respR`.
+Data is first checked for errors before the main functions are used to extract and analyse segments of the data.
+Summarised results and diagnostic plots provide immediate visual feedback on the outcome of the analyses.
+Once the rate estimates are obtained, they can be converted into volume and/or mass-specific rates.
+
+<!-- FIG 2 ----------------------------------------------------->
+
+```{r, echo = F, out.width = "650px"}
+knitr::include_graphics("figs/Figure 2.png")
+```
+
+**Fig.
+2** Illustration showing the sampling method used by `auto.rate()` during **(a)** interval regression analysis, where sample windows do not overlap and **(b)** rolling regression analysis, where sample windows overlap and move forward by one sample unit at a time.
+<!-- FIG 3 ----------------------------------------------------->
+
+```{r, echo = F, , out.width = "800px"}
+knitr::include_graphics("figs/Figure 3.png")
+```
+
+**Fig.
+3** Typical graphical outputs of the function `auto.rate()`.
+**(a)** Depending on the `method` input, the function is able to show: (clockwise form top left) the maximum rate detected, the minimum rate detected, calculation of rate by fixed-width intervals, and the most linear region as determined by kernel density estimates.
+**(b)** In addition, the outputs will also print diagnostic plots of: (clockwise from top left) the rolling regression of rate as a function of time with dotted lines indicating the region used to estimate the parameter, a kernel density plot that shows only when the `linear` method is called, a residual plot of the regression, and a normal Q-Q plot of the residuals of the regression.
+Analysis is performed using the example dataset `sardine`, which is included in the package.
+Code to reproduce the plots is available in the Supplementary Material.
+
+<!-- FIG 44 ----------------------------------------------------->
+
+```{r, echo = F, , out.width = "800px"}
+knitr::include_graphics("figs/Figure 4.png")
+```
+
+**Fig.
+4** Output plots of the function `pcrit()`, used in the calculation of critical oxygen tension, $P_{crit}$.
+**(a)** To determine $P_{crit}$ in a regular $O_2$ timeseries, the function performs a rolling regression againt the rolling mean of the data.
+This produces **(b)** data that represents the rate of $O_2$ consumption as a function of $O_2$ concentration.
+Regression lines are then fitted iteratively, based on the methods of Yeager and Ultsch (1989) and Muggeo (2003), until convergence is achieved.
+The output then highlights the critical points by colour.
+
+# Supplementary material
+```{r, eval = F}
+# Measuring intermittent data
+# Load data:
+data("intermittent") 
+int <- inspect.data(intermittent) # check for errors
+# Calculate rate in 3 locations:
+int.rate <- calc.rate(int$df, c(200,2300,4100), c(1800,3200,4600), by = 'time')
+plot(int.rate, rep = 1)
+
+# Estimating max, min, interval, most linear (best-fit) rates
+data("sardine")
+sar <- inspect.data(sardine) # check for errors
+sar.max <- auto_rate(sar$df, width = 1500, method = "max") # maximum rate
+sar.min <- auto_rate(sar$df, width = 1500, method = "min") # minimum rate
+
+sdd.int <- auto_rate(sar$df, width = 1500, method = "interval") # interval sampling
+print(sar.int)
+summary(sar.int)
+plot(sar.int, pos = 3) # view the plots for the third interval of the data
+
+sar.lin <- auto_rate(sar$df, width = 1500, method = "linear") # most linear rate
+sar.lin # same as print(sdd.max)
+print(sar.lin, pos = 2) # view rank 2 results
+plot(sar.lin, pos = 2) # plot rank 2 results
+summary(sar.lin) # summary call
+
+# NOTE: using using sample number as roll width is possible and exponentially faster.
+# We can switch to `by = "row"` if time is evenly-spaced.
+sar.max.row <- auto_rate(sar$df, width = 1500, by = "row", method = "max") # maximum rate
+
+# Pcrit analysis
+data(squid)
+sqd <- inspect.data(squid, 2, 3)
+# Perfomr pcrit using defaut
+sqd.pcrit <- pcrit2(sqd$df)
+sqd.pcrit # same as print(sqd.pcrit)
+
+```

### 2017-11-09 manuscript/manuscript-HTML.Rmd
manuscript updates

@@
 1.
-Estimating metabolic rate from respirometry is a conceptually simple, but fundamentally difficult technique where researchers often arbitrarily extract one or more segments of a $O_2$ timeseries to estimate the parameter of interest.
-As respirometry data becomes increasingly large and analytical approaches become more complex, manipulating it remains a challenge for current tools, with important limitations that prevent objectively reproducible research.
+Respirometry is a ubiquitous practice in experimental biology, but there is a lack of standard practices when analysing the resulting data, limiting transparency and reproducibility.
+As respirometry datasets become increasingly large and analytical approaches more complex, manipulating the data remains a challenge and often intractable with existing tools.
 
@@
 Here we describe the `respR` R package, a collection of functions that implement a workflow-based approach to automate the analysis and visualisation of respirometry data.
-The package can be used for closed, intermittent flow, flowthrough and open respirometry and uses well-defined sets of rules to realiably and rapidly facilitate reproducible results.
+The package can be used for closed, intermittent flow, flow-through and open respirometry and uses well-defined sets of rules to reliably and rapidly generate reproducible results.
 
 3.
-We demonstrate how `respR` uses novel and parallelised computing methods such as rolling regressions and kernel density estimates to sample the data iteratively to realiably detect maximum, minimum & most linear sections of the data, and critical oxygen tension, $P_{crit}$.
+We demonstrate how `respR` uses novel and parallelised computing methods such as rolling regressions and kernel density estimates to reliably detect maximum, minimum and most linear sections of the data, and critical oxygen tension, $P_{crit}$.
 
 4.
-Although designed specifically with aquatic respirometry in mind, the object-oriented approach of the package the "dimensionless" nature of its analytical functions mean that parts of the package could easily be used to estimate any other linear relationships in other research disciplines.
+Although designed specifically with aquatic respirometry in mind, the object-oriented approach of the package and the unitless nature of its analytical functions mean that parts of the package can easily be used to estimate any other linear relationships in other research disciplines.
 
@@
 Pörtner, 2002, McElroy et al., 2012, Carey et al., 2016, Delorme & Sewell, 2016).
-By far the most common metric used to assess physiological performance in animals, metabolic rate has been determined from whole organisms, to the level of cells and tissues (White & Kearney, 2013).
-They are typically quantified using respirometry, which measures oxygen uptake over time as a proxy for metabolic rate, since oxygen is rarely stored and anaerobic metabolism is typically sustained over only brief periods (Lighton, 2008).
+By far the most common metric used to assess physiological performance in animals, metabolic rate has been determined from whole organisms to the level of cells and tissues (White & Kearney, 2013).
+Metabolic rate is typically quantified using respirometry, which measures oxygen uptake over time (Lighton, 2008).
 
-There are four broad methodological approaches to respirometry: closed, intermittent flow, flowthrough and open.
+There are four broad methodological approaches in respirometry: closed, intermittent flow, flow-through and open.
 In closed respirometry, $O_2$ decrease is measured within a hermetically sealed chamber of known volume, sometimes set within a closed loop to allow mixing of the environment within the chamber.
 Oxygen recordings may be continuous through use of an oxygen probe, periodic through withdrawing water or gas samples at set intervals, or a two-point measurement consisting of the initial and final concentrations.
-Metabolic rates are estimated from respirometry data by assuming a linear relationship between variables, and estimates of metabolic rate are straightforward in constant volume respirometry using the equation: $$VO_2 = \Delta O_2V$$ where $\Delta O_2V$ is the slope of the regression that describes the rate of change in $O_2$ concentration over time, or in the case of a two-point measurement, the difference in $O_2$ concentration divided by time elapsed, and $V$ is the volume of the container (Lighton, 2008).
+Metabolic rates are estimated from the $O_2$ timeseries by assuming a linear relationship between variables, and estimates of metabolic rate are straightforward in constant volume respirometry using the equation: $$VO_2 = \Delta O_2V$$ where $\Delta O_2V$ is the slope of the regression that describes the rate of change in $O_2$ concentration over time, or in the case of a two-point measurement, the difference in $O_2$ concentration divided by time elapsed, and $V$ is the volume of the container (Lighton, 2008).
 <!-- Note, we use here and in the package the term $RO_2$ to denote rate of $O_2$ depletion or uptake within an experiment.
@@
 
-Flowthrough respirometry involves a closed chamber, but with a regulated flow of air or water through it at a precisely determined rate.
+Flow-through respirometry involves a closed chamber, but with a regulated flow of air or water through it at a precisely determined rate.
 After equilibrium has been achieved, the oxygen concentration differential between the incurrent and excurrent channels, along with the flow rate, allows calculation of the oxygen extracted from the flow volume per unit time: $$VO_2 = (C_iO_2 - C_eO_2)FR$$ where $VO_2$ is the rate of $O_2$ consumption over time, $C_iO_2$ and $C_eO_2$ are the incurrent and excurrent $O_2$ concentrations, and $FR$ is the flow rate of water through the system  (Lighton, 2008).
@@
 Depending on the experiment, different metabolic rate parameters may be of interest to researchers.
-These include maximum metabolic rates (MMR, or $MO_{2,max}$), which define rates under high activity or exhaustive exercise, and minimal metabolic rates ($MO_{2,min}$), which may be termed standard (SMR), basal (BMR), or resting metabolic rates, and typically represent the minimum metabolic cost of maintaining biological functioning (White and Kearney, 2014).
-Routine metabolic rate (RMR), is similar to $MO_{2,min}$, but takes into account that in some organisms energy is expended on small, sponateous movements to maintain posture or position (Rogers et al.
+These include maximum metabolic rates (MMR, or $MO_{2,max}$), rates under high activity or exhaustive exercise, and minimal metabolic rates ($MO_{2,min}$), which may be termed standard (SMR), basal (BMR), or resting metabolic rates, and typically represents the minimum metabolic cost of maintaining biological functioning (White and Kearney, 2014).
+Routine metabolic rate (RMR), is similar to $MO_{2,min}$, but takes into account that in some organisms energy is expended on small, spontaneous movements to maintain posture or position (Rogers et al.
 2016).
@@
 
-Respirometry experiments are increasingly assimilating large, high-resolution datasets and are run for longer periods (e.g.
+Respirometry studies are increasingly outputting large, high-resolution datasets and conducted over longer periods (e.g.
 20 h, Norin & Malte, 2012).
-In most cases, processing the data involves an *ad hoc* selection of data points with undefined criteria, and subsequent manual processing of the data subset(s) using a spreadsheet program (e.g.
+In most cases, processing the data involves an *ad hoc* selection of data points with poorly reported criteria, and subsequent manual processing of the data subset(s) using a spreadsheet program (e.g.
 Microsoft Excel) or an integrated development environment (IDE, e.g.
@@
 Dedicated software are also available to perform metabolic rate analyses, but many have costly licensing requirements (e.g.
-AutoResp by Loligo Systems), complicating or preventing their use on multiple machines, and are proprietary and closed-source, hindering scientific reproducability and transparency.
+AutoResp by Loligo Systems), complicating or preventing their use on multiple machines, and are proprietary and closed-source, hindering scientific reproducibility and transparency.
 
@@
 
-Further, a number of open-source R software packages have recently become available which are designed for, or are suitable to analyse respirometry data.
+A number of open-source R software packages have recently become available which are designed for, or are suitable to analyse respirometry data.
 The `respirometry` package (https://CRAN.R-project.org/package=respirometry) contains a comprehensive collection of tools to explore and evaluate experimental parameters in aquatic respirometry and is useful for the design and diagnosis of experimental setups.
 The `rMR` package (https://CRAN.R-project.org/package=rMR) performs interval-based metabolic rate calculations and has a method to automatically detect $P_{crit}$ using the "broken-stick" regression method (Yeager & Ultsch, 1989).
-`LoLinR` (Olito et al., 2017), while not strictly coded for respirometry, provides a statistically robust method of detecting a "best-fit" regression that applies to the calculation of metabolic rate, and performs very well at identifying truly linear subsets of a data series.
-Importantly, all of these packages assume that data are monotonic with evenly-spaced sampling periods with respect to time, and do not have the tools to analyse complicated datasets to automatically determine useful metric such as $MO_{2,max}$ and $MO_{2,min}$.
+`LoLinR` (Olito et al., 2017), while not specifically coded for respirometry, provides a statistically robust method of detecting a "best-fit" regression, and performs very well at identifying truly linear subsets of a data series.
+Importantly, all of these packages assume that data are monotonic with evenly-spaced sampling periods with respect to time, an ideal structure but one where respirometry data often fails to meet in reality.
+They also do not have the tools to analyse complicated datasets to automatically determine useful metric such as $MO_{2,max}$ and $MO_{2,min}$.
 
 <!-- FIGURE 1 -->
-Here we describe `respR`, an open-source R package desgined to provide an efficient and reproduceable workflow for the analysis of respirometry data.
+Here we describe `respR`, an open-source R package designed to provide an efficient and reproducible workflow for the analysis of respirometry data.
 The package contains utilities to: (1) analyse closed, intermittent, flow-through and open respirometry data, (2) determine volume, mass, and surface area specific oxygen uptake rates, (3) automatically detect maximal and minimal rates, $P_{crit}$, and (4) detect the most linear sections of data using traditional rolling regression and kernel density estimation techniques.
+The package facilitates the output as supplementary material of all stages of respirometry analyses; raw data, subsetting criteria, background corrections, units of measure, and resulting metabolic rates.
+This streamlines the reporting of analytical methods and allows reviewers or other investigators to fully reproduce or scrutinise the results.
 Other functions useful in processing respirometry data are avaliable and are described in more detail in our online html vignette (https://januarharianto.github.io/respR/).
-We demonstrate the utility of the package by analysing past data that were collected using closed, intermittent and flow-through respirometry as proof of concept, and show how `respR` might be used to determine $MO_{2,max}$, $MO_{2,min}$, and $P_{crit}$.
+The utility of the `respR` package is shown here using  data collected for closed, intermittent and flow-through respirometry.
+We show how `respR` might be used to determine $MO_{2,max}$, $MO_{2,min}$, and $P_{crit}$.
 
@@
 
-The `respR` package streamlines the analysis of respirometry data by partitioning the data processing workflow into several independent actions: (1) initially, data is imported, checked for errors, automatically plotted for visual inspection and then exported as an object; (2) the object is then analysed to estimate the desired rate parameter; (3) diagnostic plots are automatically generated and the data can be further explored using the generic S3 `print()`, `plot()`, and `summary()` commands to verify the results; (4) adjustments to the data, either from background respiration and/or oxygen flux in open respirometry, are applied; (5) volume or mass-specific rate conversions, specific to the respiratory chamber and specimen(s), are calculated where necessary (**Fig.
+The `respR` package streamlines the analysis of respirometry data by partitioning the data processing workflow into several independent actions: (1) initially, data are imported, checked for common errors, plotted for visual inspection and exported as an object; (2) the object is then analysed to estimate the desired rate parameter; (3) diagnostic plots are automatically generated and the data can be further explored using the generic S3 `print()`, `plot()`, and `summary()` commands to verify the results; (4) adjustments to the data, either from background respiration and/or oxygen flux in open respirometry, are applied; (5) volume or mass-specific rate conversions, specific to the respiratory chamber and specimen(s), are calculated where necessary (**Fig.
 1**).
-By design, each step of the workflow requires minimal manual manipulation of the data, and the user can pass the output object of a function to the next.
-The object-oriented approach makes `respR` easy to use, even for novice users of R, and allows us (or anyone interested in the open-source project) to easily extend the functionality of the package, either by adding more intermediate actions or substituting current actions with other code without breaking the rest of the package.
+By design, each step of the workflow requires minimal manual manipulation of the data, and the user can pass the output of each function to the next.
+This object-oriented approach makes `respR` easy to use, even for novice users of R, and allows us (or anyone interested in the open-source project) to easily extend the functionality of the package, either by adding additional intermediate actions or substituting current actions with other code without breaking the rest of the package.
+Note, we use $RO_2$ in the following text to represent the dimensionless rate of $O_2$ use with regard to the time unit, and to distinguish it from other commonly reported metrics (e.g.
+$MO_2$, $VO_2$).
 
-Installation of `respR` can be performed directly in the R console using `devtools`:
+Installation of `respR` can be performed directly via the R console using `devtools`:
 
@@
 The function `inspect.data()` extracts a two-column data frame from a multi-column dataset and performs error checks on a data frame of any size, while plotting the data for quick data visualisation.
-Ideally all data are numeric, because `respR` has limited support for date/time data (e.g.
-`POSIXct` and `POSIXlt` classes), and performing calculations on different time formats have many caveats that are beyond the scope of the package.
+All data should be numeric; respR at this time has limited support for date/time data (e.g.
+`POSIXct` and `POSIXt` classes).
+Parsing  different time formats is extremely complex and beyond the scope of this package, particularly because such packages already exist (e.g.
+`lubridate`).
 Time data, being a continuous variable, should be sequential, without duplicates.
-Ideally, sampling frequency should also be monotonic, but in some cases this may be difficult as machine drift and/or data dropouts can occur during regular sampling.
-In any case, `respR` uses time-based data extraction methods by default in most of its analytical functions (except `pcrit()`), and correctly handle irregular timeseries data by matching the input times to the closes matching values within the data, rounded down.
+Ideally, sampling frequency will be monotonic, but in many cases this may not be the case, such as experiments with irregular sampling intervals, or because machine drift and/or data dropouts frequently occur during regular sampling.
+Therefore, `respR` uses time-based data extraction methods by default in most of its analytical functions (except `pcrit()`), and correctly handles irregular timeseries data.
 
@@
 
-It should be noted that using `inspect.data()` is optional -- the main functions in our package will readily work with data frames that were not processed by the function.
-Running `inspect.data()` is an exploratory step that flags potential issues about the data before they are analysed.
-Error check results are summarised in the R console and all tests are saved in the output for us to view the locations (i.e.
+Using `inspect.data()` is optional -- the main functions in our package will readily work with data frames that were not processed by the function.
+Running `inspect.data()` is an exploratory step that flags potential issues before analysis.
+Error check results are summarised in the R console and all tests are saved in the output object with the locations (i.e.
 row numbers) of the errors:
@@
 ```{r}
-inspect.data(sardine, plot = FALSE)
+inspect.data(squid, xcol = 2, ycol = 3, plot = FALSE, highlight = T)
 ```
@@
 
-Background $RO_2$ is often measured to account for the influence of bacterial respiration in the chamber, and applied as a correction to the main experiment (Rodgers et al., 2016).
-Since background rates typically account for a small percentage of experimental rates, multiple “blank” experiments are routinely conducted and the rates are averaged across several datasets to obtain a more accurate estimate of the correction (e.g.
+Background oxygen consumption rate ($RO_({2(bg)}$) is often measured to account for the contribution of microbial respiration, and applied as a correction to experimental rates (Rodgers et al., 2016).
+Background rates typically account for a small percentage of experimental rates, and multiple “blank” experiments are usually conducted and the rates averaged to obtain a more accurate estimate of the correction (e.g.
 Carey and Sigwart, 2014; Daoud et al., 2007).
 The function `calc.rate.bg()` uses simple linear regression to simultaneously process multiple background rate measurements.
-Data segments can be truncated before analysis, by either time period or row numbers.
+Data segments can be truncated before analysis by time period or row numbers.
 The results are stored as an object for later use.
@@
 Gordon et al., 1989, Dorey et al., 2014, Chu & Gal, 2017).
-The function `calc.rate()` can manually extract and process data segments from respiromety data.
-Data segments can be selected by (1) time period, (2) row numbers, (3) $O_2$ decrease or (4) proportion, which should accomodate most, if not all, data selection requirements and allow for consistent reporting of methods and results.
+The function `calc.rate()` can manually extract and process data subset from respiromety data.
+Data subsets can be selected by (1) time period, (2) row numbers, (3) $O_2$ decrease or (4) proportion, which should accommodate most, if not all, manual data selection requirements and allow for consistent reporting of methods and results.
 Rates are determined using linear regression, however a two-point analysis is also included alongside all regression results.
@@
 2a**).
-Thus, a total of $(m−n)+1$ number of overlapping regressions are fitted, which are then ranked by size to obtain maximum and minimum values.
-If an interval-based regression is required, we selecth the `"interval"` method when running the function and it automatically selects non-overlapping sections of the data for regressions (**Fig.
+Thus, a total of $(m−n)+1$ number of overlapping regressions are fitted, which are then ranked to obtain maximum and minimum values.
+If an interval-based regression is required, we select the `"interval"` method when calling the function, and it automatically selects non-overlapping sections of the data for regressions (**Fig.
 2b**).
@@
 flat) rate.
-If the rate changes at some point during the sampling period, thent the rolling estimate should detect the variability.
-The function takes advantage of this relationship and performs a Gaussian kernel density estimate (KDE) of all regressions (see Silverman, 1986).
+If the rate changes at some point during the sampling period, then the rolling estimate should detect the variability.
+This function takes advantage of this relationship and performs a Gaussian kernel density estimate (KDE) of all regressions (see Silverman, 1986).
 Peaks in the density thus reflect stable rates, and new segments of the data within the kernel bandwidth of the peak values are resampled and re-analysed to determine the most linear sections.
@@
 
-The function `calc.rate.ft()` is used to specifically analyse flowthrough respirometry data, and is similar to `calc.rate()` in functionality.
-However it accepts additional inputs of incurrent and excurrent oxygen concentration, and flow rate, that are not available in `calc.rate()`.`calc.rate.ft()` can process data to determine the mean or modal $RO_2$ over a period.
+The function `calc.rate.ft()` is used to specifically analyse flow-through respirometry data, and is similar to `calc.rate()` in functionality.
+However it accepts additional inputs of incurrent and excurrent oxygen concentration, and flow rate.`calc.rate.ft()` can process data to determine the mean or modal $RO_2$ over a period.
 
@@
 The first is a "broken-stick" regression (BSR) approach, adopted from Yeager and Ultsch (1989) in which two segments of the data are iteratively fitted and the intersection with the smallest sum of the residual sum of squares between the two linear models is the estimated critical point.
-The second method is a wrapper for the segmented, or "broken-line" regression approach, presented by Muggeo (2003) and avaliable as an R package (Muggeo 2008), which estimates the critical point by iteratively fitting two intersecting models on the data and picks the point that minimises the "gap" between the two fitted lines.
+The second method is a wrapper for the segmented, or "broken-line" regression approach, presented by Muggeo (2003) and available as an R package (Muggeo 2008), which estimates the critical point by iteratively fitting two intersecting models on the data and selecting the point that minimises the "gap" between the two fitted lines.
 
-To determine $P_{crit}$, $RO_2$ (or $MO_2$) and ambient oxygen concentration ($PO_2$) must be provided.
-If $RO_2$ is not available, but the respirometry data exists to calculate it, `pcrit()` can automatically perform a rolling regression directly on the data to generate $RO_2$ and arrange it against the rolling mean of $PO_2$ (**Fig.
-4**), before calculating $P_{crit}$.
+As $P_{crit}$ is the break-point in the relationship between ambient dissolved oxygen ($DO_2$) and $RO_2$/$MO_2$, the user has two options to provide the required data.
+Loading an $O_2$ timeseries data into the function will prompt it to automatically perform a rolling mean and a rolling regression of $DO_2$ to extract those variables (**Fig.
+4**).
+Alternatively, the user already has the required data, and the function may run the data directly using the argument `has.rate = TRUE`.
 
@@
 
-The function `adjust.rate()` can be used to correct $RO_2$ estimates to oxygen flux and/or background respirometry data.
-As the oxygen consumption units are not mass-specific at the point of correction, the function adjusts $RO_2$ directly without bias using the equation: $$RO_{2(correction)} = RO_2 - (\phi_d + RO_{2(bg)})$$ where $\phi_d$ is the oxygen flux and $RO_{2(bg)}$ is background respiration.
+The function `adjust.rate()` can be used to correct $RO_2$ estimates for oxygen flux and/or background respirometry data.
+As the oxygen consumption units are not mass specific at the point of correction, the function adjusts RO_2 directly without bias using the equation: $$RO_{2(correction)} = RO_2 - (\phi_d + RO_{2(bg)})$$ where $\phi_d$ is the oxygen flux and $RO_{2(bg)}$ is background respiration.
 
@@
 
-For delibrate reasons, units of measurements are not used when estimating rates in R.
-Instead, two functions, `convert.rate()` and `scale.rate()`, need to be called separately to convert units and to perform volumetric or mass-specific scaling.
-This unique approach has the advantage of allowing us to perform "dimensionless" rate calculations for data that may not be fully supported by the entire `respR` workflow.
+For deliberate reasons, units of measurement are not used when determining $RO_2$ in `respR`.
+Instead, two functions, `convert.rate()` and `scale.rate()`, are used to transform $RO_2$ to specific units and perform volumetric or mass-specific conversions .
+This unique approach has the advantage of allowing "dimensionless" rate calculations for data that may not be fully supported by the entire respR workflow.
 For example, measurements of respiration can be scaled to surface area (e.g.
@@
 copepod eggs, Hammervold et al., 2015), or density (e.g.
-sea urchin larvae, Stumpp et al., 2011).
-Users may also be interested in using `respR`'s analytical functions to investigate other linear relationships or perform breakpoint analyses on non-respirometry-related data.
+sea urchin larvae, Stumpp et al., 2011), which are currently unsupported.
+Users may also be interested in using respR's analytical functions to investigate other linear relationships or perform breakpoint analyses on non-respirometry-related data.
 
@@
 The package `respR` implements a number of methods not available in any other R package or software.
-These include the use of rolling regression techniques to estimate $MO_{2,max}$, $MO_{2,min}$, or $P_{crit}$, the use of kernel density estimation techniques to rank linear segments of the data, and the ability to extract and analyse data precicely by time, row, or $O_2$ concentration.
+These include the use of rolling regression techniques to estimate $MO_{2,max}$, $MO_{2,min}$, and $P_{crit}$, the use of kernel density estimation techniques to rank linear segments of the data, support for unevenly-spaced time data, and the ability to extract and analyse data precicely by time, row, or $O_2$ concentration.
 In particular, there are clear advantages to using `respR` when estimating parameters for $MO_{2,max}$ or $MO_{2,min}$.
-Unlike manual methods where data may be arbitrary selected, `respR` adheres to well-defined rules to obtain rates of interest, and its methods are results are transparent and fully reproducible.
+Unlike manual methods where data ranges may be arbitrary selected, `respR` adheres to well-defined rules to obtain rates of interest, and its methods are results are transparent and fully reproducible.
+The `respR` package allows for more effective and less time-consuming handling and processing of the very large data sets that are increasingly common in respirometry research.
 
@@
 **Fig.
-1.** Diagram showing a typical workflow in the analysis of respirometry data using `respR`.
-Data is first checked for errors before the main functions are used to extract and analyse segments of the data.
+1.** Diagram showing steps in a typical workflow in the analysis of respirometry data using `respR`.
+Data are first checked for errors before the main functions are used to extract and analyse segments of the data.
 Summarised results and diagnostic plots provide immediate visual feedback on the outcome of the analyses.
@@
 **Fig.
-2** Illustration showing the sampling method used by `auto.rate()` during **(a)** interval regression analysis, where sample windows do not overlap and **(b)** rolling regression analysis, where sample windows overlap and move forward by one sample unit at a time.
+2** Illustration showing the sampling method used by `auto.rate()` during **(A)** interval regression analysis, where sample windows do not overlap and **(B)** rolling regression analysis, where sample windows overlap and move forward by one sample unit at a time.
 <!-- FIG 3 ----------------------------------------------------->
@@
 3** Typical graphical outputs of the function `auto.rate()`.
-**(a)** Depending on the `method` input, the function is able to show: (clockwise form top left) the maximum rate detected, the minimum rate detected, calculation of rate by fixed-width intervals, and the most linear region as determined by kernel density estimates.
-**(b)** In addition, the outputs will also print diagnostic plots of: (clockwise from top left) the rolling regression of rate as a function of time with dotted lines indicating the region used to estimate the parameter, a kernel density plot that shows only when the `linear` method is called, a residual plot of the regression, and a normal Q-Q plot of the residuals of the regression.
+**(A)** Depending on the `method` input, the function is able to show: (clockwise form top left) the maximum rate detected, the minimum rate detected, calculation of rate by fixed-width intervals, and the most linear region as determined by kernel density estimates.
+**(B)** In addition, the outputs will also print diagnostic plots of: (clockwise from top left) the rolling regression of rate as a function of time with dotted lines indicating the region used to estimate the parameter, a kernel density plot that shows only when the `linear` method is called, a residual plot of the regression, and a normal Q-Q plot of the residuals of the regression.
 Analysis is performed using the example dataset `sardine`, which is included in the package.
@@
 4** Output plots of the function `pcrit()`, used in the calculation of critical oxygen tension, $P_{crit}$.
-**(a)** To determine $P_{crit}$ in a regular $O_2$ timeseries, the function performs a rolling regression againt the rolling mean of the data.
-This produces **(b)** data that represents the rate of $O_2$ consumption as a function of $O_2$ concentration.
-Regression lines are then fitted iteratively, based on the methods of Yeager and Ultsch (1989) and Muggeo (2003), until convergence is achieved.
-The output then highlights the critical points by colour.
+**(A)** To determine $P_{crit}$ in a regular $O_2$ timeseries, the function performs a rolling regression againt the rolling mean of the data.
+This produces **(B)** data that represents the rate of $O_2$ consumption as a function of $O_2$ concentration.
+Regression lines are then fitted iteratively, based on the broken-stick methods of Yeager and Ultsch (1989) and the segmented approach of Muggeo (2003), until convergence is achieved.
+The output then highlights the critical points as indicated by the coloured lines.
 
@@
 ```{r, eval = F}
+
 # Measuring intermittent data
@@
 # Calculate rate in 3 locations:
-int.rate <- calc.rate(int$df, c(200,2300,4100), c(1800,3200,4600), by = 'time')
+int.rate <- calc.rate(int, c(200,2300,4100), c(1800,3200,4600), by = 'time')
 plot(int.rate, rep = 1)
@@
 sar <- inspect.data(sardine) # check for errors
-sar.max <- auto_rate(sar$df, width = 1500, method = "max") # maximum rate
-sar.min <- auto_rate(sar$df, width = 1500, method = "min") # minimum rate
+sar.max <- auto.rate(sar, width = 1500, method = "max") # maximum rate
+sar.min <- auto.rate(sar, width = 1500, method = "min") # minimum rate
 
-sdd.int <- auto_rate(sar$df, width = 1500, method = "interval") # interval sampling
+sar.int <- auto.rate(sar, width = 1500, method = "interval") # interval sampling
 print(sar.int)
@@
 
-sar.lin <- auto_rate(sar$df, width = 1500, method = "linear") # most linear rate
+sar.lin <- auto.rate(sar, width = 1500, method = "linear") # most linear rate
 sar.lin # same as print(sdd.max)
@@
 # We can switch to `by = "row"` if time is evenly-spaced.
-sar.max.row <- auto_rate(sar$df, width = 1500, by = "row", method = "max") # maximum rate
+sar.max.row <- auto.rate(sar, width = 1500, by = "row", method = "max") # maximum rate
 
@@
 # Perfomr pcrit using defaut
-sqd.pcrit <- pcrit2(sqd$df)
+sqd.pcrit <- pcrit(sqd) # this may take a while as over 30k regressions are performed
 sqd.pcrit # same as print(sqd.pcrit)

### 2017-11-12 manuscript/manuscript-HTML.Rmd
manuscript update

@@
 3.
-We demonstrate how `respR` uses novel and parallelised computing methods such as rolling regressions and kernel density estimates to reliably detect maximum, minimum and most linear sections of the data, and critical oxygen tension, $P_{crit}$.
+We demonstrate how `respR` uses novel computing methods such as rolling regressions and kernel density estimates to reliably detect maximum, minimum and most linear sections of the data, and critical oxygen tension, $P_{crit}$.
 
@@
 ```{r}
-inspect.data(squid, xcol = 2, ycol = 3, plot = FALSE, highlight = T)
+inspect.data(squid, plot = F)
 ```

### 2017-11-13 manuscript/LaTeX/manuscript.Rmd
manuscript updates
- new LaTeX template.tex for pdf output formatting
- updated figures
- updated tables
- new .csl file for referencing style (Methods Eco Evol)
- new .bib file to store references
- REMOVED manuscript-HTML, but HTML output still exists in new manuscript.Rmd

--- 2017-11-12 manuscript/manuscript-HTML.Rmd
@@
 title: "`respR` - An R package for the analysis of respirometry data"
-author: "Januar Harianto, Nicholas Carey, Maria Byrne"
+
+author:
+- name       : Januar Harianto
+  affiliation: University of Sydney, School of Medical Science, NSW Australia 2006
+- name       : Nicholas Carey
+  affiliation: Scottish Association for Marine Science, Oban, Argyll, UK, PA37 1QA
+- name       : Maria Byrne
+  affiliation: University of Sydney, School of Life & Environmental Sciences (SOLES), NSW Australia 2006
+
 output:
+  pdf_document:
+    fig_caption    : yes
+    keep_tex       : yes
+    number_sections: yes
+    latex_engine   : xelatex
+    template       : template.tex
   html_document:
-    toc: false
-    number_sections: true
-    fig_caption: true
-    theme: "default"
----
+    fig_caption    : yes
+    force_captions : yes
+    highlight      : pygments
+    number_sections: yes
+    theme          : default
 
-```{r, echo = F}
-library(respR)
-library(knitr)
-library(readr)
-Sys.setenv(TZ="Australia/Sydney")
-# set pander table-layout options
-library(pander)
-panderOptions('table.alignment.default', function(df)
-    ifelse(sapply(df, is.numeric), 'right', 'left'))
-panderOptions('table.split.table', Inf)
-panderOptions('big.mark', ",")
-panderOptions('keep.trailing.zeros', TRUE)
-
-opts_chunk$set(collapse = TRUE, comment = "#>", tidy = TRUE, highlight = TRUE)
-```
-
-# Abstract
-
-1.
+abstract: "1.
 Respirometry is a ubiquitous practice in experimental biology, but there is a lack of standard practices when analysing the resulting data, limiting transparency and reproducibility.
@@
 4.
-Although designed specifically with aquatic respirometry in mind, the object-oriented approach of the package and the unitless nature of its analytical functions mean that parts of the package can easily be used to estimate any other linear relationships in other research disciplines.
+Although designed specifically with aquatic respirometry in mind, the object-oriented approach of the package and the dimensionless nature of its analytical functions mean that parts of the package can easily be used to estimate any other linear relationships in other research disciplines."
+
+keywords: "R, software, respirometry, metabolic rate"
+thanks: "Replication files are available on the first author's Github account (http://github.com/januarharianto).
+**Current version**: `r format(Sys.time(), '%B %d, %Y')`; **Corresponding author**: januar.harianto@sydney.edu.au."
+
+fontfamily: mathpazo
+fontsize: 10pt
+geometry: margin = 1.6in
+csl: methods-in-ecology-and-evolution.csl
+bibliography: refs.bib
+---
+
+```{r, echo=FALSE}
+library(knitcitations); cleanbib()
+cite_options(citation_format = "pandoc", check.entries = TRUE)
+library(bibtex)
+library(respR)
+library(knitr)
+library(readr)
+library(readxl)
+library(kableExtra)
+Sys.setenv(TZ="Australia/Sydney")
+# set pander table-layout options
+library(pander)
+panderOptions('table.alignment.default', function(df)
+    ifelse(sapply(df, is.numeric), 'right', 'left'))
+panderOptions('table.split.table', Inf)
+panderOptions('big.mark', ",")
+panderOptions('keep.trailing.zeros', TRUE)
+
+listfunctions <- readxl::read_excel('../tables/list-of-functions.xlsx')
+conversions <- readxl::read_excel("../tables/conversion-units.xlsx")
+
+opts_chunk$set(collapse = TRUE, comment = "#>", tidy = TRUE, highlight = TRUE)
+```
 
 # Introduction
-
-Metabolic rate is a fundamental trait associated with virtually all biological functions and is key in predicting patterns in ecology and conservation biology, from populations (Seibel and Drazen, 2007, Barneche et al., 2014) to ecosystems (Brown et al., 2004).
-It has been widely investigated in studies of the effects of external stressors on organisms, and is increasingly used to study resilience in climate change-related studies where warming is expected to drive increased metabolism in ectotherms (e.g.
-Pörtner, 2002, McElroy et al., 2012, Carey et al., 2016, Delorme & Sewell, 2016).
-By far the most common metric used to assess physiological performance in animals, metabolic rate has been determined from whole organisms to the level of cells and tissues (White & Kearney, 2013).
-Metabolic rate is typically quantified using respirometry, which measures oxygen uptake over time (Lighton, 2008).
+Metabolic rate is a fundamental trait associated with virtually all biological functions and is key in predicting patterns in ecology and conservation biology, from populations [@seibel_rate_2007; @barneche_scaling_2014] to ecosystems [@brown_toward_2004].
+It has been widely investigated in studies of the effects of external stressors on organisms, and is increasingly used to study resilience in climate change-related studies where warming is expected to drive increased metabolism in ectotherms [e.g.
+@portner_climate_2002; @mcelroy_respiratory_2012; @carey_sea_2016; @delorme_effects_2016].
+By far the most common metric used to assess physiological performance in animals, metabolic rate has been determined from whole organisms to the level of cells and tissues [@white_determinants_2013].
+Metabolic rate is typically quantified using respirometry, which measures oxygen uptake over time [@lighton_measuring_2008].
 
@@
 Oxygen recordings may be continuous through use of an oxygen probe, periodic through withdrawing water or gas samples at set intervals, or a two-point measurement consisting of the initial and final concentrations.
-Metabolic rates are estimated from the $O_2$ timeseries by assuming a linear relationship between variables, and estimates of metabolic rate are straightforward in constant volume respirometry using the equation: $$VO_2 = \Delta O_2V$$ where $\Delta O_2V$ is the slope of the regression that describes the rate of change in $O_2$ concentration over time, or in the case of a two-point measurement, the difference in $O_2$ concentration divided by time elapsed, and $V$ is the volume of the container (Lighton, 2008).
-<!-- Note, we use here and in the package the term $RO_2$ to denote rate of $O_2$ depletion or uptake within an experiment.
-We use this term used to distinguish it from other common terms, for example $VO_2$ which is can be used to denote both volumetric O2 uptake and whole animal uptake, or $MO_2$ which typically denotes mass-specific uptake.
--->
+Metabolic rates are estimated from the $O_2$ timeseries by assuming a linear relationship between variables, and estimates of metabolic rate are straightforward in constant volume respirometry using the equation: $$VO_2 = \Delta O_2V$$ where $\Delta O_2$ is the slope of the regression that describes the rate of change in $O_2$ concentration over time, or in the case of a two-point measurement, the difference in $O_2$ concentration divided by time elapsed, and $V$ is the volume of the container [@lighton_measuring_2008].
 
@@
 This technique is essentially closed respirometry, but with the incorporation of repeated measures.
-Depending on the metabolic rate metric being investigated, final respiration rate can be calculated as the mean of the measures (e.g.
-Carey et al., 2016), or the lowest or highest rates recorded in any trial (e.g.
-Stoffels, 2015).
-<!--[nc-will try and find a better ref for that]
-
-Stoffels, R.
-J.
-(2015).
-Physiological trade-offs along a fast-slow lifestyle continuum in fishes: What do they tell us about resistance and resilience to hypoxia?
-Plos One, 10(6), e0130303.
-https://doi.org/10.1371/journal.pone.0130303
--->
+Depending on the metabolic rate metric being investigated, final respiration rate can be calculated as the mean of the measures [e.g.
+@carey_sea_2016], or the lowest or highest rates recorded in any trial [e.g.
+@stoffels_physiological_2015].
 
 Flow-through respirometry involves a closed chamber, but with a regulated flow of air or water through it at a precisely determined rate.
-After equilibrium has been achieved, the oxygen concentration differential between the incurrent and excurrent channels, along with the flow rate, allows calculation of the oxygen extracted from the flow volume per unit time: $$VO_2 = (C_iO_2 - C_eO_2)FR$$ where $VO_2$ is the rate of $O_2$ consumption over time, $C_iO_2$ and $C_eO_2$ are the incurrent and excurrent $O_2$ concentrations, and $FR$ is the flow rate of water through the system  (Lighton, 2008).
+After equilibrium has been achieved, the oxygen concentration differential between the incurrent and excurrent channels, along with the flow rate, allows calculation of the oxygen extracted from the flow volume per unit time: $$VO_2 = (C_iO_2 - C_eO_2)FR$$ where $VO_2$ is the rate of $O_2$ consumption over time, $C_iO_2$ and $C_eO_2$ are the incurrent and excurrent $O_2$ concentrations, and $FR$ is the flow rate of water through the system [@lighton_measuring_2008].
 
-A final method is open respirometry, in which an open tank or semi-enclosed area is used, but the input or mixing rate of oxygen from the surroundings is known or found to be negligible relative to oxygen consumption of the specimens (Leclercq et al., 1999).
-It is seldom used, but for some applications it is a sufficient and practical methodology (e.g.
-Gamble et al., 2014).
-The common equation used for open respirometry is: $$VO_2 = \Delta O_2V + \phi_d$$ where $\Delta O_2V$ is the slope of the regression that relates $O_2$ concentration to time, $V$ is the volume of the container and $\phi_d$ is the oxygen flux as determined by Fick's Law (Leclercq et al., 1999).
-
-<!--[nc - so the following needs to be spot on.
-Physiologists get very worked up (if not confused themselves) about the differences between these, esp.
-SMR v Routine v Resting v Basal!
-Especially those who work on ecto vs endo therms]-->
+A final method is open respirometry, in which an open tank or semi-enclosed area is used, but the input or mixing rate of oxygen from the surroundings is known or found to be negligible relative to oxygen consumption of the specimens [@leclercq_measurement_1999].
+It is seldom used, but for some applications it is a sufficient and practical methodology [@gamble_open-top_2014].
+The common equation used for open respirometry is: $$VO_2 = \Delta O_2V + \phi_d$$ where $\Delta O_2V$ is the slope of the regression that relates $O_2$ concentration to time, $V$ is the volume of the container and $\phi_d$ is the oxygen flux as determined by Fick's Law [@leclercq_measurement_1999].
 
 Depending on the experiment, different metabolic rate parameters may be of interest to researchers.
-These include maximum metabolic rates (MMR, or $MO_{2,max}$), rates under high activity or exhaustive exercise, and minimal metabolic rates ($MO_{2,min}$), which may be termed standard (SMR), basal (BMR), or resting metabolic rates, and typically represents the minimum metabolic cost of maintaining biological functioning (White and Kearney, 2014).
-Routine metabolic rate (RMR), is similar to $MO_{2,min}$, but takes into account that in some organisms energy is expended on small, spontaneous movements to maintain posture or position (Rogers et al.
-2016).
-Another metric of great interest is the critical oxygen tension, which represents the lowest level of oxygen at which aerobic metabolism is independent of the ambient partial pressure of oxygen ($P_{crit}$; Yeager & Ultsch, 1989, Hochachka and Somero, 2002).
+These include maximum metabolic rates (MMR, or $MO_{2,max}$), rates under high activity or exhaustive exercise, and minimal metabolic rates ($MO_{2,min}$), which may be termed standard (SMR), basal (BMR), or resting metabolic rates, and typically represents the minimum metabolic cost of maintaining biological functioning [@white_determinants_2013].
+Routine metabolic rate (RMR), is similar to $MO_{2,min}$, but takes into account that in some organisms energy is expended on small, spontaneous movements to maintain posture or position [@rogers_new_2016].
+Another metric of great interest is the critical oxygen tension, which represents the lowest level of oxygen at which aerobic metabolism is independent of the ambient partial pressure of oxygen [$P_{crit}$; @yeager_physiological_1989; @hochachka_biochemical_2002].
 
-<!--White, C.
-R., & Kearney, M.
-R.
-(2014).
-Metabolic scaling in animals: methods, empirical results, and theoretical explanations.
-Comprehensive Physiology, 4(1), 231–256.
-https://doi.org/10.1002/cphy.c110049-->
-
-Respirometry studies are increasingly outputting large, high-resolution datasets and conducted over longer periods (e.g.
-20 h, Norin & Malte, 2012).
+Respirometry studies are increasingly outputting large, high-resolution datasets and conducted over longer periods [e.g.
+20 h ; @norin_intraspecific_2012].
 In most cases, processing the data involves an *ad hoc* selection of data points with poorly reported criteria, and subsequent manual processing of the data subset(s) using a spreadsheet program (e.g.
@@
 
-<!-- r.e. below para. LoLinR is not strictly for respirometry - just monotonic, something-changing-over-time data -->
-
 A number of open-source R software packages have recently become available which are designed for, or are suitable to analyse respirometry data.
@@
 The `rMR` package (https://CRAN.R-project.org/package=rMR) performs interval-based metabolic rate calculations and has a method to automatically detect $P_{crit}$ using the "broken-stick" regression method (Yeager & Ultsch, 1989).
-`LoLinR` (Olito et al., 2017), while not specifically coded for respirometry, provides a statistically robust method of detecting a "best-fit" regression, and performs very well at identifying truly linear subsets of a data series.
+The package `LoLinR` [@olito_estimating_2017], while not specifically coded for respirometry, provides a statistically robust method of detecting a "best-fit" regression, and performs very well at identifying truly linear subsets of a data series.
 Importantly, all of these packages assume that data are monotonic with evenly-spaced sampling periods with respect to time, an ideal structure but one where respirometry data often fails to meet in reality.
@@
 
-<!-- FIGURE 1 -->
-Here we describe `respR`, an open-source R package designed to provide an efficient and reproducible workflow for the analysis of respirometry data.
+Here we describe `respR`, an open-source R package designed to provide an efficient and reproducible workflow for the analysis of respirometry data (Table 1).
 The package contains utilities to: (1) analyse closed, intermittent, flow-through and open respirometry data, (2) determine volume, mass, and surface area specific oxygen uptake rates, (3) automatically detect maximal and minimal rates, $P_{crit}$, and (4) detect the most linear sections of data using traditional rolling regression and kernel density estimation techniques.
@@
 
+\singlespacing
+
+```{r Table-1, echo=F, results='asis'}
+pander(listfunctions, caption = "Main functions available in `respR`.", 
+  split.cells = 50, split.table = 80, floating = FALSE)
+```
+
+\doublespacing
+
 # Package overview
 
-The `respR` package streamlines the analysis of respirometry data by partitioning the data processing workflow into several independent actions: (1) initially, data are imported, checked for common errors, plotted for visual inspection and exported as an object; (2) the object is then analysed to estimate the desired rate parameter; (3) diagnostic plots are automatically generated and the data can be further explored using the generic S3 `print()`, `plot()`, and `summary()` commands to verify the results; (4) adjustments to the data, either from background respiration and/or oxygen flux in open respirometry, are applied; (5) volume or mass-specific rate conversions, specific to the respiratory chamber and specimen(s), are calculated where necessary (**Fig.
-1**).
+The `respR` package streamlines the analysis of respirometry data by partitioning the data processing workflow into several independent actions: (1) initially, data are imported, checked for common errors, plotted for visual inspection and exported as an object; (2) the object is then analysed to estimate the desired rate parameter; (3) diagnostic plots are automatically generated and the data can be further explored using the generic S3 `print()`, `plot()`, and `summary()` commands to verify the results; (4) adjustments to the data, either from background respiration and/or oxygen flux in open respirometry, are applied; (5) volume or mass-specific rate conversions, specific to the respiratory chamber and specimen(s), are calculated where necessary (Fig.
+1).
 By design, each step of the workflow requires minimal manual manipulation of the data, and the user can pass the output of each function to the next.
@@
 
+\singlespacing
+
 ```{r eval=F}
@@
 
-To illustrate the main features of the package, 4 example datasets, `sardine.rd`, `urchin.rd`, `intermittent.rd` and `squid.rd` are distributed with `respR`.
+\doublespacing
+
+To illustrate the main features of the package, four example datasets, `sardine.rd`, `urchin.rd`, `intermittent.rd` and `squid.rd` are distributed with `respR`.
 Detailed information about the example data, including their source and methods, can be obtained with the `?` command in the R console (e.g.
 `?sardine.rd`).
+
+```{r Figure1, fig.cap="Diagram showing steps in a typical workflow in the analysis of respirometry data using `respR`.
+**(A)** Data are first checked for errors before **(B)** main functions are used to extract and analyse segments of the data.
+**(C)** Summarised results and diagnostic plots provide immediate visual feedback on the outcome of the analyses.
+**(D)** Once the rate estimates are obtained, they can be converted into volume and/or mass-specific rates.", echo=F}
+
+knitr::include_graphics("../figs/Figure 1.png")
+
+```
 
@@
 
-<!-- ...Identifying irregularities in the sampling time is therefore important as it as variations in data intervals may cause systemic bias in the analysis, especially if data loss or machine drift during sampling is persistent through time and time-based comparisons are made between experiments.
--->
-
 Using `inspect.data()` is optional -- the main functions in our package will readily work with data frames that were not processed by the function.
@@
 
+\singlespacing
+
 ```{r}
-inspect.data(squid, plot = F)
+inspect.data(squid.rd, plot = F)
 ```
+
+\doublespacing
 
@@
 
-Background oxygen consumption rate ($RO_({2(bg)}$) is often measured to account for the contribution of microbial respiration, and applied as a correction to experimental rates (Rodgers et al., 2016).
-Background rates typically account for a small percentage of experimental rates, and multiple “blank” experiments are usually conducted and the rates averaged to obtain a more accurate estimate of the correction (e.g.
-Carey and Sigwart, 2014; Daoud et al., 2007).
+Background oxygen consumption rate, $RO_{2(bg)}$, is often measured to account for the contribution of microbial respiration, and applied as a correction to experimental rates [@rogers_new_2016].
+Background rates typically account for a small percentage of experimental rates, and multiple “blank” experiments are usually conducted and the rates averaged to obtain a more accurate estimate of the correction [e.g.
+@carey_size_2014; @daoud_temperature_2007].
 The function `calc.rate.bg()` uses simple linear regression to simultaneously process multiple background rate measurements.
@@
 
-Typically, the analysis respirometry data requires the selection one or more linear sections of the raw data for more accurate calculations of $RO_2$ (e.g.
-Gordon et al., 1989, Dorey et al., 2014, Chu & Gal, 2017).
+Typically, the analysis respirometry data requires the selection one or more linear sections of the raw data for more accurate calculations of $RO_2$ [e.g.
+@gordon_energetics_1989; @chu_ecophysiological_2017].
 The function `calc.rate()` can manually extract and process data subset from respiromety data.
@@
 2a**).
-Thus, a total of $(m−n)+1$ number of overlapping regressions are fitted, which are then ranked to obtain maximum and minimum values.
+Thus, a total of $(m - n) + 1$ number of overlapping regressions are fitted, which are then ranked to obtain maximum and minimum values.
 If an interval-based regression is required, we select the `"interval"` method when calling the function, and it automatically selects non-overlapping sections of the data for regressions (**Fig.
 2b**).
+
+```{r Figure2, echo = F, fig.cap="Illustration showing the sampling method used by `auto.rate()` during **(A)** interval regression analysis, where sample windows do not overlap and **(B)** rolling regression analysis, where sample windows overlap and move forward by one sample unit at a time."}
+
+knitr::include_graphics("../figs/Figure 2.png")
+
+```
 
@@
 If the rate changes at some point during the sampling period, then the rolling estimate should detect the variability.
-This function takes advantage of this relationship and performs a Gaussian kernel density estimate (KDE) of all regressions (see Silverman, 1986).
+This function takes advantage of this relationship and performs a Gaussian kernel density estimate (KDE) of all regressions [see @silverman_density_1986].
 Peaks in the density thus reflect stable rates, and new segments of the data within the kernel bandwidth of the peak values are resampled and re-analysed to determine the most linear sections.
@@
 
+```{r Figure3, echo = F, fig.cap="Typical graphical outputs of the function `auto.rate()`.
+**(A)** Depending on the `method` input, the function is able to show: (clockwise form top left) the maximum rate detected, the minimum rate detected, calculation of rate by fixed-width intervals, and the most linear region as determined by kernel density estimates.
+**(B)** In addition, the outputs will also print diagnostic plots of: (clockwise from top left) the rolling regression of rate as a function of time with dotted lines indicating the region used to estimate the parameter, a kernel density plot that shows only when the `linear` method is called, a residual plot of the regression, and a normal Q-Q plot of the residuals of the regression.
+Analysis is performed using the example dataset `sardine`, which is included in the package.
+Code to reproduce the plots is available in the Supplementary Material."}
+
+knitr::include_graphics("../figs/Figure 3.png")
+
+```
+
 ### Estimating critical oxygen tension, $P_{crit}$
@@
 We provide two methods of estimating $P_{crit}$.
-The first is a "broken-stick" regression (BSR) approach, adopted from Yeager and Ultsch (1989) in which two segments of the data are iteratively fitted and the intersection with the smallest sum of the residual sum of squares between the two linear models is the estimated critical point.
-The second method is a wrapper for the segmented, or "broken-line" regression approach, presented by Muggeo (2003) and available as an R package (Muggeo 2008), which estimates the critical point by iteratively fitting two intersecting models on the data and selecting the point that minimises the "gap" between the two fitted lines.
+The first is a "broken-stick" regression (BSR) approach, adopted from @yeager_physiological_1989 in which two segments of the data are iteratively fitted and the intersection with the smallest sum of the residual sum of squares between the two linear models is the estimated critical point.
+The second method is a wrapper for the segmented, or "broken-line" regression approach, presented by @muggeo_estimating_2003 and available as part of the `segmented` R package [@muggeo_segmented:_2008], which estimates the critical point by iteratively fitting two intersecting models on the data and selecting the point that minimises the "gap" between the two fitted lines.
 
@@
 Alternatively, the user already has the required data, and the function may run the data directly using the argument `has.rate = TRUE`.
+
+```{r, echo = F, fig.cap="Output plots of the function `pcrit()`, used in the calculation of critical oxygen tension, $P_{crit}$.
+**(A)** To determine $P_{crit}$ in a regular $O_2$ timeseries, the function performs a rolling regression againt the rolling mean of the data.
+This produces **(B)** data that represents the rate of $O_2$ consumption as a function of $O_2$ concentration.
+Regression lines are then fitted iteratively, based on the broken-stick methods of Yeager and Ultsch (1989) and the segmented approach of Muggeo (2003), until convergence is achieved.
+The output then highlights the critical points as indicated by the coloured lines."}
+
+knitr::include_graphics("../figs/Figure 4.png")
+
+```
 
@@
 copepod eggs, Hammervold et al., 2015), or density (e.g.
-sea urchin larvae, Stumpp et al., 2011), which are currently unsupported.
+sea urchin larvae, Stumpp et al., 2011), all of which are currently not supported in `respR`.
 Users may also be interested in using respR's analytical functions to investigate other linear relationships or perform breakpoint analyses on non-respirometry-related data.
@@
 
+\singlespacing
+
+```{r Table2, echo = F, results='asis'}
+
+pander(conversions, caption = "Units available for conversions in `calc.rate()` and `scale.rate()`.")
+
+```
+
+\doublespacing
+
 # Concluding remarks and future improvements
@@
 
-# Tables & Figures
+# References
+\singlespacing
+\setlength{\parindent}{-0.2in}
+\setlength{\leftskip}{0.2in}
+\setlength{\parskip}{8pt}
+\vspace*{-0.2in}
+\noindent
 
-<!-- TAB 1 ----------------------------------------------------->
-**Table 1.** Main functions available in `respR`.
-
-```{r, echo=F, message=F}
-# use this later to show table
-library(readr)
-listfunctions <- readxl::read_excel('tables/list-of-functions.xlsx')
-conversions <- readxl::read_excel("tables/conversion-units.xlsx")
+```{r, warning=FALSE, message=FALSE, echo=FALSE} 
+  # write.bibtex(file="refs.bib")
 ```
-
-```{r, echo=F, results='asis'}
-pander(listfunctions)
-```
-
-<!-- TAB 2 ----------------------------------------------------->
-**Table 2.** Units available for conversions in `calc.rate()` and `scale.rate()`.
-
-```{r, echo = F, results='asis'}
-pander(conversions)
-
-```
-
-<!-- FIG 1 ----------------------------------------------------->
-
-```{r, echo = F, out.width = "800px"}
-knitr::include_graphics("figs/Figure 1.png")
-```
-
-**Fig.
-1.** Diagram showing steps in a typical workflow in the analysis of respirometry data using `respR`.
-Data are first checked for errors before the main functions are used to extract and analyse segments of the data.
-Summarised results and diagnostic plots provide immediate visual feedback on the outcome of the analyses.
-Once the rate estimates are obtained, they can be converted into volume and/or mass-specific rates.
-
-<!-- FIG 2 ----------------------------------------------------->
-
-```{r, echo = F, out.width = "650px"}
-knitr::include_graphics("figs/Figure 2.png")
-```
-
-**Fig.
-2** Illustration showing the sampling method used by `auto.rate()` during **(A)** interval regression analysis, where sample windows do not overlap and **(B)** rolling regression analysis, where sample windows overlap and move forward by one sample unit at a time.
-<!-- FIG 3 ----------------------------------------------------->
-
-```{r, echo = F, , out.width = "800px"}
-knitr::include_graphics("figs/Figure 3.png")
-```
-
-**Fig.
-3** Typical graphical outputs of the function `auto.rate()`.
-**(A)** Depending on the `method` input, the function is able to show: (clockwise form top left) the maximum rate detected, the minimum rate detected, calculation of rate by fixed-width intervals, and the most linear region as determined by kernel density estimates.
-**(B)** In addition, the outputs will also print diagnostic plots of: (clockwise from top left) the rolling regression of rate as a function of time with dotted lines indicating the region used to estimate the parameter, a kernel density plot that shows only when the `linear` method is called, a residual plot of the regression, and a normal Q-Q plot of the residuals of the regression.
-Analysis is performed using the example dataset `sardine`, which is included in the package.
-Code to reproduce the plots is available in the Supplementary Material.
-
-<!-- FIG 44 ----------------------------------------------------->
-
-```{r, echo = F, , out.width = "800px"}
-knitr::include_graphics("figs/Figure 4.png")
-```
-
-**Fig.
-4** Output plots of the function `pcrit()`, used in the calculation of critical oxygen tension, $P_{crit}$.
-**(A)** To determine $P_{crit}$ in a regular $O_2$ timeseries, the function performs a rolling regression againt the rolling mean of the data.
-This produces **(B)** data that represents the rate of $O_2$ consumption as a function of $O_2$ concentration.
-Regression lines are then fitted iteratively, based on the broken-stick methods of Yeager and Ultsch (1989) and the segmented approach of Muggeo (2003), until convergence is achieved.
-The output then highlights the critical points as indicated by the coloured lines.
-
-# Supplementary material
-```{r, eval = F}
-
-# Measuring intermittent data
-# Load data:
-data("intermittent") 
-int <- inspect.data(intermittent) # check for errors
-# Calculate rate in 3 locations:
-int.rate <- calc.rate(int, c(200,2300,4100), c(1800,3200,4600), by = 'time')
-plot(int.rate, rep = 1)
-
-# Estimating max, min, interval, most linear (best-fit) rates
-data("sardine")
-sar <- inspect.data(sardine) # check for errors
-sar.max <- auto.rate(sar, width = 1500, method = "max") # maximum rate
-sar.min <- auto.rate(sar, width = 1500, method = "min") # minimum rate
-
-sar.int <- auto.rate(sar, width = 1500, method = "interval") # interval sampling
-print(sar.int)
-summary(sar.int)
-plot(sar.int, pos = 3) # view the plots for the third interval of the data
-
-sar.lin <- auto.rate(sar, width = 1500, method = "linear") # most linear rate
-sar.lin # same as print(sdd.max)
-print(sar.lin, pos = 2) # view rank 2 results
-plot(sar.lin, pos = 2) # plot rank 2 results
-summary(sar.lin) # summary call
-
-# NOTE: using using sample number as roll width is possible and exponentially faster.
-# We can switch to `by = "row"` if time is evenly-spaced.
-sar.max.row <- auto.rate(sar, width = 1500, by = "row", method = "max") # maximum rate
-
-# Pcrit analysis
-data(squid)
-sqd <- inspect.data(squid, 2, 3)
-# Perfomr pcrit using defaut
-sqd.pcrit <- pcrit(sqd) # this may take a while as over 30k regressions are performed
-sqd.pcrit # same as print(sqd.pcrit)
-
-```

### 2017-11-13 manuscript/LaTeX/manuscript.Rmd
manuscript updates

@@
 thanks: "Replication files are available on the first author's Github account (http://github.com/januarharianto).
-**Current version**: `r format(Sys.time(), '%B %d, %Y')`; **Corresponding author**: januar.harianto@sydney.edu.au."
+**Current version**: `r format(Sys.time(), '%B %d, %Y')`; **Corresponding author**: nicholas.carey@sams.ac.uk."
 
@@
 
-In intermittent flow respirometry, $O_2$ concentration is measured as described above, but periodically the chamber is flushed with new water or air, returning it to initial conditions, resealed, and the experiment repeated (Svendsen et al., 2016).
+In intermittent flow respirometry, $O_2$ concentration is measured as described above, but periodically the chamber is flushed with new water or air, returning it to initial conditions, resealed, and the experiment repeated [@svendsen_design_2016].
 This technique is essentially closed respirometry, but with the incorporation of repeated measures.
@@
 **(B)** In addition, the outputs will also print diagnostic plots of: (clockwise from top left) the rolling regression of rate as a function of time with dotted lines indicating the region used to estimate the parameter, a kernel density plot that shows only when the `linear` method is called, a residual plot of the regression, and a normal Q-Q plot of the residuals of the regression.
-Analysis is performed using the example dataset `sardine`, which is included in the package.
+Analysis is performed using the example dataset `sardine.rd`, which is included in the package.
 Code to reproduce the plots is available in the Supplementary Material."}

### 2017-11-15 manuscript/LaTeX/manuscript-171115.Rmd
manuscript update

--- 2017-11-13 manuscript/LaTeX/manuscript.Rmd
@@
 - name       : Maria Byrne
-  affiliation: University of Sydney, School of Life & Environmental Sciences (SOLES), NSW Australia 2006
+  affiliation: University of Sydney, School of Medical Science and School of Life & Environmental Sciences (SOLES), NSW Australia 2006
 
@@
     latex_engine   : xelatex
+    highlight      : tango
     template       : template.tex
@@
     force_captions : yes
-    highlight      : pygments
+    highlight      : seashell
     number_sections: yes
@@
 
-keywords: "R, software, respirometry, metabolic rate"
+keywords: "R, software, respirometry, ecophysiology, metabolic rate"
 thanks: "Replication files are available on the first author's Github account (http://github.com/januarharianto).
@@
 Data should be formatted correctly before use in `respR`.
-The function `inspect.data()` extracts a two-column data frame from a multi-column dataset and performs error checks on a data frame of any size, while plotting the data for quick data visualisation.
+The function `inspect_data()` extracts a two-column data frame from a multi-column dataset and performs error checks on a data frame of any size, while plotting the data for quick data visualisation.
 All data should be numeric; respR at this time has limited support for date/time data (e.g.
@@
 
-Using `inspect.data()` is optional -- the main functions in our package will readily work with data frames that were not processed by the function.
-Running `inspect.data()` is an exploratory step that flags potential issues before analysis.
+Using `inspect_data()` is optional -- the main functions in our package will readily work with data frames that were not processed by the function.
+Running `inspect_data()` is an exploratory step that flags potential issues before analysis.
 Error check results are summarised in the R console and all tests are saved in the output object with the locations (i.e.
@@
 ```{r}
-inspect.data(squid.rd, plot = F)
+inspect_data(squid.rd, plot = F)
 ```
@@
 @carey_size_2014; @daoud_temperature_2007].
-The function `calc.rate.bg()` uses simple linear regression to simultaneously process multiple background rate measurements.
+The function `calc_rate_bg()` uses simple linear regression to simultaneously process multiple background rate measurements.
 Data segments can be truncated before analysis by time period or row numbers.
@@
 @gordon_energetics_1989; @chu_ecophysiological_2017].
-The function `calc.rate()` can manually extract and process data subset from respiromety data.
+The function `calc_rate()` can manually extract and process data subset from respiromety data.
 Data subsets can be selected by (1) time period, (2) row numbers, (3) $O_2$ decrease or (4) proportion, which should accommodate most, if not all, manual data selection requirements and allow for consistent reporting of methods and results.
@@
 
-For more complex respirometry parameters, the function `auto.rate()` uses novel rolling regression and kernel density estimate algorithms to detect patterns in respirometry data.
+For more complex respirometry parameters, the function `auto_rate()` uses novel rolling regression and kernel density estimate algorithms to detect patterns in respirometry data.
 It can calculate maximum or minimum rates, extract rates at non-overlapping intervals, and has an exploratory method to detect the most linear sections of the data.
@@
 
-```{r Figure2, echo = F, fig.cap="Illustration showing the sampling method used by `auto.rate()` during **(A)** interval regression analysis, where sample windows do not overlap and **(B)** rolling regression analysis, where sample windows overlap and move forward by one sample unit at a time."}
+```{r Figure2, echo = F, fig.cap="Illustration showing the sampling method used by `auto_rate()` during **(A)** interval regression analysis, where sample windows do not overlap and **(B)** rolling regression analysis, where sample windows overlap and move forward by one sample unit at a time."}
 
@@
 
-The function `calc.rate.ft()` is used to specifically analyse flow-through respirometry data, and is similar to `calc.rate()` in functionality.
-However it accepts additional inputs of incurrent and excurrent oxygen concentration, and flow rate.`calc.rate.ft()` can process data to determine the mean or modal $RO_2$ over a period.
+The function `calc_rate_ft()` is used to specifically analyse flow-through respirometry data, and is similar to `calc_rate()` in functionality.
+However it accepts additional inputs of incurrent and excurrent oxygen concentration, and flow rate.`calc_rate_ft()` can process data to determine the mean or modal $RO_2$ over a period.
 
-```{r Figure3, echo = F, fig.cap="Typical graphical outputs of the function `auto.rate()`.
+```{r Figure3, echo = F, fig.cap="Typical graphical outputs of the function `auto_rate()`.
 **(A)** Depending on the `method` input, the function is able to show: (clockwise form top left) the maximum rate detected, the minimum rate detected, calculation of rate by fixed-width intervals, and the most linear region as determined by kernel density estimates.
@@
 
-As $P_{crit}$ is the break-point in the relationship between ambient dissolved oxygen ($DO_2$) and $RO_2$/$MO_2$, the user has two options to provide the required data.
-Loading an $O_2$ timeseries data into the function will prompt it to automatically perform a rolling mean and a rolling regression of $DO_2$ to extract those variables (**Fig.
-4**).
-Alternatively, the user already has the required data, and the function may run the data directly using the argument `has.rate = TRUE`.
+As $P_{crit}$ is the break-point in the relationship between ambient dissolved oxygen ($DO_2$) and rate of O2 uptake (e.g.
+$RO_2$), the user has two options to input data.
+With an $O_2 \sim$ timeseries dataset the function will automatically perform a rolling mean $RO_2$ and a rolling regression of $DO_2$ (Fig.
+4) in order to calculate $P_{crit}$ Alternatively, the user can input already calculated $RO_2/MO_2 \sim O_2$ data and the function will determine $P_{crit}$ using these data directly.
 
@@
 For deliberate reasons, units of measurement are not used when determining $RO_2$ in `respR`.
-Instead, two functions, `convert.rate()` and `scale.rate()`, are used to transform $RO_2$ to specific units and perform volumetric or mass-specific conversions .
+Instead, two functions, `convert_DO()` and `convert_rate()`, are used to transform $RO_2$ to specific units and perform volumetric or mass-specific conversions .
 This unique approach has the advantage of allowing "dimensionless" rate calculations for data that may not be fully supported by the entire respR workflow.
-For example, measurements of respiration can be scaled to surface area (e.g.
-diatoms, Iversen & Ploug, 2013, Naumann et al., 2014), volume (e.g.
-copepod eggs, Hammervold et al., 2015), or density (e.g.
-sea urchin larvae, Stumpp et al., 2011), all of which are currently not supported in `respR`.
+For example, measurements of respiration can be scaled to surface area [e.g.
+diatoms & coral, @iversen_temperature_2013; @naumann_species-specific_2014], volume [e.g.
+copepod eggs, @hammervold_new_2015], or density [e.g.
+sea urchin larvae, @stumpp_co2_2011], all of which are currently not supported in `respR`.
 Users may also be interested in using respR's analytical functions to investigate other linear relationships or perform breakpoint analyses on non-respirometry-related data.
 
-The function `convert.rate()` is used to convert units of oxygen concentration only (e.g.
+The function `convert_DO()` is used to convert units of oxygen concentration only (e.g.
 $mg/L$ or %) and has support for various unit combinations (Table 2).
@@
 "$mg/L$", "$mg/l$", "$mg L^{-1}$" and "$mg l^{-1}$" are identified as the same units).
-The function `scale.rate()` is conceptually similar to `convert.rate()`, but it focuses volumetric and mass-specific conversions (e.g.
+The function `convert_rate()` is conceptually similar to `convert_DO()`, but it focuses volumetric and mass-specific conversions (e.g.
 $mg h^{-1} kg^{-1}$).
@@
 
-pander(conversions, caption = "Units available for conversions in `calc.rate()` and `scale.rate()`.")
+pander(conversions, caption = "Units available for conversions in `convert_DO()` and `convert_rate()`.")
 

### 2017-11-17 manuscript/LaTeX/manuscript.Rmd
manuscript updates

@@
 4.
-Although designed specifically with aquatic respirometry in mind, the object-oriented approach of the package and the dimensionless nature of its analytical functions mean that parts of the package can easily be used to estimate any other linear relationships in other research disciplines."
+Although designed specifically with aquatic respirometry in mind, the object-oriented approach of the package and the dimensionless nature of its analytical functions mean that parts of the package can easily be used to estimate linear relationships from a range of applications in many research disciplines."
 
-keywords: "R, software, respirometry, metabolic rate"
+keywords: "R, software, respirometry, ecophysiology, metabolic rate"
 thanks: "Replication files are available on the first author's Github account (http://github.com/januarharianto).
@@
 
-opts_chunk$set(collapse = TRUE, comment = "#>", tidy = TRUE, highlight = TRUE)
+opts_chunk$set(collapse = TRUE, comment = "#>", tidy = TRUE, cache = TRUE, 
+  highlight = TRUE)
 ```
@@
 It has been widely investigated in studies of the effects of external stressors on organisms, and is increasingly used to study resilience in climate change-related studies where warming is expected to drive increased metabolism in ectotherms [e.g.
-@portner_climate_2002; @mcelroy_respiratory_2012; @carey_sea_2016; @delorme_effects_2016].
+@portner_climate_2002; @carey_sea_2016; @delorme_effects_2016].
 By far the most common metric used to assess physiological performance in animals, metabolic rate has been determined from whole organisms to the level of cells and tissues [@white_determinants_2013].
-Metabolic rate is typically quantified using respirometry, which measures oxygen uptake over time [@lighton_measuring_2008].
+Metabolic rate is typically quantified using respirometry, which measures the rate of oxygen uptake over time, $RO_2$ [@lighton_measuring_2008].
 
-There are four broad methodological approaches in respirometry: closed, intermittent flow, flow-through and open.
+<!-- There are four broad methodological approaches in respirometry: closed, intermittent flow, flow-through and open.
 In closed respirometry, $O_2$ decrease is measured within a hermetically sealed chamber of known volume, sometimes set within a closed loop to allow mixing of the environment within the chamber.
@@
 Metabolic rates are estimated from the $O_2$ timeseries by assuming a linear relationship between variables, and estimates of metabolic rate are straightforward in constant volume respirometry using the equation: $$VO_2 = \Delta O_2V$$ where $\Delta O_2$ is the slope of the regression that describes the rate of change in $O_2$ concentration over time, or in the case of a two-point measurement, the difference in $O_2$ concentration divided by time elapsed, and $V$ is the volume of the container [@lighton_measuring_2008].
+-->
 
-In intermittent flow respirometry, $O_2$ concentration is measured as described above, but periodically the chamber is flushed with new water or air, returning it to initial conditions, resealed, and the experiment repeated [@svendsen_design_2016].
+<!-- In intermittent flow respirometry, $O_2$ concentration is measured as described above, but periodically the chamber is flushed with new water or air, returning it to initial conditions, resealed, and the experiment repeated [@svendsen_design_2016].
 This technique is essentially closed respirometry, but with the incorporation of repeated measures.
@@
 @stoffels_physiological_2015].
+-->
 
-Flow-through respirometry involves a closed chamber, but with a regulated flow of air or water through it at a precisely determined rate.
+<!-- Flow-through respirometry involves a closed chamber, but with a regulated flow of air or water through it at a precisely determined rate.
 After equilibrium has been achieved, the oxygen concentration differential between the incurrent and excurrent channels, along with the flow rate, allows calculation of the oxygen extracted from the flow volume per unit time: $$VO_2 = (C_iO_2 - C_eO_2)FR$$ where $VO_2$ is the rate of $O_2$ consumption over time, $C_iO_2$ and $C_eO_2$ are the incurrent and excurrent $O_2$ concentrations, and $FR$ is the flow rate of water through the system [@lighton_measuring_2008].
+-->
 
-A final method is open respirometry, in which an open tank or semi-enclosed area is used, but the input or mixing rate of oxygen from the surroundings is known or found to be negligible relative to oxygen consumption of the specimens [@leclercq_measurement_1999].
+<!-- A final method is open respirometry, in which an open tank or semi-enclosed area is used, but the input or mixing rate of oxygen from the surroundings is known or found to be negligible relative to oxygen consumption of the specimens [@leclercq_measurement_1999].
 It is seldom used, but for some applications it is a sufficient and practical methodology [@gamble_open-top_2014].
 The common equation used for open respirometry is: $$VO_2 = \Delta O_2V + \phi_d$$ where $\Delta O_2V$ is the slope of the regression that relates $O_2$ concentration to time, $V$ is the volume of the container and $\phi_d$ is the oxygen flux as determined by Fick's Law [@leclercq_measurement_1999].
+-->
 
@@
 
-Respirometry studies are increasingly outputting large, high-resolution datasets and conducted over longer periods [e.g.
+Respirometry studies are increasingly outputting large, high-resolution datasets  conducted over longer periods [e.g.
 20 h ; @norin_intraspecific_2012].
@@
 A number of open-source R software packages have recently become available which are designed for, or are suitable to analyse respirometry data.
-The `respirometry` package (https://CRAN.R-project.org/package=respirometry) contains a comprehensive collection of tools to explore and evaluate experimental parameters in aquatic respirometry and is useful for the design and diagnosis of experimental setups.
-The `rMR` package (https://CRAN.R-project.org/package=rMR) performs interval-based metabolic rate calculations and has a method to automatically detect $P_{crit}$ using the "broken-stick" regression method (Yeager & Ultsch, 1989).
-The package `LoLinR` [@olito_estimating_2017], while not specifically coded for respirometry, provides a statistically robust method of detecting a "best-fit" regression, and performs very well at identifying truly linear subsets of a data series.
-Importantly, all of these packages assume that data are monotonic with evenly-spaced sampling periods with respect to time, an ideal structure but one where respirometry data often fails to meet in reality.
+The "`respirometry`"" package (https://CRAN.R-project.org/package=respirometry) contains a comprehensive collection of tools to explore and evaluate experimental parameters in aquatic respirometry and is useful for the design and diagnosis of experimental setups.
+The "`rMR`"" package (https://CRAN.R-project.org/package=rMR) performs interval-based metabolic rate calculations and has a method to automatically detect $P_{crit}$ using the "broken-stick" regression method (Yeager & Ultsch, 1989).
+The package "`LoLinR`"" [@olito_estimating_2017], while not specifically coded for respirometry, provides a statistically robust method of detecting a "best-fit" regression, and performs well at identifying truly linear subsets of a data series.
+Importantly, all of these packages assume that data are monotonic with evenly-spaced sampling periods with respect to time, an ideal structure but one where respirometry data often fails to meet.
 They also do not have the tools to analyse complicated datasets to automatically determine useful metric such as $MO_{2,max}$ and $MO_{2,min}$.
@@
 Here we describe `respR`, an open-source R package designed to provide an efficient and reproducible workflow for the analysis of respirometry data (Table 1).
-The package contains utilities to: (1) analyse closed, intermittent, flow-through and open respirometry data, (2) determine volume, mass, and surface area specific oxygen uptake rates, (3) automatically detect maximal and minimal rates, $P_{crit}$, and (4) detect the most linear sections of data using traditional rolling regression and kernel density estimation techniques.
+The package contains utilities to: (1) analyse closed, intermittent, flow-through and open respirometry data [see @lighton_measuring_2008; @svendsen_design_2016], (2) determine volume, mass, and surface area specific oxygen uptake rates, (3) automatically detect maximal and minimal rates, $P_{crit}$, and (4) detect the most linear sections of data using traditional rolling regression and kernel density estimation techniques.
 The package facilitates the output as supplementary material of all stages of respirometry analyses; raw data, subsetting criteria, background corrections, units of measure, and resulting metabolic rates.
@@
 By design, each step of the workflow requires minimal manual manipulation of the data, and the user can pass the output of each function to the next.
-This object-oriented approach makes `respR` easy to use, even for novice users of R, and allows us (or anyone interested in the open-source project) to easily extend the functionality of the package, either by adding additional intermediate actions or substituting current actions with other code without breaking the rest of the package.
+This object-oriented approach makes `respR` easy to use, even for novice users of R, and allows any interested R programer to easily extend the functionality of the package, either by adding additional intermediate actions or substituting current actions with other code without breaking the rest of the package.
 Note, we use $RO_2$ in the following text to represent the dimensionless rate of $O_2$ use with regard to the time unit, and to distinguish it from other commonly reported metrics (e.g.
@@
 Data should be formatted correctly before use in `respR`.
-The function `inspect.data()` extracts a two-column data frame from a multi-column dataset and performs error checks on a data frame of any size, while plotting the data for quick data visualisation.
-All data should be numeric; respR at this time has limited support for date/time data (e.g.
+The function `inspect_data()` extracts a two-column data frame from a multi-column dataset and performs simple error checks while plotting the data for quick data visualisation.
+All data should be numeric; `respR` at this time has limited support for date/time data (e.g.
 `POSIXct` and `POSIXt` classes).
@@
 @carey_size_2014; @daoud_temperature_2007].
-The function `calc.rate.bg()` uses simple linear regression to simultaneously process multiple background rate measurements.
+The function `calc_rate.bg()` uses simple linear regression to simultaneously process multiple background rate measurements.
 Data segments can be truncated before analysis by time period or row numbers.
@@
 
-Typically, the analysis respirometry data requires the selection one or more linear sections of the raw data for more accurate calculations of $RO_2$ [e.g.
-@gordon_energetics_1989; @chu_ecophysiological_2017].
-The function `calc.rate()` can manually extract and process data subset from respiromety data.
+Typically, the analysis of respirometry data requires the selection one or more linear sections of the data for more accurate calculations of $RO_2$ [e.g.
+@gordon_energetics_1989; @chu_ecophysiological_2017; @carey_sea_2016].
+The function `calc_rate()` can manually extract and process data subset from respiromety data.
 Data subsets can be selected by (1) time period, (2) row numbers, (3) $O_2$ decrease or (4) proportion, which should accommodate most, if not all, manual data selection requirements and allow for consistent reporting of methods and results.
 Rates are determined using linear regression, however a two-point analysis is also included alongside all regression results.
+Generally, $RO_2$ is calculated using the equation: $$RO_2 = \Delta O_2V$$ where $\Delta O_2$ is the slope of the regression that describes the rate of change in $O_2$ concentration over time, or in the case of a two-point measurement, the difference in $O_2$ concentration divided by time elapsed, and $V$ is the volume of the container [@lighton_measuring_2008].
 
-For more complex respirometry parameters, the function `auto.rate()` uses novel rolling regression and kernel density estimate algorithms to detect patterns in respirometry data.
+### Estimating maximum, minimum, most linear and interval $RO_2$
+
+For more complex respirometry parameters, the function `auto_rate()` uses novel rolling regression and kernel density estimate algorithms to detect patterns in respirometry data.
 It can calculate maximum or minimum rates, extract rates at non-overlapping intervals, and has an exploratory method to detect the most linear sections of the data.
 When called, the function takes the data frame of length $m$, and performs regressions of sample size $n$ iteratively across the length of the data (i.e.
-rolling regressions), one sample at a time (**Fig.
-2a**).
+rolling regressions), one sample at a time (Fig.
+2a).
 Thus, a total of $(m - n) + 1$ number of overlapping regressions are fitted, which are then ranked to obtain maximum and minimum values.
-If an interval-based regression is required, we select the `"interval"` method when calling the function, and it automatically selects non-overlapping sections of the data for regressions (**Fig.
-2b**).
+If an interval-based regression is required, we select the `"interval"` method when calling the function, and it automatically selects non-overlapping sections of the data for regressions (Fig.
+2b).
 
-```{r Figure2, echo = F, fig.cap="Illustration showing the sampling method used by `auto.rate()` during **(A)** interval regression analysis, where sample windows do not overlap and **(B)** rolling regression analysis, where sample windows overlap and move forward by one sample unit at a time."}
+```{r Figure2, echo = F, fig.cap="Illustration showing the sampling method used by `auto_rate()` during **(A)** interval regression analysis, where sample windows do not overlap and **(B)** rolling regression analysis, where sample windows overlap and move forward by one sample unit at a time."}
 
@@
 Peaks in the density thus reflect stable rates, and new segments of the data within the kernel bandwidth of the peak values are resampled and re-analysed to determine the most linear sections.
-Summary diagnostics and visual plots provide supporting information on the technique and the validity of the results (**Fig.
-3**).
+Summary diagnostics and visual plots provide supporting information on the technique and the validity of the results (Fig.
+3).
 
-The function `calc.rate.ft()` is used to specifically analyse flow-through respirometry data, and is similar to `calc.rate()` in functionality.
-However it accepts additional inputs of incurrent and excurrent oxygen concentration, and flow rate.`calc.rate.ft()` can process data to determine the mean or modal $RO_2$ over a period.
+### Calcularing $RO_2$ from flow-through respirometry
 
-```{r Figure3, echo = F, fig.cap="Typical graphical outputs of the function `auto.rate()`.
+The function `calc_rate.ft()` is used to specifically analyse flow-through respirometry data, and is similar to `calc_rate()` in functionality.
+However it accepts additional inputs of incurrent and excurrent oxygen concentration, and flow rate, based on the equation: $$VO_2 = (C_iO_2 - C_eO_2)FR$$ where $VO_2$ is the rate of $O_2$ consumption over time, $C_iO_2$ and $C_eO_2$ are the incurrent and excurrent $O_2$ concentrations, and $FR$ is the flow rate of water through the system [@lighton_measuring_2008].
+
+```{r Figure3, echo = F, fig.cap="Typical graphical outputs of the function `auto_rate()`.
 **(A)** Depending on the `method` input, the function is able to show: (clockwise form top left) the maximum rate detected, the minimum rate detected, calculation of rate by fixed-width intervals, and the most linear region as determined by kernel density estimates.
@@
 The function `adjust.rate()` can be used to correct $RO_2$ estimates for oxygen flux and/or background respirometry data.
-As the oxygen consumption units are not mass specific at the point of correction, the function adjusts RO_2 directly without bias using the equation: $$RO_{2(correction)} = RO_2 - (\phi_d + RO_{2(bg)})$$ where $\phi_d$ is the oxygen flux and $RO_{2(bg)}$ is background respiration.
+As the oxygen consumption units are not mass specific at the point of correction, the function adjusts RO_2 directly without bias using the equation: $$RO_{2(correction)} = RO_2 - (\phi_d + RO_{2(bg)})$$ where $\phi_d$ is the oxygen flux as determined by Fick's Law [@leclercq_measurement_1999] and $RO_{2(bg)}$ is background respiration.
 
@@
 
-For deliberate reasons, units of measurement are not used when determining $RO_2$ in `respR`.
-Instead, two functions, `convert.rate()` and `scale.rate()`, are used to transform $RO_2$ to specific units and perform volumetric or mass-specific conversions .
+Units of measurement are not used when determining $RO_2$ in `respR`.
+Instead, two functions, `convert_DO()` and `convert_rate()`, are used to transform $RO_2$ to specific units and perform volumetric or mass-specific conversions .
 This unique approach has the advantage of allowing "dimensionless" rate calculations for data that may not be fully supported by the entire respR workflow.
-For example, measurements of respiration can be scaled to surface area (e.g.
-diatoms, Iversen & Ploug, 2013, Naumann et al., 2014), volume (e.g.
-copepod eggs, Hammervold et al., 2015), or density (e.g.
-sea urchin larvae, Stumpp et al., 2011), all of which are currently not supported in `respR`.
+For example, measurements of respiration can be scaled to surface area [e.g.
+diatoms & coral, @iversen_temperature_2013; @naumann_species-specific_2014], volume [e.g.
+copepod eggs, @hammervold_new_2015], or density [e.g.
+sea urchin larvae, @stumpp_co2_2011], all of which are currently not supported in `respR`.
 Users may also be interested in using respR's analytical functions to investigate other linear relationships or perform breakpoint analyses on non-respirometry-related data.
 
-The function `convert.rate()` is used to convert units of oxygen concentration only (e.g.
+The function `convert_DO()` is used to convert units of oxygen concentration only (e.g.
 $mg/L$ or %) and has support for various unit combinations (Table 2).
 For ease of use, we have implemented a simple fuzzy string-matching algorithm to match relevant units even when they have been provided in different formats (e.g.
-"$mg/L$", "$mg/l$", "$mg L^{-1}$" and "$mg l^{-1}$" are identified as the same units).
-The function `scale.rate()` is conceptually similar to `convert.rate()`, but it focuses volumetric and mass-specific conversions (e.g.
-$mg h^{-1} kg^{-1}$).
+"$mg/L$", "$mg/l$", "$mg\ L^{-1}$" and "$mg\ l^{-1}$" are identified as the same units).
+The function `convert_rate()` is conceptually similar to `convert_DO()`, but it focuses volumetric and mass-specific conversions (e.g.
+$mg\ h^{-1}\ kg^{-1}$).
 
@@
 
+# Acknowledgements
+This work was supported by an NSW Environmental Trust grant (MB), and Endeavour Postdoctoral Fellowship (NC) and a Professor NGW & Mrs Ann Macintosh Memorial Scholarship from the University of Sydney (JH) and is Sydney Institue of Marine Science contribution number XXXX (to be confirmed).
+
+# Author contributions
+JH, NC and MB conceived the ideas and designed methodology; JH and NC wrote the code and analysed the example data; JH, NC and MB led the writing of the manuscript.
+All authors contributed critically to the drafts and gave final approval for publication.
+
 # References

### 2017-11-17 R/auto_rate.R
Revert "Merge branch 'develop'"
This reverts commit e4280a5feb04f74b9a099c84410fc510596ca50d, reversing
changes made to c2a5ed8e39c83fab98347ac4e97a434dd95f97b8.

@@ -7,12 +7,8 @@
 #' possible values. The computations are then ranked (or, arranged), based on
 #' the "`logic`" argument, and the output is summarised.
 #'
-#' **Units** There are no units of measurements involved in `auto_rate`. This is
-#' a deliberate decision. Units are called in a later function when volume-
-#' and/or weight-specific rates of oxygen concentration are computed in
-#' [convert_rate()] and [convert_DO()].
+#' **Units**
 #'
-<<<<<<< HEAD:R/auto_rate.R
 #' There are no units of measurements involved in `auto_rate`. This is a
 #' deliberate decision. Units are called in a later function when volume- and/or
 #' weight-specific rates of oxygen concentration are computed in
@@ -23,44 +19,31 @@
 #'
 #' For now, `auto_rate()` contains four ranking algorithms that can be called
 #' with the argument "`logic`":
-=======
-#'
-#' ***Ranking algorithms*** For now, `auto_rate()` contains four ranking
-#' algorithms that can be called with the argument "`logic`":
->>>>>>> develop:R/auto_rate.R
 #'
 #' - `max`: regressions are arranged from highest absolute values, to the
 #' lowest.
-#'
 #' - `min`: regressions are arranged from lowest absolute values, to the
 #' highest.
-#'
-#' - `interval`: non-overlapping regressions are extracted from the
-#' rolled regrssions. They are not ranked.
-#'
-#' - `linear`: Buses kernel density
-#' estimation to detect the most "linear" sections of the timeseries in
-#' descending order.
+#' - `interval`: non-overlapping regressions are extracted from the rolled
+#' regrssions. They are not ranked.
+#' - `linear`: Buses kernel density estimation to detect the most "linear"
+#' sections of the timeseries in descending order.
 #'
 #' @param df data frame.
 #' @param width numeric. Defaults to 25 percent of width if NULL.
 #' @param by string. "row" or "time". Defaults to "row".
 #' @param method string. "linear", "max", "min" or "interval".
-#' @param plot logical. Defaults to TRUE. Automatically plot the results.
 #'
[…]

### 2017-11-17 R/calc_rate.R

@@ -17,10 +17,6 @@
 #' @param plot logical. Defaults to TRUE.
 #'
 #' @importFrom data.table data.table rbindlist
-#' @import utils
-#' @import stats
-#' @import graphics
-#' @import grDevices
 #'
 #' @return A list object of class `calc_rate`.
 #' @export
@@ -61,7 +57,7 @@ calc_rate <- function(x, from = NULL, to = NULL, by = "time", plot = T) {
   }
 
   # Subset the data:
-  dt <- lapply(1:length(from), function(z) subset_data(x, from[z], to[z], by))
+  dt <- lapply(1:length(from), function(z) subset.data(x, from[z], to[z], by))
 
   # Perform lm on data and extract coefficients
   coefs <- lapply(1:length(to), function(z) linear_fit(dt[[z]]))
@@ -102,33 +98,20 @@ calc_rate <- function(x, from = NULL, to = NULL, by = "time", plot = T) {
 
 
 #' @export
-<<<<<<< HEAD:R/calc_rate.R
 print.calc_rate <- function(x) {
-=======
-print.calc_rate <- function(x, ...) {
->>>>>>> develop:R/calc_rate.R
   cat("Rate(s):\n")
   print(x$rate)
 }
 
 
 #' @export
-<<<<<<< HEAD:R/calc_rate.R
 summary.calc_rate <- function(x) {
-=======
-summary.calc_rate <- function(x, ...) {
->>>>>>> develop:R/calc_rate.R
   cat("Summary:\n")
   print(x$summary)
-  return(invisible(x$summary))
 }
 
 #' @export
-<<<<<<< HEAD:R/calc_rate.R
 plot.calc_rate <- function(x, rep = 1) {
-=======
-plot.calc_rate <- function(x, rep = 1, ...) {
->>>>>>> develop:R/calc_rate.R
   message('Plotting...this may take a while for large datasets.')
   df  <- x$data
   sdf <- x$subsets[[rep]]

### 2017-11-17 R/respR.R

@@ -8,5 +8,5 @@
 #' for background corrections and unit conversions are included.
 #'
 #' @author Januar Harianto & Nicholas Carey
-#' @docType package
-#' @name respR
+#'
+"_PACKAGE"

### 2017-11-17 R/subsample.R

@@ -8,7 +8,6 @@
 #' @param df data frame.
 #' @param n numeric.
 #' @param random_start logical.
-#' @param plot logical. Defaults to TRUE. Plots the data automatically.
 #'
 #' @return A data frame object.
 #' @export
@@ -19,20 +18,16 @@
 #'
 #' # Subsample with random first value:
 #' subsample(sardine.rd, 3, T)
-<<<<<<< HEAD
 subsample <- function(df, n = 5, random_start = F) {
-=======
-subsample <- function(df, n = 5, random_start = F, plot = T) {
->>>>>>> develop
   # First check if [random_start] is true. If true, the
   # function will randomise the value taken from row 1:[n] and
   # insert it into the [from] argument in [seq].
-  if (random_start) {
+  if (random_start == T) {
     start <- sample(1:n, 1)
 
     # If set to false, the function will sample the first row for
     # the [from] argument in [seq].
-  } else {
+  } else if (random_start == F) {
     start <- 1
   }
   # Sample every [n] rows from dataframe, starting from [rs],
@@ -40,26 +35,22 @@ subsample <- function(df, n = 5, random_start = F, plot = T) {
   # performance.
   subset <- df[seq.int(start, nrow(df), n), ]
   # Create the plots
-
-  if (plot) {
-
-    pardefault <- par(no.readonly = T)  # save original par settings
-    par(mfrow=c(1,2))  # replace par settings
-    plot(df, xlab = "Time", ylab = "Oxygen", col = r1, pch = 16,
-      panel.first = c(rect(par("usr")[1],
-        par("usr")[3],
-        par("usr")[2],
-        par("usr")[4], col = r3), grid(col = "white",
-          lty = 1, lwd = 1.5)))
-    title(main = "Full Timeseries", line = 0.5)
-    plot(subset, xlab = "Time", ylab = "Oxygen", col = r1, pch = 16,
-      panel.first = c(rect(par("usr")[1],
-        par("usr")[3],
-        par("usr")[2],
-        par("usr")[4], col = r3), grid(col = "white",
-          lty = 1, lwd = 1.5)))
-    title(main = "Subsample Timeseries", line = 0.5)
-    par(pardefault)  # revert par settings to original
-  }
+  pardefault <- par(no.readonly = T)  # save original par settings
+  par(mfrow=c(1,2))  # replace par settings
[…]

### 2017-11-17 R/unit_args.R

@@ -1,4 +1,3 @@
-<<<<<<< HEAD:R/unit_args.R
 #' Print examples of unit arguments for use in [convert_rate()] and
 #' [scale_rate()]
 #'
@@ -6,24 +5,11 @@
 #' that can be used in the functions [convert_rate()] and [scale_rate()].
 #'
 #' Available units for [convert_rate()]:
-=======
-#' Print examples of unit arguments for use in [convert_DO()] and
-#' [convert_rate()]
-#'
-#' This is a simple function, requiring no argument. It shows available units
-#' that can be used in the functions [convert_DO()] and [convert_rate()].
-#'
-#' Available units for [convert_DO()]:
->>>>>>> develop:R/unit_args.R
 #'
 #' "mg/L", "ug/L", "mmol/L", "umol/L", "mL/L", "mg/kg", "ug/kg", "mmol/kg",
 #' "umol/kg", "mL/kg", "%", "Torr", "hPa", "kPa", "mmHg", "inHg"
 #'
-<<<<<<< HEAD:R/unit_args.R
 #' Available units for [scale_rate()]:
-=======
-#' Available units for [convert_rate()]:
->>>>>>> develop:R/unit_args.R
 #'
 #' **O2**
 #'

### 2017-11-17 manuscript/LaTeX/manuscript.Rmd

@@
 4.
-Although designed specifically with aquatic respirometry in mind, the object-oriented approach of the package and the dimensionless nature of its analytical functions mean that parts of the package can easily be used to estimate linear relationships from a range of applications in many research disciplines."
+Although designed specifically with aquatic respirometry in mind, the object-oriented approach of the package and the dimensionless nature of its analytical functions mean that parts of the package can easily be used to estimate any other linear relationships in other research disciplines."
 
@@
 
-opts_chunk$set(collapse = TRUE, comment = "#>", tidy = TRUE, cache = TRUE, 
-  highlight = TRUE)
+opts_chunk$set(collapse = TRUE, comment = "#>", tidy = TRUE, highlight = TRUE)
 ```
@@
 It has been widely investigated in studies of the effects of external stressors on organisms, and is increasingly used to study resilience in climate change-related studies where warming is expected to drive increased metabolism in ectotherms [e.g.
-@portner_climate_2002; @carey_sea_2016; @delorme_effects_2016].
+@portner_climate_2002; @mcelroy_respiratory_2012; @carey_sea_2016; @delorme_effects_2016].
 By far the most common metric used to assess physiological performance in animals, metabolic rate has been determined from whole organisms to the level of cells and tissues [@white_determinants_2013].
@@
 
-Respirometry studies are increasingly outputting large, high-resolution datasets  conducted over longer periods [e.g.
+Respirometry studies are increasingly outputting large, high-resolution datasets and conducted over longer periods [e.g.
 20 h ; @norin_intraspecific_2012].
@@
 A number of open-source R software packages have recently become available which are designed for, or are suitable to analyse respirometry data.
-The "`respirometry`"" package (https://CRAN.R-project.org/package=respirometry) contains a comprehensive collection of tools to explore and evaluate experimental parameters in aquatic respirometry and is useful for the design and diagnosis of experimental setups.
-The "`rMR`"" package (https://CRAN.R-project.org/package=rMR) performs interval-based metabolic rate calculations and has a method to automatically detect $P_{crit}$ using the "broken-stick" regression method (Yeager & Ultsch, 1989).
-The package "`LoLinR`"" [@olito_estimating_2017], while not specifically coded for respirometry, provides a statistically robust method of detecting a "best-fit" regression, and performs well at identifying truly linear subsets of a data series.
-Importantly, all of these packages assume that data are monotonic with evenly-spaced sampling periods with respect to time, an ideal structure but one where respirometry data often fails to meet.
+The `respirometry` package (https://CRAN.R-project.org/package=respirometry) contains a comprehensive collection of tools to explore and evaluate experimental parameters in aquatic respirometry and is useful for the design and diagnosis of experimental setups.
+The `rMR` package (https://CRAN.R-project.org/package=rMR) performs interval-based metabolic rate calculations and has a method to automatically detect $P_{crit}$ using the "broken-stick" regression method (Yeager & Ultsch, 1989).
+The package `LoLinR` [@olito_estimating_2017], while not specifically coded for respirometry, provides a statistically robust method of detecting a "best-fit" regression, and performs very well at identifying truly linear subsets of a data series.
+Importantly, all of these packages assume that data are monotonic with evenly-spaced sampling periods with respect to time, an ideal structure but one where respirometry data often fails to meet in reality.
 They also do not have the tools to analyse complicated datasets to automatically determine useful metric such as $MO_{2,max}$ and $MO_{2,min}$.
@@
 
-Typically, the analysis of respirometry data requires the selection one or more linear sections of the data for more accurate calculations of $RO_2$ [e.g.
+Typically, the analysis respirometry data requires the selection one or more linear sections of the raw data for more accurate calculations of $RO_2$ [e.g.
 @gordon_energetics_1989; @chu_ecophysiological_2017; @carey_sea_2016].
@@
 
-Units of measurement are not used when determining $RO_2$ in `respR`.
+For deliberate reasons, units of measurement are not used when determining $RO_2$ in `respR`.
 Instead, two functions, `convert_DO()` and `convert_rate()`, are used to transform $RO_2$ to specific units and perform volumetric or mass-specific conversions .

### 2017-11-19 index.Rmd
initial commit

---
author: "Januar Harianto"
title: "Introduction to respR"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{respR}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = F}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

# Introduction to `respR`
`respR` is an R package that provides a structural, reproducible workflow for the processing and analysis of respirometry and other related data. You are expected to at least have a basic understanding of R to use this package. If you know how to read Excel and/or .csv files and store them as `data.frame` objects in R, you should be able to use `respR` with relative ease.

It is recommend that you download the latest versions of RStudio (https://shiny.rstudio.com/) and R (http://www.r-project.org/). If you already have RStudio installed, go to the Help menu and click "Check for Updates".

```{r, echo=FALSE}
library(knitcitations); cleanbib()
cite_options(citation_format = "pandoc", check.entries = TRUE)
library(bibtex)
library(respR)
library(knitr)
library(readr)
library(readxl)
library(kableExtra)
Sys.setenv(TZ="Australia/Sydney")
# set pander table-layout options
library(pander)
panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)

# listfunctions <- readxl::read_excel('../tables/list-of-functions.xlsx')
# conversions <- readxl::read_excel("../tables/conversion-units.xlsx")

opts_chunk$set(collapse = TRUE, comment = "#>", tidy = TRUE, cache = TRUE, 
  highlight = TRUE, fig.width = 7, fig.height = 5)
```

# Package Installation

The latest version of `respR` can be installed from our GitHub repository. The installation process is easy with the `devtools` package, which needs to be installed first:

```{r, eval = F}
install.packages("devtools")
```

You can then install and load `respR` by running the following code:
```{r, eval = F}
devtools::install_github("januarharianto/respR") # Install respR
library(respR) # Load the package

```

Please note that the installation will also download all the dependencies required to run the package.

# Usage

When working with respirometry data, you will often need to:

1. Ensure that the data, or at least a subset of the data, is representative of the research question of interest.
2. Perform an initial analysis of the data to estimate the rate of change in oxygen concentration or amount.
3. Correct for background usage of oxygen by micro-organisms.
4. Scale the resulting usage rate to the volume (of the chamber/respirometry medium) and mass (of the animal/specimen), and convert to appropriate units.

The `respR` package contains several functions to make these actions quick and easy:

- It provides visual feedbacks and diagnostic plots to help you subset and analyse your data.
- It uses computational techniques such as rolling regressions and kernel density estimates to estimate ranked maximum, minimum or best fit sections of the data.
- By separating the workflow into a series of connected functions, you can "mix and match" functions to help you achieve your result.

# A typical workflow

To explore `respR` and its main functions we have provided example data that can be used immediately once the package is loaded. The data are `urchins.rd`, `intermittent.rd`, `sardine.rd` and `squid.rd`. These data were obtained from past experiments and more information can be obtained by invoking the `?` command, for instance, `?urchins.rd`.

## Step 1: Use `inspect_data()` to check for common errors
`respR` was designed to work best with `data.frame` objects. This includes objects of class `data.frame`, `data.table` and `tibble`. The user should already know how to import data into this format as it is the fundemental data structure in R. Once data is imported, we use `inspect_data()` to quickly scan it for common issues.

```{r}
# Check data for errors
int <- inspect_data(urchins.rd)
```

<!-- [inspect]: https://januarharianto.github.io/respR/reference/inspect_data.html -->
<!-- [bg]: https://januarharianto.github.io/respR/reference/calc_rate.bg.html -->
<!-- [calc]: https://januarharianto.github.io/respR/reference/calc_rate.html -->
<!-- [auto]: https://januarharianto.github.io/respR/reference/auto_rate.html -->
<!-- [adjust]: https://januarharianto.github.io/respR/reference/adjust_rate.html -->
<!-- [rate]: https://januarharianto.github.io/respR/reference/convert_rate.html -->

### 2017-11-21 R/adjust_rate.R
A few more edits to .R files

@@ -7,11 +7,11 @@
 #' To account for background respiration and/or oxygen flux in open respirometry
 #' experiments, we have provided a simple function to perform either, or both,
 #' corrections. As a separate function we may develop this to support more
-#' complex correction methods in the future.
+#' complex correction methods (e.g. non-linear) in the future.
 #'
 #' @md
 #' @param df data frame, or any object of class `calc_rate` or `auto_rate`. This
-#'   object contains the value that needs to be adjusted.
+#'   object contains the rate value that needs to be adjusted.
 #' @param by either a numeric, or an object of class `calc_rate.bg`. This is the
 #'   value that is used to perform the correction/adjustment.
 #'

### 2017-11-21 R/data.R

@@ -66,6 +66,16 @@
 
 
 #' Respirometry data of the sardine, *Sardinops sagax*
+#' 
+#' Dissolved oxygen units: %
+#' Temperature: ??
+#' Salinity: 35
+#' Atm. Pressure: ??
+#' Time units: seconds
+#' Chamber volume (L): ??
+#' Specimen wet mass (g): ??
+#' 
+#' Conducted at Hopkins Marine Station, Stanford University. 
 #'
 #' @md
 #' @author Nicholas Carey

### 2017-11-21 R/pcrit.R

@@ -12,7 +12,7 @@
 #' This is also the default method. If you already have calculated RO2/MO2
 #' elsewhere, the argument `has.rate = "TRUE"` must be used. The width of the
 #' rolling regression is determined by the `width` argument. In most cases, the
-#' default width works well.
+#' default width (10% of the data length) works well.
 #'
 #' @param df data frame or object of class `inspect_data`. This is the data to
 #'   analyse.

### 2021-06-25 pkgdown_script.R
pkgdown interim commit

@@ -1,0 +1,13 @@
+## Notes and commands
+##
+## For now - we keep README.md updated in here, and copy to README.md in respR directory.
+## Probably a better way of keeping them in sync, but for now do this.
+
+pkgdown::build_site()
+
+pkgdown::build_site(lazy = TRUE, devel = TRUE)
+
+pkgdown::preview_site(pkg = ".", path = ".", preview = NA)
+
+pkgdown::build_site(preview = TRUE)
+

### 2021-06-27 R/util_plots.R
inspect & inspect.ft - added check for numeric data

@@ -154,7 +154,7 @@ rollreg.p <- function(rolldf, ranked.b1) {
 
 plot_multi_ar <- function(x, n = 9){
   parorig <- par(no.readonly = TRUE) # save original par settings
-  on.exit(suppressWarnings(par(parorig))) # revert par settings to original
+  on.exit(par(parorig)) # revert par settings to original
 
   nres <- length(x$rate) ## no. of results
   df <- x$dataframe

### 2022-01-25 R/subset_rate.R
plot_ar interim commit

@@ -47,16 +47,17 @@
 #'
 #'  ## `NULL`, `unique`
 #'
-#'  Subsets only unique results, that is removes any duplicate regressions from
-#'  the input `$summary`. `n` is ignored.
+#'  Subsets only unique results, that is for any regressions in the input
+#'  `$summary` table that are identical (except the `$density` column), it
+#'  retains the top ranked one only. `n` is ignored.
 #'
 #'  ## `positive`, `negative`
 #'
-#'  Subsets all `positive` (>0) or `negative` (<0) rates. `n` is ignored. Useful
-#'  in intermittent respirometry where `auto_rate` may identify linear regions
-#'  of oxygen increase during flushes. Note, to keep mathematically consistent,
-#'  `respR` outputs oxygen consumption (i.e. respiration) rates as *negative*
-#'  values.
+#'  Subsets all `positive` (>0) or `negative` (<0) rates. `n` is ignored.
+#'  Useful, for example, in intermittent respirometry where `auto_rate` may
+#'  identify linear regions of oxygen increase during flushes. Note, `respR`
+#'  outputs oxygen consumption (i.e. respiration) rates as *negative* values,
+#'  production rates as *positive*.
 #'
 #'  ## `nonzero`, `zero`
 #'
@@ -71,8 +72,8 @@
 #'  rates are all negative, `method = 'highest'` will retain the highest
 #'  magnitude rates regardless of the sign, that is the *most negative*.
 #'  Essentially, these options ignore the sign of the rate. `n` indicates how
-#'  many of the lowest/highest rates to retain See `min` and `max` options for
-#'  extracting numerical low/high rates.
+#'  many of the lowest/highest rates to retain See `minimum` and `maximum`
+#'  options for extracting *numerically* lowest and highest rates.
 #'
 #'  ## `lowest_percentile`, `highest_percentile`
 #'
@@ -160,38 +161,47 @@
 #'  taking an average of the rate values may be questionable, as certain values
 #'  will be weighted higher due to these multiple, overlapping results. This
 #'  method removes overlapping rates, using `n` as a threshold to determine
-#'  degree of permitted overlap. It is recommended this method be used after all
+#'  degree of permitted overlap. It is recommended this method be used after
 #'  other selection criteria have been applied, as it is quite aggressive about
-#'  removing rates, including high-ranked ones, and can be computationally
-#'  intensive when there are many results. While it can be used with `auto_rate`
[…]

### 2022-01-25 R/util_plots.R

@@ -25,7 +25,10 @@ tck <- -0.01 # size of axis ticks
 mgp <- c(0, 0.1, 0)
 
 
-multi.p <- function(df, sdf, rsq, title = TRUE, xl = '', yl = '', legend = TRUE) {
+# axes = which axes to draw
+multi.p <- function(df, sdf, rsq, title = TRUE, xl = '', yl = '',
+                    axes = c(1,2,3), legend = TRUE) {
+
   names(df) <- c("x", "y")
   if (!is.null(nrow(sdf)))
     sdf <- list(sdf)
@@ -45,19 +48,21 @@ multi.p <- function(df, sdf, rsq, title = TRUE, xl = '', yl = '', legend = TRUE)
     abline(lm(y ~ x, z), lwd = 1.2, lty = 3)
   }))
 
-  axis(side = 1, col.axis = "blue", tck = tck, mgp = mgp)
-  axis(side = 2, col.axis = "black", tck = tck, mgp = mgp)
+  if(1 %in% axes) axis(side = 1, col.axis = "blue", tck = tck, mgp = mgp)
+  if(2 %in% axes) axis(side = 2, col.axis = "black", tck = tck, mgp = mgp)
+
   box()
-  if(legend) legend("bottomleft",
-                    "Time",
-                    text.col = "blue",
-                    bg = "gray90",
-                    cex = 0.5)
-  if(legend) legend("topright",
-                    "Row",
-                    text.col = "red",
-                    bg = "gray90",
-                    cex = 0.5)
+
+  if(legend && 1 %in% axes) legend("bottomleft",
+                                   "Time",
+                                   text.col = "blue",
+                                   bg = "gray90",
+                                   cex = 0.5)
+  if(legend && 3 %in% axes) legend("topright",
+                                   "Row",
+                                   text.col = "red",
+                                   bg = "gray90",
+                                   cex = 0.5)
   ## add row index axis
   par(new = TRUE)
   plot(seq(1, nrow(df)),
@@ -67,7 +72,7 @@ multi.p <- function(df, sdf, rsq, title = TRUE, xl = '', yl = '', legend = TRUE)
        pch = "",
        cex = .5,
        axes = FALSE)
-  axis(side = 3, col.axis = "red", tck = tck, mgp = mgp)
+  if(3 %in% axes) axis(side = 3, col.axis = "red", tck = tck, mgp = mgp)
 
   if (title == T)
[…]

### 2022-01-25 pkgdown/pkgdown_script.R

@@ -11,8 +11,10 @@ pkgdown::build_site(lazy = TRUE, devel = TRUE)
 
 pkgdown::preview_site(pkg = ".", path = ".", preview = NA)
 
+
 pkgdown::build_site(preview = TRUE)
 
 pkgdown::build_home(preview = TRUE)
 
 pkgdown::build_article("closed")
+pkgdown::build_article("subset_rate")

### 2022-02-03 pkgdown/pkgdown_script.R
various o2 to oxygen

@@ -9,7 +9,6 @@ pkgdown::build_site(lazy = TRUE, devel = TRUE)
 
 pkgdown::preview_site(pkg = ".", path = ".", preview = NA)
 
-
 pkgdown::build_site(preview = TRUE)
 
 pkgdown::build_home(preview = TRUE)

### 2022-02-03 R/convert_val.R
convert_rate - o2.unit renamed to oxy.unit

@@ -86,7 +86,7 @@
 #' # Here, chamber volume is 200 ml, pressure measured in mbar.
 #' x <- suppressWarnings(inspect(urchins.rd, 1, 2))
 #' rate <- calc_rate(x, from = 20, to = 30)
-#' convert_rate(rate, o2.unit = "ml/l", time.unit = "min",
+#' convert_rate(rate, oxy.unit = "ml/l", time.unit = "min",
 #'              output.unit = "mg/h",
 #'              volume = convert_val(200, "ml", "L"),
 #'              S = 35, t = 15,
@@ -97,7 +97,7 @@
 #' # to be specified:
 #' x <- suppressWarnings(inspect(urchins.rd, 1, 2))
 #' rate <- calc_rate(x, from = 20, to = 30)
-#' convert_rate(rate, o2.unit = "ml/l", time.unit = "min",
+#' convert_rate(rate, oxy.unit = "ml/l", time.unit = "min",
 #'              output.unit = "mg/h",
 #'              volume = convert_val(200, "ml"),
 #'              S = 35, t = 15,

### 2022-02-03 tests/testthat/test-convert_rate.ft.R

@@ -47,76 +47,76 @@ S=30
 t=15
 P=1.01
 
-conv.crft1 <- convert_rate.ft(crft1, o2.unit = "mg/l", flowrate.unit = "L/m",
+conv.crft1 <- convert_rate.ft(crft1, oxy.unit = "mg/l", flowrate.unit = "L/m",
                               output.unit = "mg/h",
                               mass = NULL, area = NULL,
                               S=S, t=t, P=P)
-conv.crft1.ms <- convert_rate.ft(crft1, o2.unit = "mg/l", flowrate.unit = "L/m",
+conv.crft1.ms <- convert_rate.ft(crft1, oxy.unit = "mg/l", flowrate.unit = "L/m",
                                  output.unit = "mg/h/g",
                                  mass = 0.4, area = NULL,
                                  S=S, t=t, P=P)
-conv.crft1.as <- convert_rate.ft(crft1, o2.unit = "mg/l", flowrate.unit = "L/m",
+conv.crft1.as <- convert_rate.ft(crft1, oxy.unit = "mg/l", flowrate.unit = "L/m",
                                  output.unit = "mg/h/cm2",
                                  mass = NULL, area = 0.01,
                                  S=S, t=t, P=P)
-conv.crftmany <- convert_rate.ft(crftmany, o2.unit = "mg/l", flowrate.unit = "L/m",
+conv.crftmany <- convert_rate.ft(crftmany, oxy.unit = "mg/l", flowrate.unit = "L/m",
                                  output.unit = "mg/h",
                                  mass = NULL, area = NULL,
                                  S=S, t=t, P=P)
-conv.crftmany.ms <- convert_rate.ft(crftmany, o2.unit = "mg/l", flowrate.unit = "L/m",
+conv.crftmany.ms <- convert_rate.ft(crftmany, oxy.unit = "mg/l", flowrate.unit = "L/m",
                                     output.unit = "mg/h/g",
                                     mass = 0.4, area = NULL,
                                     S=S, t=t, P=P)
-conv.crftmany.as <- convert_rate.ft(crftmany, o2.unit = "mg/l", flowrate.unit = "L/m",
+conv.crftmany.as <- convert_rate.ft(crftmany, oxy.unit = "mg/l", flowrate.unit = "L/m",
                                     output.unit = "mg/h/cm2",
                                     mass = NULL, area = 0.01,
[…]

### 2022-02-03 tests/testthat/test-calc_rate.ft.R
various o2 to oxy

@@ -14,17 +14,17 @@ capture.output({  ## stops printing console outputs on assigning
 df <-  data.frame(time = flowthrough.rd[[2]], delta = flowthrough.rd[[3]])
 
 ## inspect.ft objects
-insp.ft.obj.outo2.1col.ino2.1col<- suppressWarnings(suppressMessages(inspect.ft(flowthrough_mult.rd, time = 1,
+insp.ft.obj.outO2.1col.inO2.1col<- suppressWarnings(suppressMessages(inspect.ft(flowthrough_mult.rd, time = 1,
                                                                out.o2 = 2, in.o2 = 6, plot = F)))
-insp.ft.obj.outo2.1col.ino2.value <- suppressWarnings(suppressMessages(inspect.ft(flowthrough_mult.rd, time = 1, out.o2 = 2,
+insp.ft.obj.outO2.1col.inO2.value <- suppressWarnings(suppressMessages(inspect.ft(flowthrough_mult.rd, time = 1, out.o2 = 2,
                                                                  in.o2.value = 9.05, delta.o2 = NULL, plot = F)))
 insp.ft.obj.delta.1col <- suppressWarnings(suppressMessages(inspect.ft(flowthrough_mult.rd, time = 1, out.o2 = NULL,
                                                       in.o2 = NULL, delta.o2 = 10, plot = F)))
 
-insp.ft.obj.outo2.multcols.ino2.multcols <-
+insp.ft.obj.outO2.multcols.inO2.multcols <-
   suppressWarnings(suppressMessages(inspect.ft(flowthrough_mult.rd, time = 1, out.o2 = 2:4, in.o2 = 6:8, plot = F)))
 
-insp.ft.obj.outo2.multcols.ino2.value <-
+insp.ft.obj.outO2.multcols.inO2.value <-
   suppressWarnings(suppressMessages(inspect.ft(flowthrough_mult.rd, time = 1, out.o2 = 2:4, in.o2.value = 9.05, plot = F)))
 
 insp.ft.obj.delta.multcols <-
@@ -33,19 +33,19 @@ insp.ft.obj.delta.multcols <-
 ## calc_rate.ft objects for testing plotting, printing etc.
 crft.obj.value <- calc_rate.ft(-0.8,
                                flowrate = 2, plot = F)
-crft.obj.vector <- calc_rate.ft(flowthrough.rd$o2.delta,
+crft.obj.vector <- calc_rate.ft(flowthrough.rd$oxy.delta,
                                flowrate = 2, plot = F)
[…]

### 2022-02-03 R/adjust_rate.ft.R
flowthrough inputs renamed .o2 to .oxy

@@ -37,7 +37,7 @@
 #' over time (or there may be other variations such as in the inflow oxygen
 #' concentrations), it is recommended you *NOT* use this function. Instead, the
 #' paired blank oxygen concentration data should be used in [`inspect.ft`] as
-#' the `in.o2` input. In this way, the calculated specimen delta oxygen values
+#' the `in.oxy` input. In this way, the calculated specimen delta oxygen values
 #' take account of whatever background or other variation in oxygen is occurring
 #' in the blank chamber with respect to time.
 #'
@@ -169,7 +169,7 @@ adjust_rate.ft <- function(x, by) {
                 dataframe = x$dataframe,
                 data = x$data,
                 subsets = x$subsets,
-                delta.o2 = x$delta.o2,
+                delta.oxy = x$delta.oxy,
                 input_type = x$input_type,
                 summary = cbind(x$summary,
                                 adjustment = adjustment,

### 2022-02-03 R/calc_rate.ft.R

@@ -91,7 +91,7 @@
 #' saved objects can be used as the `by` input in [`adjust_rate.ft`]. For
 #' experiments in which the specimen data is to be corrected by a
 #' concurrently-run control experiment, best option is to use this as the
-#' `in.o2` input in [`inspect.ft()`]. See help file for that function, or the
+#' `in.oxy` input in [`inspect.ft()`]. See help file for that function, or the
 #' vignettes on the website for examples.
 #'
 #' ## S3 Generic Functions
@@ -208,11 +208,11 @@ calc_rate.ft <- function(x = NULL, flowrate = NULL, from = NULL, to = NULL,
     ## elements for output
     dfs <- NULL
     dt <- data.table::data.table(time = NA,
-                                 delta.o2 = delta)
-    data <- data.frame(delta.o2 = x)
+                                 delta.oxy = delta)
+    data <- data.frame(delta.oxy = x)
 
     ## summary for output
-    summary <- data.frame(out.o2 = NA, in.o2 = NA, delta.o2 = delta)
+    summary <- data.frame(out.oxy = NA, in.oxy = NA, delta.oxy = delta)
 
     # no need to plot
     if(plot) {
@@ -232,12 +232,12 @@ calc_rate.ft <- function(x = NULL, flowrate = NULL, from = NULL, to = NULL,
     ## empty elements for output
     dfs <- NULL
     dt <- data.table::data.table(time = NA,
-                                 delta.o2 = delta)
-    data <- data.frame(out.o2 = x[[1]],
-                       in.o2 = x[[2]])
+                                 delta.oxy = delta)
+    data <- data.frame(out.oxy = x[[1]],
+                       in.oxy = x[[2]])
 
     ## summary for output
-    summary <- data.frame(out.o2 = x[[1]], in.o2 = x[[2]], delta.o2 = delta)
+    summary <- data.frame(out.oxy = x[[1]], in.oxy = x[[2]], delta.oxy = delta)
 
     if(plot) {
       message("calc_rate.ft: plot only available for 'inspect.ft' inputs.")
@@ -256,8 +256,8 @@ calc_rate.ft <- function(x = NULL, flowrate = NULL, from = NULL, to = NULL,
 
     ## put data in dt
     time <- data$time[[1]]
-    delta.o2 <- data$delta.o2[[1]]
[…]

### 2022-02-03 tests/testthat/test-adjust_rate.ft.R

@@ -9,10 +9,10 @@ capture.output({  ## stops printing outputs on assigning
 
 # Create testing objects --------------------------------------------------
 
-insp.ft_obj <- inspect.ft(flowthrough.rd, delta.o2 = 4, plot = FALSE)
-## o2 production data
+insp.ft_obj <- inspect.ft(flowthrough.rd, delta.oxy = 4, plot = FALSE)
+## oxygen production data
 insp.ft_obj_prod <- inspect.ft(cbind(flowthrough.rd[,1], -1*(flowthrough.rd[,4])),
-                               time = 1, delta.o2 = 2, plot = FALSE)
+                               time = 1, delta.oxy = 2, plot = FALSE)
 
 # x objects
 x1 <- calc_rate.ft(insp.ft_obj,
@@ -36,12 +36,12 @@ xprod <- calc_rate.ft(insp.ft_obj_prod,
 
 by_val <- -0.7
 by_vec <- seq(-0.75, -0.85, -0.01)
-by_crft <- calc_rate.ft(suppressWarnings(inspect.ft(flowthrough_mult.rd, out.o2 = 4, in.o2 = 8, plot = FALSE)),
+by_crft <- calc_rate.ft(suppressWarnings(inspect.ft(flowthrough_mult.rd, out.oxy = 4, in.oxy = 8, plot = FALSE)),
                         flowrate = 2, plot = FALSE)
-  by_crft_mult <- calc_rate.ft(suppressWarnings(inspect.ft(flowthrough_mult.rd, out.o2 = 4, in.o2 = 8, plot = FALSE)),
+  by_crft_mult <- calc_rate.ft(suppressWarnings(inspect.ft(flowthrough_mult.rd, out.oxy = 4, in.oxy = 8, plot = FALSE)),
                              from = c(2000, 2500), to = c(2400, 2900), by = "row",
                              flowrate = 2, plot = FALSE)
-# by_crft_nl <- calc_rate.ft(inspect.ft(flowthrough_sim.rd, out.o2 = 2, in.o2 = 3, plot = FALSE),
+# by_crft_nl <- calc_rate.ft(inspect.ft(flowthrough_sim.rd, out.oxy = 2, in.oxy = 3, plot = FALSE),
 #                         flowrate = 2, plot = FALSE)
 
 adj.ft_obj.insp.1 <- adjust_rate.ft(x1, by = by_val)
[…]

### 2022-02-03 tests/testthat/test-calc_rate.R

@@ -206,7 +206,7 @@ test_that("calc_rate - correctly handles 'from' and 'to' NULL", {
                urch[[2]][1])
   expect_equal(calc_rate(urch, by = "row")$summary$endoxy,
                urch[[2]][nrow(urch)])
-  # by "o2"
+  # by "oxygen"
   expect_error(calc_rate(urch, by = "oxygen"),
                regexp = NA)
   expect_equal(calc_rate(urch, by = "oxygen")$summary$row,
@@ -471,8 +471,8 @@ test_that("calc_rate - by = 'row' outputs correct results with oxygen production
 
 # by = "oxygen" checks --------------------------------------------------------
 
-test_that("calc_rate - by = 'o2' stops with paired values of from and to *both* below or *both* above o2 data range", {
-  by <- "o2"
+test_that("calc_rate - by = 'oxygen' stops with paired values of from and to *both* below or *both* above oxygen data range", {
+  by <- "oxygen"
   expect_error(calc_rate(sardine.rd[1000:2000,], from = 90, to = 85, plot = F, by = by),
                "calc_rate: some paired 'from' and 'to' values are both below the range of oxygen data in 'x'.")
   expect_error(calc_rate(sardine.rd[1000:2000,], from = c(90,93.9), to = c(85,94.4), plot = F, by = by),
@@ -483,8 +483,8 @@ test_that("calc_rate - by = 'o2' stops with paired values of from and to *both* 
                "calc_rate: some paired 'from' and 'to' values are both above the range of oxygen data in 'x'.")
 })
 
-test_that("calc_rate - by = 'o2' warns with o2 values out of range, but uses closest available value", {
-  by <- "o2"
+test_that("calc_rate - by = 'oxygen' warns with oxygen values out of range, but uses closest available value", {
+  by <- "oxygen"
   # range is 89.7 to 95.7
   # single values, one above
   expect_message(calc_rate(sardine.rd, from = 100, to = 93, plot = F, by = by),
@@ -549,8 +549,8 @@ test_that("calc_rate - by = 'o2' warns with o2 values out of range, but uses clo
                nrow(sardine.rd))
 })
 
-test_that("calc_rate - by = 'o2' outputs correct results", {
-  by <- "o2"
+test_that("calc_rate - by = 'oxygen' outputs correct results", {
+  by <- "oxygen"
 
   from <- 94
[…]

### 2022-02-03 tests/testthat/test-calc_rate.ft.R

@@ -15,20 +15,20 @@ df <-  data.frame(time = flowthrough.rd[[2]], delta = flowthrough.rd[[3]])
 
 ## inspect.ft objects
 insp.ft.obj.outO2.1col.inO2.1col<- suppressWarnings(suppressMessages(inspect.ft(flowthrough_mult.rd, time = 1,
-                                                               out.o2 = 2, in.o2 = 6, plot = F)))
-insp.ft.obj.outO2.1col.inO2.value <- suppressWarnings(suppressMessages(inspect.ft(flowthrough_mult.rd, time = 1, out.o2 = 2,
-                                                                 in.o2.value = 9.05, delta.o2 = NULL, plot = F)))
-insp.ft.obj.delta.1col <- suppressWarnings(suppressMessages(inspect.ft(flowthrough_mult.rd, time = 1, out.o2 = NULL,
-                                                      in.o2 = NULL, delta.o2 = 10, plot = F)))
+                                                               out.oxy = 2, in.oxy = 6, plot = F)))
+insp.ft.obj.outO2.1col.inO2.value <- suppressWarnings(suppressMessages(inspect.ft(flowthrough_mult.rd, time = 1, out.oxy = 2,
+                                                                 in.oxy.value = 9.05, delta.oxy = NULL, plot = F)))
+insp.ft.obj.delta.1col <- suppressWarnings(suppressMessages(inspect.ft(flowthrough_mult.rd, time = 1, out.oxy = NULL,
+                                                      in.oxy = NULL, delta.oxy = 10, plot = F)))
 
 insp.ft.obj.outO2.multcols.inO2.multcols <-
-  suppressWarnings(suppressMessages(inspect.ft(flowthrough_mult.rd, time = 1, out.o2 = 2:4, in.o2 = 6:8, plot = F)))
+  suppressWarnings(suppressMessages(inspect.ft(flowthrough_mult.rd, time = 1, out.oxy = 2:4, in.oxy = 6:8, plot = F)))
 
 insp.ft.obj.outO2.multcols.inO2.value <-
-  suppressWarnings(suppressMessages(inspect.ft(flowthrough_mult.rd, time = 1, out.o2 = 2:4, in.o2.value = 9.05, plot = F)))
[…]

### 2022-02-08 R/adjust_rate.R
Huge documentation revision

@@ -1,18 +1,12 @@
-#' Adjustment function to account for background respiration or oxygen flux in
-#' respirometry experiments.
+#' Adjust rates to account for background respiration or oxygen flux.
 #'
 #' The `adjust_rate` function adjusts an oxygen uptake or production rate (for
 #' example, as determined in [`calc_rate()`] or [`auto_rate()`]) for background
-#' oxygen use by microbial organisms, or other removal (or input) of oxygen
+#' oxygen use by microbial organisms, or for other removal or input of oxygen
 #' during a respirometry experiment. The function accepts numeric values, as
 #' well as regular `respR` objects, and data frames. See [`calc_rate.bg()`] for
 #' determining background rates, which is the recommended way of passing
 #' background rates to `adjust_rate`.
-#'
-#' ***Note:*** take special care with the *sign* of the rate used for
-#' adjustments. In `respR` oxygen uptake rates are negative, as they represent a
-#' negative slope of oxygen against time. Background rates will normally also be
-#' a negative value. See Examples.
 #'
 #' `adjust_rate` allows the rate, or multiple rates, in `x` to be adjusted in a
 #' number of ways, as detailed below. Note that for those methods which accept
@@ -24,6 +18,11 @@
 #' exception to this is the `"paired"` method, where each rate in `by` (i.e.
 #' rate in each oxygen column) is paired with the rate at the same position in
 #' `x` and used to adjust it.
+#'
+#' ***Note:*** take special care with the *sign* of the rate used for
+#' adjustments. In `respR` oxygen uptake rates are negative, as they represent a
+#' negative slope of oxygen against time. Background rates will normally also be
+#' a negative value, while any input of oxygen would be positive. See Examples.
 #'
 #' ***Methods***
 #'
@@ -126,22 +125,22 @@
 #' numeric value, numeric vector, or a `calc_rate` or `auto_rate` object
 #' containing single or multiple rates. The `by` input is the first background
 #' recording or rate value, and `by2` the second background recording or rate
-#' value. Again, these must all share the *same numeric time data or scale, and
-#' time and oxygen units with each other, and with the rates in `x`. The `x`
-#' rates do not necessarily need to be at intermediate timepoints to the
-#' `by/by2` times. these are used only to establish a *time~background rate*
[…]

### 2022-02-08 R/auto_rate.R

@@ -3,15 +3,14 @@
 #'
 #' `auto_rate` performs rolling regressions on a dataset to determine the *most
 #' linear, highest, lowest, maximum, minimum, rolling*, and *interval* rates of
-#' change in oxygen concentration over time. Initially, a rolling regression of
-#' the specified `width` is performed on the entire dataset. Then, based on the
-#' "`method`" argument, the resulting regressions are ranked or ordered, and the
-#' output is summarised.
+#' change in oxygen against time. A rolling regression of the specified `width`
+#' is performed on the entire dataset, then based on the "`method`" input,
+#' the resulting regressions are ranked or ordered, and the output summarised.
 #'
 #' ## Ranking and ordering algorithms
 #'
 #' Currently, `auto_rate` contains seven ranking and ordering algorithms that
-#' can be applied using the `method` argument:
+#' can be applied using the `method` input:
 #'
 #' - `linear`: Uses kernel density estimation (KDE) to learn the shape of the
 #' entire dataset and *automatically identify* the most linear regions of the
@@ -67,13 +66,14 @@
 #' across the entire timeseries. No reordering of results is performed.
 #'
 #' - `interval`: multiple, successive, non-overlapping regressions of the
-#' specified 'width' are extracted from the rolled regressions, ordered by time.
+#' specified 'width' are extracted from the rolling regressions, ordered by
+#' time.
 #'
 #' - NOTE: `max`, `min`: These methods were used in previous versions of `respR`
 #' but have been deprecated. They were intended to order oxygen uptake
 #' (negative) rates by magnitude, but this resulted in incorrect ordering of
 #' oxygen production (positive) rates. They have been retained for code
-#' compatibility, but may be removed in a future version of `respR`, and so
+#' compatibility, but will be removed in a future version of `respR`, and so
 #' *should not be used*.
 #'
 #' ## Further selection and filtering of results
@@ -92,10 +92,10 @@
 #'
 #' The `width` input defaults to 0.2. If it is a value between 0 and 1, it
 #' represents a proportion of the total data length, as in the equation
-#' `floor(width * number of data rows)`. For example, 0.2 represents a
-#' regression rolling window of 20% of the data. Otherwise, if 1 or greater, the
+#' `floor(width * number of data rows)`. For example, 0.2 represents a rolling
[…]

### 2022-02-08 R/calc_rate.ft.R

@@ -8,12 +8,12 @@
 #' returned from different regions of continuous data, or a rolling rate of a
 #' specific window size performed across the whole dataset.
 #'
-#' The function calculates rates by averaging the delta oxygen values across the
-#' dataset, or from specified subsets of the data, and then using the `flowrate`
-#' to convert these to rates. There are no units involved in `calc_rate.ft`.
-#' This is a deliberate decision. The units of oxygen concentration and flowrate
-#' will be specified later in [`convert_rate.ft()`] when rates are converted to
-#' specific output units.
+#' `calc_rate.ft` calculates rates by averaging delta oxygen values across the
+#' whole dataset, or from specified subsets of the data. The `flowrate` is then
+#' used to convert these average delta values to rates. There are no units
+#' involved in `calc_rate.ft`. This is a deliberate decision. The units of
+#' oxygen concentration and flowrate will be specified later in
+#' [`convert_rate.ft()`] when rates are converted to specific output units.
 #'
 #' For continuous data recordings, it is recommended a `data.frame` containing
 #' the data be prepared via [`inspect.ft()`], and entered as the `x` input.
@@ -28,8 +28,8 @@
 #' ## Specifying regions
 #'
 #' For calculating rates over specific regions of the data, the `from` and `to`
-#' inputs in the `by` units of `"time"` (the default) or `"row"`) can be used
-#' for `inspect.ft()` inputs. All delta oxygen values within this region are
+#' inputs in the `by` units of `"time"` (the default) or `"row"` can be used for
+#' [`inspect.ft()`] inputs. All delta oxygen values within this region are
 #' converted to rates, and averaged to produce a overall rate for the region
 #' (`$rate` in the output). Multiple regions can be examined within the same
 #' dataset by entering `from` and `to` as vectors of paired values to specify
@@ -47,7 +47,7 @@
 #' In order to convert delta oxygen values to a oxygen uptake or production
 #' rate, the `flowrate` input is required. This must be in a volume (L, ml, or
 #' ul) per unit time (s,m,h,d), for example in `L/s`. The units are not required
-#' to be entered in here; you will specify them in [`convert_rate.ft()`] to
[…]

### 2022-02-08 R/convert_rate.R

@@ -14,29 +14,31 @@
 #'
 #' ## Respirometer volume
 #'
-#' Note, the `volume` input represents the *effective volume* of the
-#' respirometer, that is volume of fluid in the respirometry chamber, **not**
-#' the total respirometer volume or the specimen volume. Generally, this should
-#' be the volume of the respirometer minus the volume of the specimen.
+#' The `volume` of the respirometer is required and should be in litres (`L`).
+#' Note, the `volume` represents the *effective volume* of the respirometer,
+#' that is *volume of water* in the respirometry chamber. This is not
+#' necessarily the same as the volume of the respirometer. Typically, it is the
+#' volume of the respirometer *minus* the volume of the specimen.
 #' \href{https://github.com/nicholascarey/respfun#eff_vol}{See here} for help
-#' with calculating effective volumes.
+#' with calculating effective volumes. It also does not refer to the specimen
+#' volume.
 #'
 #' ## Units
 #'
-#' The `oxy.unit` of the original data used to calculated the rate is required.
-#' Concentration units should use only SI units (`L` or `kg`) for the
+#' The `oxy.unit` of the original raw data used to calculate the rate is
+#' required. Concentration units should use only SI units (`L` or `kg`) for the
 #' denominator, e.g. `"mg/L"`, `"mmol/kg"`. Percentage saturation of air or
 #' oxygen is accepted, as are oxygen pressure units. See [`unit_args()`] for
 #' details.
 #'
-#' The `time.unit` of the original data used to calculated the rate is also
+#' The `time.unit` of the original raw data used to calculate the rate is also
 #' required.
 #'
-#' An `output.unit` is also required. If left NULL, The default of `"mgO2/h"` is
-#' used, or `"mgO2/h/kg"` or `"mgO2/h/m2"` if a `mass` or `area` respectively
-#' has been entered. The `output.unit` must be in the sequence *O2-Time* (e.g.
-#' `"mg/h"`) for absolute rates, *O2-Time-Mass* (e.g. `"mg/h/kg"`) for
-#' mass-specific rates, and *O2-Time-Area* (e.g. `"mg/h/cm2"`) for surface
+#' An `output.unit` is also required. If left `NULL`, The default of `"mgO2/h"`
+#' is used, or `"mgO2/h/kg"` or `"mgO2/h/m2"` if a `mass` or `area` respectively
+#' has been entered. The `output.unit` must be in the sequence *Oxygen-Time*
[…]

### 2022-02-08 R/convert_rate.ft.R

@@ -5,7 +5,8 @@
 #' [`adjust_rate.ft()`] into an absolute rate (i.e. whole specimen or whole
 #' chamber), mass-specific rate (i.e. normalised by specimen mass), or
 #' area-specific rate (i.e. normalised by specimen surface area) in any common
-#' unit.
+#' unit. These should be rates calculated as an oxygen delta (inflow minus
+#' outflow oxygen) multiplied by the flowrate.
 #'
 #' By default, `convert_rate.ft` converts the `$rate` element from
 #' `calc_rate.ft` objects, or the `$rate.adjusted` element from `adjust_rate.ft`
@@ -14,43 +15,48 @@
 #'
 #' ## Units
 #'
-#' The `oxy.unit` of the original data used to calculated the rate is required.
-#' Concentration units should use only SI units (`L` or `kg`) for the
+#' The `oxy.unit` of the original raw data used to calculated the rate is
+#' required. Concentration units should use only SI units (`L` or `kg`) for the
 #' denominator, e.g. `"mg/L"`, `"mmol/kg"`. Percentage saturation of air or
 #' oxygen is accepted, as are oxygen pressure units. See [`unit_args()`] for
 #' details.
 #'
-#' An `output.unit` is also required. If left NULL, The default of `"mgO2/h"` is
-#' used, or `"mgO2/h/kg"` or `"mgO2/h/m2"` if a `mass` or `area` respectively
-#' has been entered. The `output.unit` must be in the sequence *O2-Time* (e.g.
-#' `"mg/h"`) for absolute rates, *O2-Time-Mass* (e.g. `"mg/h/kg"`) for
-#' mass-specific rates, and *O2-Time-Area* (e.g. `"mg/h/cm2"`) for surface
+#' An `output.unit` is also required. If left `NULL`, The default of `"mgO2/h"`
+#' is used, or `"mgO2/h/kg"` or `"mgO2/h/m2"` if a `mass` or `area` respectively
+#' has been entered. The `output.unit` must be in the sequence *Oxygen-Time*
+#' (e.g. `"mg/h"`) for absolute rates, *Oxygen-Time-Mass* (e.g. `"mg/h/kg"`) for
+#' mass-specific rates, and *Oxygen-Time-Area* (e.g. `"mg/h/cm2"`) for surface
 #' area-specific rates.
 #'
 #' Note, some oxygen input or output units require temperature (`t`) and
[…]

### 2022-02-08 R/convert_val.R

@@ -3,56 +3,62 @@
 #'
 #' This is a basic function that converts values of temperature, volume, mass,
 #' area, and atmospheric pressure to different units. This can be useful in
-#' [convert_DO()], [convert_rate()], and [convert_rate.ft()] in which some
-#' inputs must be in specific units (e.g. temperature in °C, atmospheric
-#' pressure in bar, area in m2). See Examples.
+#' [convert_DO()], [convert_rate()], and [convert_rate.ft()] where some inputs
+#' must be in specific units (e.g. temperature in °C, atmospheric pressure in
+#' bar, area in m2). See Examples.
+#'
+#' Note the type of unit does not need to be specified. The function will
+#' automatically recognise it using the `from` unit.
 #'
 #' If the `'to'` input is left `NULL`, the following defaults are applied
 #' depending on the unit type of the `from` input:
 #'
-#' - volume:        "L"
+#' - volume:        `"L"`
 #'
-#' - temperature:   "C"
+#' - temperature:   `"C"`
 #'
-#' - mass:          "kg"
+#' - mass:          `"kg"`
 #'
-#' - area:          "m2"
+#' - area:          `"m2"`
 #'
-#' - pressure:      "bar"
+#' - pressure:      `"bar"`
 #'
 #' A fuzzy string matching algorithm is used to accept different unit formatting
-#' styles. For example, `'msq', 'm2', 'M2', 'sqm'` are all parsed as metres
+#' styles. For example, `"msq"` `"m2"`, `"M2"`, `"sqm"` are all parsed as metres
 #' squared of area.
 #'
-#' **Units Accepted**
+#' ## Accepted Units
 #'
 #' *Temperature:*
 #'
-#' - "C", "K", "F"
+#' - `"C"`, `"K"`, `"F"`
 #'
 #' *Pressure:*
 #'
-#' - "kPa", "hPa", "Pa", "ubar", "mbar", "bar", "atm" (standard atmospheres),
-#' "Torr"
+#' - `"kPa"`, `"hPa"`, `"Pa"`, `"ubar"`, `"mbar"`, `"bar"`, `"Torr"`, `"atm"`
+#' (note, this is standard atmospheres).
 #'
 #' *Volume:*
 #'
-#' - "uL", "mL", "L"
+#' - `"uL"`, `"mL"`, `"L"`
 #'
 #' *Mass:*
 #'
[…]

### 2022-02-08 R/import_file.R

@@ -1,49 +1,63 @@
 #' Import respirometry system output files
 #'
 #' Automatically import data from different respirometry hardware and software
-#' systems. The aim is to work with most commercial DO sensors available in the
-#' market with little input from the user. The function extracts data columns
-#' from the file and generally cleans up column names (e.g. removes whitespace
-#' and characters which cause text encoding issues) to make the data easier to
-#' work with. Files should be raw output files where possible; files opened and
-#' re-saved in a different format or filetype will likely fail to import.
+#' systems. The aim is to work with most commercial oxygen sensors available in
+#' the market with minimal input from the user. The function extracts data
+#' columns from the file, removes redundant rows of metadata, and generally
+#' cleans up column names (e.g. removes whitespace and characters which cause
+#' text encoding issues) to make the data easier to work with. Files should be
+#' sensor system raw output files where possible; files opened and re-saved in a
+#' different format will likely fail to import.
 #'
-#' Currently works for:
-#'  - Firesting
-#'  - Pyro (another name for Firesting)
-#'  - PreSens OXY10
-#'  - PreSens OXY4
-#'  - PreSens (OxyView generic, including multiplate systems)
-#'  - PreSens/Loligo 24-Well Multiplate System (Excel files)
-#'  - MiniDOT
-#'  - Loligo AutoResp ('_raw' files output, not metadata files)
-#'  - Loligo Witrox (same as AutoResp, without metadata)
-#'  - Vernier (raw qmbl, csv, txt, (not yet gmbl))
-#'  - NeoFox
-#'  - Qbox Aqua
+#' Currently tested and working for:
+#'
+#' - Firesting
+#'
+#' - Pyro (another name for Firesting)
+#'
+#' - PreSens OXY10
+#'
+#' - PreSens OXY4
+#'
+#' - PreSens (OxyView generic, including multiplate systems)
+#'
+#' - PreSens/Loligo 24-Well Multiplate System (output Excel files)
+#'
+#' - MiniDOT
+#'
+#' - Loligo AutoResp ('_raw' files output, *not* metadata files)
+#'
+#' - Loligo Witrox (same as AutoResp, without metadata)
+#'
+#' - Vernier (raw qmbl, csv, or txt, (gmbl not yet supported))
+#'
+#' - NeoFox
+#'
+#' - Qbox Aqua
+#'
+#' Files with European numeric formatting (i.e. commas instead of points to
+#' denote decimals) are supported, and will be converted to point decimals on
[…]

### 2022-02-08 R/inspect.R

@@ -54,10 +54,10 @@
 #' oxygen against time. In these plots the axis is reversed so that higher
 #' uptake rates (i.e. more negative) will be higher on these plots. If you are
 #' interested instead in oxygen production rates, which are positive, the
-#' `rate.rev = FALSE` argument can be passed in either the `inspect` call, or
-#' when using `plot()` on the output object. In this case, the rate values will
-#' be plotted numerically, and higher oxygen *production* rates will be higher
-#' on the plot.
+#' `rate.rev = FALSE` input can be passed in either the `inspect` call, or when
+#' using `plot()` on the output object. In this case, the rate values will be
+#' plotted numerically, and higher oxygen *production* rates will be higher on
+#' the plot.
 #'
 #' ## Plot an additional data source
 #'
@@ -86,7 +86,7 @@
 #' ## Multiple Columns of Oxygen Data
 #'
 #' For a quick overview of larger datasets, multiple oxygen columns can be
-#' inspected for errors and plotted by using the `oxygen` argument to select
+#' inspected for errors and plotted by using the `oxygen` input to select
 #' multiple columns (by default, all non-time columns are inspected as oxygen
 #' columns if no `oxygen` input is specified). These must share the same `time`
 #' column. In this case, data checks are performed, with a plot of each oxygen
@@ -166,13 +166,14 @@
 #'   `tibble`, etc.).
 #' @param time integer. Defaults to 1. Specifies the column number of the Time
 #'   data.
-#' @param oxygen vector of integers. Defaults to 2. Specifies the column
-#'   number(s) of the Oxygen data.
-#' @param width numeric. Defaults to 0.1. Width used in the rolling regression
-#'   plot as proportion of total length of data (0.01 to 1).
-#' @param plot logical. Defaults to `TRUE`. Plots the data. If 2 columns
-#'   selected, plots timeseries data, plus plot of rolling rate. If multiple
-#'   columns, plots all timeseries data only.
+#' @param oxygen integer or vector of integers. Defaults to all non-time columns
+#'   if no other inputs given. Specifies the column number(s) of the Oxygen
+#'   data.
+#' @param width numeric, 0.01 to 1. Defaults to 0.1. Width used in the rolling
+#'   regression plot as proportion of total length of data.
[…]

### 2022-02-08 R/oxy_crit.R

@@ -1,16 +1,16 @@
 #' Calculate critical oxygen values, such as PCrit
 #'
-#' A function to calculate the critical oxygen value, the oxygen tension or
+#' A function to calculate critical oxygen values, the oxygen tension or
 #' concentration below which an uptake rate transitions from independent to
 #' dependent on the oxygen supply, typically known as *PCrit*.
 #'
-#' In earlier versions of `respR`, this function was known as `calc_pcrit`. It
-#' was renamed to avoid conflicts with a function of the same name in another
-#' package, and also because technically the *P* in  *PCrit* stands for the
-#' partial *pressure* of oxygen. Since the function returns the value in the
-#' units of the data as entered, whether they are concentration or pressure
-#' units, this terminology can be technically in error. Instead, for the
-#' purposes of the documentation we refer to this as the *Critical Oxygen
+#' In earlier versions of `respR`, this function was known as `pcrit` or
+#' `calc_pcrit`. It was renamed to avoid conflicts with functions of the same
+#' name in another package, and also because technically the *P* in  *PCrit*
+#' stands for the partial *pressure* of oxygen. Since the function returns the
+#' value in the units of the data as entered, whether they are concentration or
+#' pressure units, this terminology can be technically in error. Instead, for
+#' the purposes of the documentation we refer to this as the *Critical Oxygen
 #' Value*, or "*COV*". If the units of oxygen are partial pressure units (e.g.
 #' kPa), this is equivalent to PCrit, otherwise they should be reported with
 #' this in mind.
@@ -49,7 +49,7 @@
 #' ## Inputs
 #'
 #' The data input `x` should be an `inspect` object or `data.frame` containing
-#' oxygen~time data, *or* a `data.frame` containing rate~oxygen data.
+#' oxygen~time data, or a `data.frame` containing rate~oxygen data.
 #'
 #' ### Oxygen ~ Time data
 #'
@@ -68,7 +68,7 @@
 #' regression on the oxygen~time data to determine rates, and pairs these
 #' against a rolling mean of the oxygen data. The function then performs the
[…]

### 2022-02-08 R/subset_data.R

@@ -1,53 +1,54 @@
 #' Subset a `data.frame`, `inspect`, or `inspect.ft` object
 #'
 #' `subset_data` subsets a `data.frame`, `inspect`, or `inspect.ft` object based
-#' on a given set of criteria. The function is ideal for passing only some
-#' regions of your data to other functions such as [`calc_rate()`] and
-#' [`auto_rate()`], either by saving them as a new object or through the use of
-#' pipes (`%>%` or `|>`). It is also very useful with intermittent-flow data in
-#' loops, where each replicate can be extracted and passed to an analytical
-#' function such as `calc_rate` or `auto_rate`. See examples and vignettes.
+#' on a given set of criteria. The function is ideal for passing only selected
+#' regions of data to other functions such as [`calc_rate()`] and
+#' [`auto_rate()`], either by saving the output as a new object or via the use
+#' of pipes (`%>%` or `|>`). It is also very useful in analysis of
+#' intermittent-flow data, where in a loop each replicate can be extracted and
+#' passed to an analytical function such as `calc_rate` or `auto_rate`. See
+#' examples and vignettes.
 #'
-#' The function can subset data based on ranges of: `"time"`, `"oxygen"`,
-#' `"row"` , or `"proportion"` of total oxygen used or produced (note, this last
-#' option works poorly with noisy or fluctuating data). For data frames, to
-#' subset by `"time"`, `"oxygen"`, or `"proportion"`, the time data is assumed
-#' to be in the first column, and oxygen data in the second column. For
-#' `inspect()` and `inspect.ft()` objects, the data will have been coerced to
-#' this structure already. In these cases the `$dataframe` element in the output
-#' is replaced by the subset, and in `inspect.ft()` the `$inputs` element is
-#' also subset and replaced. Note for `inspect.ft()` objects, the oxygen data in
-#' column 2 will be either `out.oxy` data or `delta.oxy` data depending on what
-#' was inspected. The function can subset any data frame by `row`.
+#' The function can subset data based on ranges of `"time"`, `"oxygen"`, `"row"`
+#' , or `"proportion"` of total oxygen used or produced (note, this last option
+#' works poorly with noisy or fluctuating data). For data frames, to subset by
+#' `"time"`, `"oxygen"`, or `"proportion"`, the time data is assumed to be in
+#' the first column, and oxygen data in the second column. For [`inspect()`] and
[…]

### 2022-02-08 R/subset_rate.R

@@ -21,7 +21,7 @@
 #'  *Note:* when choosing a `method`, keep in mind that to remain mathematically
 #'  consistent, `respR` outputs oxygen consumption (i.e. respiration) rates as
 #'  negative values. This is particularly important in the difference between
-#'  `highest/lowest` and `min/max` methods. See Details.
+#'  `highest/lowest` and `minimum/maximum` methods. See Details.
 #'
 #'  When a `$rate` result is omitted by the subsetting criteria, all associated
 #'  data in `$summary` (i.e. the associated row) is removed. Summary results are
@@ -32,13 +32,13 @@
 #'  removal of duplicate results. Due to the way it works, the `auto_rate`
 #'  kernel-density analysis occasionally identifies linear regions that are
 #'  identical, and these appear as identical rows in the output `$summary`,
-#'  except for different `$density` values. For all methods, lower ranked
-#'  regressions are removed and only unique regressions are retained. To apply
-#'  only this duplicate removal method, use the default `method = NULL` or
-#'  `method = "unique"`.
+#'  except for different `$density` values. The lower ranked duplicates are
+#'  removed and only unique regressions are retained. To apply only this
+#'  duplicate removal method, use the default `method = NULL` or `method =
+#'  "unique"`.
 #'
 #'  Generally speaking, for most large datasets we recommend using
-#'  [subset_data()] and then running `auto_rate` on the subset(s) of the data
+#'  [`subset_data()`] and then running `auto_rate` on the subset(s) of the data
 #'  you are interested in, rather than run it on the whole dataset and relying
 #'  on `subset_rate` to filter it afterwards.
 #'
@@ -173,7 +173,7 @@
 #'  used in combination with the `auto_rate` `linear` results, unless you have a
 #'  specific reason for doing so.
 #'
-#'  The `plot_ar()` function is very useful for plotting `auto_rate` objects,
+#'  The [`plot_ar()`] function is very useful for plotting `auto_rate` objects,
 #'  and the results of `subset_rate` operations upon them, to visualise where
 #'  regression results in the summary table occur in relation to the original
 #'  dataset. See Examples.
@@ -777,7 +777,7 @@ subset_rate <- function(x, method = NULL, n = NULL, plot = TRUE){
 #     message("Over 20 results found. Plotting first 20 only...")
 #   }
 #   par(mar = c(0.1,0.1,0.1,0.1))
[…]

### 2022-02-08 R/unit_args.R

@@ -1,19 +1,25 @@
-#' Print examples of unit arguments for use in `convert_DO`, `convert_rate`, and
-#' `convert_rate.ft`
+#' Print examples of unit inputs for use in convert_DO, convert_rate, and
+#' convert_rate.ft
 #'
-#' This is a simple function, with no arguments. It prints to the console the
-#' units that can be used in the functions [convert_DO()], [convert_rate()], and
-#' [convert_rate.ft()]. Some oxygen unit conversions also require temperature
-#' (`t`), salinity (`S`), and atmospheric pressure (`P`) to be specified.
+#' This is a basic function with no inputs. It prints to the console the units
+#' that can be used in the functions [convert_DO()], [convert_rate()], and
+#' [convert_rate.ft()].
+#'
+#' Note that some oxygen unit conversions require temperature (`t`), salinity
+#' (`S`), and atmospheric pressure (`P`) to be specified.
 #'
 #' Note the difference between percent air saturation (`%Air`), where air
-#' saturated water is ~100%, and percent oxygen saturation (`%O2`), where air
-#' saturated water is ~20.946% *oxygen* saturated. In other words, `%O2 = %Air *
-#' 0.20946`.
+#' saturated water is ~100%, and percent oxygen saturation (`%Oxy`), where air
+#' saturated water is ~20.946% *oxygen* saturated. In other words, `%Oxy = %Air
+#' x 0.20946`.
 #'
-#' **[convert_DO()]**
+#' For most units a fuzzy string matching algorithm is used to accept different
+#' formatting styles. For example, `"mg/l"`, `"mg/L"`, `"mgL-1"`, `"mg l-1"`,
+#' `"mg.l-1"` are all parsed the same.
 #'
-#' **Oxygen concentration or pressure units for `from` and `to` arguments:**
+#' # [convert_DO()]
+#'
+#' ## Oxygen concentration or pressure units for `from` and `to`s:
 #'
 #' Oxygen concentration units should use SI units (`L` or `kg`) for the
 #' denominator.
@@ -25,12 +31,12 @@
[…]

### 2022-02-08 tests/testthat/test-auto_rate.R

@@ -34,10 +34,10 @@ test_that("auto_rate subsets out 2 columns from larger dataset", {
 
 test_that("auto_rate stops with malformed method", {
   expect_error(auto_rate(urch_data, method = "wrong"),
-               regexp = "auto_rate: The 'method' argument is not recognised: ")
+               regexp = "auto_rate: The 'method' input is not recognised: ")
 })
 
-test_that("auto_rate works using default arguments", {
+test_that("auto_rate works using default inputs", {
   ar <- auto_rate(urch_data, plot = F)
   expect_is(ar,
             "auto_rate")
@@ -47,7 +47,7 @@ test_that("auto_rate works using default arguments", {
             "data.frame")
 })
 
-test_that("auto_rate works by default with `highest` argument", {
+test_that("auto_rate works by default with `highest` input", {
   ar <- auto_rate(urch_data, plot = F, method = "highest")
   expect_is(ar,
             "auto_rate")
@@ -56,7 +56,7 @@ test_that("auto_rate works by default with `highest` argument", {
             "auto_rate")
 })
 
-test_that("auto_rate  works by default with `lowest` argument", {
+test_that("auto_rate  works by default with `lowest` input", {
   ar <- auto_rate(urch_data, plot = F, method = "lowest")
   expect_is(ar,
             "auto_rate")
@@ -92,15 +92,15 @@ test_that("auto_rate can be plotted with highest/lowest/interval methods", {
   suppressWarnings(file.remove("Rplots.pdf"))
 })
 
-test_that("auto_rate: panels 1 to 6 can be plot with choose", {
+test_that("auto_rate: panels 1 to 6 can be plot with panel", {
   ## plots individual panels
   ar <- auto_rate(urch_data, plot = F)
-  expect_output(plot(ar, choose =1))
-  expect_output(plot(ar, choose =2))
-  expect_output(plot(ar, choose =3))
-  expect_output(plot(ar, choose =4))
-  expect_output(plot(ar, choose =5))
-  expect_output(plot(ar, choose =6))
+  expect_output(plot(ar, panel =1))
+  expect_output(plot(ar, panel =2))
+  expect_output(plot(ar, panel =3))
+  expect_output(plot(ar, panel =4))
+  expect_output(plot(ar, panel =5))
+  expect_output(plot(ar, panel =6))
 })
 
 test_that("auto_rate: different results can be plot with pos", {
[…]

### 2022-02-08 tests/testthat/test-convert_DO.R

@@ -25,14 +25,14 @@ capture.output({  ## stops printing outputs on assigning
     expect_equal(round(convert_DO(10, "%Air", "umol/kg", S = 35, t = 25, P = 1.013253), 3), 20.619)
     expect_equal(round(convert_DO(10, "%Air", "ml/kg", S = 35, t = 25, P = 1.013253), 3), 0.504)
     expect_equal(round(convert_DO(10, "%Air", "%Air", S = 35, t = 25, P = 1.013253), 3), 10)
-    expect_equal(round(convert_DO(10, "%Air", "%O2", S = 35, t = 25, P = 1.013253), 3), 2.095)
+    expect_equal(round(convert_DO(10, "%Air", "%Oxy", S = 35, t = 25, P = 1.013253), 3), 2.095)
     expect_equal(round(convert_DO(10, "%Air", "hPa", S = 35, t = 25, P = 1.013253), 3), 20.854)
     expect_equal(round(convert_DO(10, "%Air", "kPa", S = 35, t = 25, P = 1.013253), 3), 2.085)
     expect_equal(round(convert_DO(10, "%Air", "mmHg", S = 35, t = 25, P = 1.013253), 3), 15.642)
     expect_equal(round(convert_DO(10, "%Air", "inHg", S = 35, t = 25, P = 1.013253), 3), 0.616)
     expect_equal(round(convert_DO(10, "%Air", "Torr", S = 35, t = 25, P = 1.013253), 3), 15.642)
   })
-  ## check that `from` arguments do not produce error
+  ## check that `from` inputs do not produce error
 
   ## regexp: regular expression to test against. If omitted, just asserts that
   ## code produces some output, messsage, warning or error. Alternatively, you can
@@ -50,7 +50,7 @@ capture.output({  ## stops printing outputs on assigning
     expect_error(convert_DO(10, "umol/kg", "mg/l", S = 35, t = 25, P = 1.013253), regexp = NA)
     expect_error(convert_DO(10, "ml/kg", "mg/l", S = 35, t = 25, P = 1.013253), regexp = NA)
     expect_error(convert_DO(10, "%Air", "mg/l", S = 35, t = 25, P = 1.013253), regexp = NA)
-    expect_error(convert_DO(10, "%O2", "mg/l", S = 35, t = 25, P = 1.013253), regexp = NA)
+    expect_error(convert_DO(10, "%Oxy", "mg/l", S = 35, t = 25, P = 1.013253), regexp = NA)
[…]

### 2022-02-08 tests/testthat/test-format_time.R

@@ -2,7 +2,7 @@
 ## testthat::test_file("tests/testthat/test-format_time.R")
 ## covr::file_coverage("R/format_time.R", "tests/testthat/test-format_time.R")
 
-test_that("format_time works with default arguments", {
+test_that("format_time works with default inputs", {
 
   x <- c("09-02-03 01:11:11", "09-02-03 02:11:11","09-02-03 02:25:11")
   expect_error(format_time(x), regexp = NA)
@@ -176,7 +176,7 @@ test_that("format_time works across midnight when no dates provided", {
   expect_error(format_time(x, time = 1, format = "HMS"),
                regexp = NA)
   expect_message(format_time(x, time = 1, format = "HMS"),
-                 regexp = "Time\\(s) cross midnight, attempting to parse correctly...")
+                 regexp = "Times cross midnight, attempting to parse correctly...")
 
   ## doesn't confuse 11 am for 11 pm
   x <- c("11:59:11", "00:11:11")
@@ -186,7 +186,7 @@ test_that("format_time works across midnight when no dates provided", {
                regexp = NA)
 
   expect_message(format_time(x, time = 1, format = "HMS"),
-                 regexp = "Time\\(s) cross midnight, attempting to parse correctly...")
+                 regexp = "Times cross midnight, attempting to parse correctly...")
   ## works with "p" suffix for AM/PM present
   x <- c("11:59:11 PM", "00:11:11 AM")
   expect_equal(format_time(x, time = 1, format = "HMSp")[2],
@@ -194,7 +194,7 @@ test_that("format_time works across midnight when no dates provided", {
   expect_error(format_time(x, time = 1, format = "HMSp"),
                regexp = NA)
   expect_message(format_time(x, time = 1, format = "HMSp"),
-                 regexp = "Time\\(s) cross midnight, attempting to parse correctly...")
+                 regexp = "Times cross midnight, attempting to parse correctly...")
 })
 
 ## This test in response to a failure to convert i still don't really understand.

### 2022-02-08 tests/testthat/test-inspect.ft.R

@@ -266,19 +266,19 @@ test_that("inspect.ft - stops if input column numbers found to conflict",{
                regexp = "inspect.ft: Input columns conflict.")
 })
 
-test_that("inspect.ft - stops if out.oxy entered and neither or in.oxy or in.oxy.value have arguments",{
+test_that("inspect.ft - stops if out.oxy entered and neither or in.oxy or in.oxy.value have inputs",{
   expect_error(inspect.ft(flowthrough.rd, time = 1, out.oxy = 2:3,
                           in.oxy = NULL, in.oxy.value = NULL, delta.oxy = NULL, plot = F),
                regexp = "inspect.ft: With 'out.oxy' data, paired 'in.oxy' columns or an 'in.oxy.value' is required.")
 })
 
-test_that("inspect.ft - stops if in.oxy entered and out.oxy does not have argument",{
+test_that("inspect.ft - stops if in.oxy entered and out.oxy does not have input",{
   expect_error(inspect.ft(flowthrough.rd, time = 1, out.oxy = NULL,
                           in.oxy = 5:7, in.oxy.value = NULL, delta.oxy = NULL, plot = F),
                regexp = "inspect.ft: An 'in.oxy' input requires paired 'out.oxy' column\\(s\\).")
 })
 
-test_that("inspect.ft - stops if both or in.oxy and in.oxy.value have arguments",{
+test_that("inspect.ft - stops if both or in.oxy and in.oxy.value have inputs",{
   expect_error(inspect.ft(flowthrough.rd, time = 1, out.oxy = 2:3,
                           in.oxy = 4, in.oxy.value = 5, delta.oxy = NULL, plot = F),
                regexp = "inspect.ft: Only one of 'in.oxy' or 'in.oxy.value' can be entered.")
@@ -308,7 +308,7 @@ test_that("inspect.ft - stops if out.oxy and in.oxy entered, but in.oxy does not
                regexp = NA)
 })
 
-test_that("inspect.ft - stops if delta.oxy entered and out.oxy, in.oxy or in.oxy.value have arguments",{
+test_that("inspect.ft - stops if delta.oxy entered and out.oxy, in.oxy or in.oxy.value have inputs",{
   expect_error(inspect.ft(flowthrough.rd, time = 1, out.oxy = 2,
                           in.oxy = NULL, in.oxy.value = 4, delta.oxy = 3, plot = F),
                regexp = "inspect.ft: With 'delta.oxy' data, 'out.oxy', 'in.oxy' and 'in.oxy.value' should be NULL.")

### 2022-02-08 tests/testthat/test-oxy_crit.R

@@ -155,7 +155,7 @@ capture.output({  ## stops printing of outputs on assigning
   })
 
   test_that("oxy_crit - oxy_crit S3 generics work with both oxygen/time and rate/oxygen inputs,
-            and both methods, export, and choose options", {
+            and both methods, export, and panel options", {
     # oxygen
     pcr <- oxy_crit(squid, plot = F)
     expect_output(print(pcr))
@@ -163,10 +163,10 @@ capture.output({  ## stops printing of outputs on assigning
     expect_is(summary(pcr, export = TRUE),
               "data.frame")
     expect_error(plot(pcr), regexp = NA)
-    expect_error(plot(pcr, choose = 1), regexp = NA)
-    expect_error(plot(pcr, choose = 2), regexp = NA)
-    expect_error(plot(pcr, choose = 3),
-                 "plot.oxy_crit: 'choose' input should be 1 to 2 or 'NULL' for both.")
+    expect_error(plot(pcr, panel = 1), regexp = NA)
+    expect_error(plot(pcr, panel = 2), regexp = NA)
+    expect_error(plot(pcr, panel = 3),
+                 "plot.oxy_crit: 'panel' input should be 1 to 2 or 'NULL' for both.")
     # rate
     pcr <- oxy_crit(squid_rate, oxygen = 1, rate = 2, plot = F)
     expect_output(print(pcr))
@@ -174,10 +174,10 @@ capture.output({  ## stops printing of outputs on assigning
     expect_is(summary(pcr, export = TRUE),
               "data.frame")
     expect_error(plot(pcr), regexp = NA)
-    expect_error(plot(pcr, choose = 1), regexp = NA)
-    expect_error(plot(pcr, choose = 2), regexp = NA)
-    expect_error(plot(pcr, choose = 3),
-                 "plot.oxy_crit: 'choose' input should be 1 to 2 or 'NULL' for both.")
+    expect_error(plot(pcr, panel = 1), regexp = NA)
+    expect_error(plot(pcr, panel = 2), regexp = NA)
+    expect_error(plot(pcr, panel = 3),
+                 "plot.oxy_crit: 'panel' input should be 1 to 2 or 'NULL' for both.")
     # segmented method
     # oxygen
     pcr <- oxy_crit(squid, method = "segmented", plot = F)
@@ -186,10 +186,10 @@ capture.output({  ## stops printing of outputs on assigning
[…]

### 2022-02-10 R/convert_DO.R
convert_DO - 'value.out' renamed to 'simplify', and mean() S3 added

@@ -20,14 +20,14 @@
 #'
 #' ## Output
 #'
-#' By default (`value.out = TRUE`) the output is a numeric vector of converted
-#' values. If `value.out = FALSE` output is a `list` object of class
+#' By default (`simplify = TRUE`) the output is a numeric vector of converted
+#' values. If `simplify = FALSE` output is a `list` object of class
 #' `convert_DO` containing five elements: `$call` the function call, `$input`
 #' values, `$output` converted values, `$input.unit` and `$output.unit`.
 #'
 #' ## S3 Generic Functions
 #'
-#' Saved output objects (if `value.out = FALSE` is used) can be entered in the
+#' Saved output objects (if `simplify = FALSE` is used) can be entered in the
 #' generic S3 functions `print()` and `summary()`.
 #'
 #' - `print()`: prints input and converted values (up to first 20), plus input
@@ -45,7 +45,7 @@
 #'   of some units. See [unit_args()] for details.
 #' @param P numeric. Pressure (bar). Defaults to 1.013253. Required for
 #'   conversion of some units. See [unit_args()] for details.
-#' @param value.out logical. Defaults to `TRUE` in which case the converted
+#' @param simplify logical. Defaults to `TRUE` in which case the converted
 #'   values are returned as a numeric vector. if `FALSE` a list object of class
 #'   `convert_DO` is returned.
 #'
@@ -71,7 +71,7 @@
 #'   }
 
 convert_DO <- function(x, from = NULL, to = NULL, S = NULL, t = NULL,
-                       P = NULL, value.out = TRUE) {
+                       P = NULL, simplify = TRUE) {
 
   ## Save function call for output
   call <- match.call()
@@ -174,7 +174,7 @@ convert_DO <- function(x, from = NULL, to = NULL, S = NULL, t = NULL,
 
   class(out) <- "convert_DO"
 
-  if(value.out) return(out$output) else
+  if(simplify) return(out$output) else
     return(out)
 }
 
@@ -211,10 +211,36 @@ plot.convert_DO <- function(x, ...) {
 }
 
 #' @export
-mean.convert_DO <- function(x, ...) {
-  message("convert_DO: mean() is not available for 'convert_DO' objects.")
-  return(invisible(x))
+mean.convert_DO <- function(x, pos = NULL, export = FALSE, ...){
+
+  cat("\n# mean.convert_DO # ---------------------\n")
[…]

### 2022-02-10 tests/testthat/test-convert_DO.R

@@ -105,8 +105,8 @@ capture.output({  ## stops printing outputs on assigning
 
   test_that("convert_DO S3 generics work", {
 
-    ob <- convert_DO(10, "inHg", "mg/l", S = 35, t = 25, P = 1.013253, value.out = FALSE)
-    ob_many <- convert_DO(10:30, "inHg", "mg/l", S = 35, t = 25, P = 1.013253, value.out = FALSE)
+    ob <- convert_DO(10, "inHg", "mg/l", S = 35, t = 25, P = 1.013253, simplify = FALSE)
+    ob_many <- convert_DO(10:30, "inHg", "mg/l", S = 35, t = 25, P = 1.013253, simplify = FALSE)
 
     expect_error(print(ob),
                  NA)

### 2022-02-15 R/adjust_rate.ft.R
CRAN updates

@@ -44,21 +44,7 @@
 #' **Note:** take special care with the *sign* of the rate used for adjustments.
 #' In `respR` oxygen uptake rates are negative, as they represent a negative
 #' slope of oxygen against time. Background rates will normally also be a
-#' negative value. See Examples.
-#'
-#' ## Output
-#'
-#' If the `x` input is a `calc_rate.ft` object, the output will be identical in
-#' structure, but of class `adjust_rate.ft` and containing the additional
-#' elements `$adjustment` and `$rate.adjusted`, with these also added to
-#' `$summary` metadata.
-#'
-#' If `x` is a numeric value or vector, the output is a `list` object of class
-#' `adjust_rate.ft` containing four elements: a `$summary` table, `$rate`,
-#' `$adjustment`, and `$rate.adjusted`.
-#'
-#' For all outputs, the `$rate.adjusted` element will be the one converted when
-#' the object is passed to `convert_rate.ft`.
+#' negative value (though not always). See Examples.
 #'
 #' ## S3 Generic Functions
 #'
@@ -76,7 +62,17 @@
 #' the `pos` input. e.g. `mean(x, pos = 1:5)` The mean can be exported as a
 #' separate value by passing `export = TRUE`.
 #'
-#' @md
+#' @return **Output**: If the `x` input is a `calc_rate.ft` object, the output will be identical
+#'   in structure, but of class `adjust_rate.ft` and containing the additional
+#'   elements `$adjustment` and `$rate.adjusted`, with these also added to
+#'   `$summary` metadata.
+#'
+#'   If `x` is a numeric value or vector, the output is a `list` object of class
+#'   `adjust_rate.ft` containing four elements: a `$summary` table, `$rate`,
+#'   `$adjustment`, and `$rate.adjusted`.
+#'
+#'   For all outputs, the `$rate.adjusted` element will be the one converted
+#'   when the object is passed to `convert_rate.ft`.
 #'
 #' @param x numeric. A single numeric value, numeric vector, or object of class
 #'   `calc_rate.ft`. This is the experimental rate value(s) to be adjusted.

### 2022-02-15 R/calc_rate.bg.R

@@ -42,16 +42,6 @@
 #' readily capable of quantifying production rates as well as consumption, and
 #' these can also be used for adjustments in [`adjust_rate()`].
 #'
-#' ## Output
-#'
-#' Output is a `list` object of class `calc_rate.bg` containing original data,
-#' linear models, summary information, and the primary output of interest
-#' `$rate.bg`, which contains a rate for each oxygen column present in the input
-#' data. There is also `$rate.bg.mean` containing the mean of all background
-#' rates. Note, this is not used in `adjust_rate`, where the `method` input
-#' there determines how `$rate.bg` is applied, but can easily be extracted and
-#' applied as an adjustment value if desired.
-#'
 #' ## Plot
 #'
 #' A plot is produced (provided `plot = TRUE`) showing all examined columns of
@@ -76,6 +66,14 @@
 #' - `mean()`: calculates the mean of all rates, or those specified by the `pos`
 #' input. e.g. `mean(x, pos = 1:5)` The mean can be exported as a separate value
 #' by passing `export = TRUE`.
+#'
+#' @return Output is a `list` object of class `calc_rate.bg` containing original
+#'   data, linear models, summary information, and the primary output of
+#'   interest `$rate.bg`, which contains a rate for each oxygen column present
+#'   in the input data. There is also `$rate.bg.mean` containing the mean of all
+#'   background rates. Note, this is not used in `adjust_rate`, where the
+#'   `method` input there determines how `$rate.bg` is applied, but can easily
+#'   be extracted and applied as an adjustment value if desired.
 #'
 #' @param x `data.frame` or `inspect` object. This is the data to extract
 #'   background rate(s) from.

### 2022-02-15 R/calc_rate.ft.R

@@ -110,19 +110,17 @@
 #' input. e.g. `mean(x, pos = 1:5)` The mean can be exported as a separate value
 #' by passing `export = TRUE`.
 #'
-#' ## Output
-#'
-#' Output is a `list` object of class `calc_rate.ft` containing input parameters
-#' and data, various summary data, metadata, and the primary output of interest
-#' `$rate`, which can be background adjusted in [`adjust_rate.ft`] or converted
-#' to units in [`convert_rate.ft`]. Note the `$summary` table contains linear
-#' regression coefficients alongside other metadata. These *should not be
-#' confused* with those in other functions such as `calc_rate` where slopes
-#' represent rates and coefficients such as a high r-squared are important.
-#' Here, they represent the stability of the data region, in that the closer the
-#' slope is to zero the less the delta oxygen values, and therefore rates, in
-#' that region vary. These are included to enable possible future functionality
-#' where stable regions may be automatically identified.
+#' @return Output is a `list` object of class `calc_rate.ft` containing input
+#'   parameters and data, various summary data, metadata, and the primary output
+#'   of interest `$rate`, which can be background adjusted in [`adjust_rate.ft`]
+#'   or converted to units in [`convert_rate.ft`]. Note the `$summary` table
+#'   contains linear regression coefficients alongside other metadata. These
+#'   *should not be confused* with those in other functions such as `calc_rate`
+#'   where slopes represent rates and coefficients such as a high r-squared are
+#'   important. Here, they represent the stability of the data region, in that
+#'   the closer the slope is to zero the less the delta oxygen values, and
+#'   therefore rates, in that region vary. These are included to enable possible
+#'   future functionality where stable regions may be automatically identified.
 #'
 #' @param x numeric value or vector of delta oxygen values, a 2-column
 #'   `data.frame` of outflow (col 1) and inflow (col 2) oxygen values, or an

### 2022-02-15 R/convert_rate.ft.R

@@ -54,12 +54,6 @@
 #' [`unit_args()`] for details of accepted units and their formatting. See also
 #' [`convert_val()`] for simple conversion between non-oxygen units.
 #'
-#' ## Output
-#'
-#' Output is a `list` object containing the `$rate.input`, and converted rate(s)
-#' in `$rate.output` in the `$output.unit`, as well as inputs and summary
-#' elements.
-#'
 #' ## S3 Generic Functions
 #'
 #' Saved output objects can be used in the generic S3 functions `print()`,
@@ -75,6 +69,10 @@
 #' - `mean()`: calculates the mean of all converted rates, or those specified by
 #' the `pos` input. e.g. `mean(x, pos = 1:5)` The mean can be exported as a
 #' separate value by passing `export = TRUE`.
+#'
+#' @return Output is a `list` object containing the `$rate.input`, and converted
+#'   rate(s) in `$rate.output` in the `$output.unit`, as well as inputs and
+#'   summary elements.
 #'
 #' @param x numeric value or vector, or object of class [calc_rate.ft()] or
 #'   [adjust_rate.ft()]. Contains the rate(s) to be converted.

### 2022-02-15 R/format_time.R

@@ -3,8 +3,11 @@
 #' A function to parse class POSIX.ct or text strings of date-time data to
 #' numeric time for use in `respR` functions.
 #'
-#' `format_time` uses conversion functionality from the \code{\link{lubridate}}
-#' package.
+#' Regardless of input, all data are parsed to numeric time data in seconds
+#' duration from the first entry starting at 1. If you want the times to start
+#' at a different time, a `start` value can be specified, in which case the
+#' series starts at that number (in seconds) and all subsequent times are
+#' shifted forward by the same amount.
 #'
 #' ## Input
 #'
@@ -16,18 +19,6 @@
 #' time in another), multiple columns can be specified (e.g. `time = c(1,2)`).
 #' In this case, the `format` setting should reflect the correct order as
 #' entered in `time`.
-#'
-#' ## Output
-#'
-#' If input is a vector, output is a vector of equal length containing the
-#' numeric time data. For data frame inputs, an identical data frame is
-#' returned, with a new column called `time_num` added as the **last** column.
-#'
-#' Regardless of input, all data are parsed to numeric time data in seconds
-#' duration from the first entry starting at 1. If you want the times to start
-#' at a different time, a `start` value can be specified, in which case the
-#' series starts at that number (in seconds) and all subsequent times are
-#' shifted forward by the same amount.
 #'
 #' ## Time only data
 #'
@@ -84,6 +75,11 @@
 #' Single experimental datasets should never span different time zones, so if a
 #' time zone is present it is ignored for the purposes of calculating numeric
 #' times.
+#'
+#' @return **Output**: If the input is a vector, output is a vector of equal
+#'   length containing the numeric time data. For data frame inputs, an
+#'   identical data frame is returned, with a new column named `time_num` added
+#'   as the **final** column.
 #'
 #' @param x vector or data frame containing strings or class POSIX.ct date-time
 #'   data to be converted to numeric.

### 2022-02-15 R/import_file.R

@@ -51,9 +51,7 @@
 #' import data by other functions such as [`read.csv()`] since it is a
 #' fundamental basic in data analyses in R.
 #'
-#' ## Output
-#'
-#' A `data.frame` object of all columned data
+#' @return A `data.frame` object of all columned data
 #'
 #' @param path string. Path to file.
 #' @param export logical. If TRUE, exports the data as a `csv` to the same

### 2022-02-15 R/inspect.ft.R

@@ -218,14 +218,6 @@
 #' **Note**: All background calculations should be from experiments done at the
 #' *same flow rate* as the specimen experiments to be corrected.
 #'
-#' ## Output
-#'
-#' Output is a `list` object of class `inspect.ft` containing input parameters
-#' and data, data check summaries, and metadata, which can be passed to
-#' [`calc_rate.ft()`] to determine rates. If there are failed checks or
-#' warnings, the row locations of the potentially problematic data can be found
-#' in `$locs`.
-#'
 #' ## S3 Generic Functions
 #'
 #' Saved output objects can be used in the generic S3 functions `plot()`,
@@ -237,6 +229,12 @@
 #' are found, locations (row numbers) are printed (up to first 20 occurrences).
 #'
 #' - `summary()`: simple wrapper for `print()` function. See above.
+#'
+#' @return Output is a `list` object of class `inspect.ft` containing input
+#'   parameters and data, data check summaries, and metadata, which can be
+#'   passed to [`calc_rate.ft()`] to determine rates. If there are failed checks
+#'   or warnings, the row locations of the potentially problematic data can be
+#'   found in `$locs`.
 #'
 #' @param x `data.frame` containing columns of `time` and `out.oxy` or
 #'   `delta.oxy` concentrations, and optionally `in.oxy`.

### 2022-02-15 R/subset_rate.R

@@ -1,229 +1,234 @@
-#'@title Subset auto_rate results based on a range of criteria
+#' @title Subset auto_rate results based on a range of criteria
 #'
-#'@description The `auto_rate` function is powerful, but the output can be large
-#'  and difficult to explore. In addition, it may identify linear data regions,
-#'  but from areas of the data that are not of experimental interest.  As an
-#'  advanced, machine learning based process, it is also somewhat fallible and
-#'  on occasion may return questionable results.
+#' @description The `auto_rate` function is powerful, but the output can be
+#'   large and difficult to explore. In addition, it may identify linear data
+#'   regions, but from areas of the data that are not of experimental interest.
+#'   As an advanced, machine learning based process, it is also somewhat
+#'   fallible and on occasion may return questionable results.
 #'
-#'  The `subset_rate` function helps explore and filter `auto_rate` results by
-#'  subsetting according to various criteria. For example, extracting only
-#'  positive or negative rates, only the highest or lowest rates, only those
-#'  from certain data regions, and numerous other methods that allow advanced
-#'  filtering of results so the rates extracted are well-defined towards the
-#'  research question of interest. This also allows for highly consistent
-#'  reporting of results and rate selection criteria.
+#'   The `subset_rate` function helps explore and filter `auto_rate` results by
+#'   subsetting according to various criteria. For example, extracting only
+#'   positive or negative rates, only the highest or lowest rates, only those
+#'   from certain data regions, and numerous other methods that allow advanced
+#'   filtering of results so the rates extracted are well-defined towards the
+#'   research question of interest. This also allows for highly consistent
+#'   reporting of results and rate selection criteria.
 #'
-#'  Multiple subsetting criteria can be applied by assigning the output and
-#'  processing it through the function multiple times with different criteria,
-#'  or alternatively via `%>%` piping. See Examples.
+#'   Multiple subsetting criteria can be applied by assigning the output and
+#'   processing it through the function multiple times with different criteria,
+#'   or alternatively via `%>%` piping. See Examples.
 #'
-#'  *Note:* when choosing a `method`, keep in mind that to remain mathematically
[…]

### 2022-02-15 R/unit_args.R

@@ -90,6 +90,8 @@
 #'
 #' - `"mm2", "cm2", "m2", "km2"`
 #'
+#' @return A print out to the console of accepted units
+#'
 #' @export
 #'
 #' @examples

### 2022-02-15 R/util_plots.R

@@ -301,6 +301,8 @@ plot_multi_ar <- function(x, n = 9, ...){
 #' @param legend logical. Suppress plot legends.
 #' @param ... Allows additional plotting controls to be passed.
 #'
+#' @return A plot of the auto_rate object results
+#'
 #' @export
 #' @keywords internal
 

### 2022-02-15 pkgdown/pkgdown_script.R

@@ -17,3 +17,5 @@ pkgdown::build_home(preview = TRUE)
 
 pkgdown::build_article("closed")
 pkgdown::build_article("subset_rate")
+
+usethis::use_vignette("subsample", "subsample: Thin larger datasets")

### 2022-02-18 pkgdown/pkgdown_script.R
vignette updates

@@ -1,22 +1,21 @@
-## Notes and commands
-##
-## For now - we keep README.md updated in here, and copy to README.md in respR directory.
-## Probably a better way of keeping them in sync, but for now do this.
+## Build site
 
 pkgdown::build_site()
+
+pkgdown::build_site(preview = TRUE)
+# d oit in the background
+job::job({pkgdown::build_site(preview = TRUE)})
 
 ## Quicker
 ## Only builds what has changed
 pkgdown::build_site(lazy = TRUE, devel = TRUE)
 
+# ?
 #pkgdown::preview_site(pkg = ".", path = ".", preview = NA)
-
-pkgdown::build_site(preview = TRUE)
-job::job({pkgdown::build_site(preview = TRUE)})
 
 pkgdown::build_home(preview = TRUE)
 
 pkgdown::build_article("closed")
 pkgdown::build_article("subset_rate")
 
-usethis::use_vignette("subsample", "subsample: Thin larger datasets")
+usethis::use_vignette("whatsnew", "What's new")

### 2022-02-20 R/subset_rate.R
Vig updates

@@ -97,16 +97,17 @@
 #'   ## `rate`, `rsq`, `row`, `time`, `density`
 #'
 #'   These methods refer to the respective columns of the `$summary` data frame.
-#'   For these, `n` should be a numeric vector of two values. Matching rates
-#'   which fall within the `n` range (inclusive) are retained. For example, to
-#'   retain only rates where the `rate` value is between 0.05 and 0.08: `method
-#'   = 'rate', n = c(0.05, 0.08)`. To retain all rates with a R-Squared above
-#'   0.90: `method = 'rsq', n = c(0.9, 1)`. The `row` and `time` ranges refer to
-#'   the `$row`-`$endrow` or `$time`-`$endtime` columns and original data source
-#'   (`$dataframe` element of the input), and can be used to constrain results
-#'   to rates from particular regions of the data (although usually a better
-#'   option is to \code{\link{subset_data}} prior to analysis). Note, `time` is
-#'   not the same as `duration` - see later section.
+#'   For these, `n` should be a vector of two values in numeric order (lowest to
+#'   highest numerically). Matching rates which fall within the `n` range
+#'   (inclusive) are retained. For example, to retain only rates where the
+#'   `rate` value is between 0.05 and 0.08: `method = 'rate', n = c(0.05,
+#'   0.08)`. To retain all rates with a R-Squared above 0.90: `method = 'rsq', n
+#'   = c(0.9, 1)`. The `row` and `time` ranges refer to the `$row`-`$endrow` or
+#'   `$time`-`$endtime` columns and original data source (`$dataframe` element
+#'   of the input), and can be used to constrain results to rates from
+#'   particular regions of the data (although usually a better option is to
+#'   \code{\link{subset_data}} prior to analysis). Note, `time` is not the same
+#'   as `duration` - see later section.
 #'
 #'   ## `time_omit`, `row_omit`
 #'

### 2022-02-20 pkgdown/pkgdown_script.R

@@ -3,14 +3,15 @@
 pkgdown::build_site()
 
 pkgdown::build_site(preview = TRUE)
-# d oit in the background
+
+# do it in the background
 job::job({pkgdown::build_site(preview = TRUE)})
 
 ## Quicker
 ## Only builds what has changed
 pkgdown::build_site(lazy = TRUE, devel = TRUE)
 
-# ?
+# ? not sure what this does
 #pkgdown::preview_site(pkg = ".", path = ".", preview = NA)
 
 pkgdown::build_home(preview = TRUE)
@@ -18,4 +19,4 @@ pkgdown::build_home(preview = TRUE)
 pkgdown::build_article("closed")
 pkgdown::build_article("subset_rate")
 
-usethis::use_vignette("whatsnew", "What's new")
+usethis::use_vignette("support", "Support Us")

### 2022-03-07 R/subset_rate.R
subset_rate plot revised

@@ -196,6 +196,18 @@
 #'   data, are retained. If `n = 1`, only rates which are 100% contained within
 #'   at least one other are removed.
 #'
+#'   ## Plot
+#'
+#'   While output objects are plotted as normal `auto_rate` objects if used in
+#'   `plot()`, `subset_rate` has its own plotting functionality. This simply
+#'   plots a grid of the remaining rates in `$summary` up to the first 20. This
+#'   is simple functionality to give the user an idea of how subsetting is
+#'   reducing the number of rates, and where the remaining rates occur within
+#'   the data. It is really only useful once rates are down to fewer than 20
+#'   remaining, or for examining the effects of different subsetting options on
+#'   a small selection of rates. Therefore, the default is `plot = FALSE` to
+#'   prevent this being produced for every single subsetting operation.
+#'
 #'   ## More
 #'
 #'   For additional help, documentation, vignettes, and more visit the `respR`
@@ -226,8 +238,8 @@
 #'   results are *retained* in the output. See Details.
 #' @param n numeric. Number, percentile, or range of results to return depending
 #'   on `method`. See Details.
-#' @param plot logical. Plots a summary of subset locations within data (up to a
-#'   maximum of the first 9 ranked results).
+#' @param plot logical. Default FALSE. Plots a summary of subset locations
+#'   within data (up to a maximum of the first 20 ranked results).
 #'
 #' @export
 #'
@@ -240,7 +252,7 @@
 #'  auto_rate(plot = FALSE)
 #'  ar_subs_neg <- subset_rate(ar_obj, method = "negative", plot = FALSE)
 
-subset_rate <- function(x, method = NULL, n = NULL, plot = TRUE){
+subset_rate <- function(x, method = NULL, n = NULL, plot = FALSE){
 
   ## Save function call for output
   call <- match.call()
@@ -723,43 +735,18 @@ subset_rate <- function(x, method = NULL, n = NULL, plot = TRUE){
   if(!("auto_rate_subset" %in% class(output))) class(output) <- c(class(output), "auto_rate_subset")
 
   # Plot --------------------------------------------------------------------
+  if(plot) plot_ar_grid(output)
 
-  ## This is SLOOOWWWWWW AS F
[…]

### 2022-03-07 R/util_plots.R

@@ -62,8 +62,14 @@ las_def <- 0
 # Plotting functions ------------------------------------------------------
 
 # axes = which axes to draw
+# mgp_bt etc = separate mgp for axes if needed
 multi.p <- function(df, sdf, rsq, title = TRUE, xl = '', yl = '',
-                    axes = c(1,2,3), legend = TRUE, ...) {
+                    axes = c(1,2,3), legend = TRUE,
+                    bt_mgp = NULL,
+                    lf_mgp = NULL,
+                    tp_mgp = NULL,
+                    rt_mgp = NULL,
+                    ...) {
 
   par(...)
   names(df) <- c("x", "y")
@@ -84,8 +90,8 @@ multi.p <- function(df, sdf, rsq, title = TRUE, xl = '', yl = '',
     abline(lm(y ~ x, z), lwd = 1.2, lty = 3)
   }))
 
-  if(1 %in% axes) axis(side = 1, col.axis = "blue")
-  if(2 %in% axes) axis(side = 2, col.axis = "black")
+  if(1 %in% axes) axis(side = 1, col.axis = "blue", mgp = bt_mgp)
+  if(2 %in% axes) axis(side = 2, col.axis = "black", mgp = lf_mgp)
 
   box()
 
@@ -108,7 +114,7 @@ multi.p <- function(df, sdf, rsq, title = TRUE, xl = '', yl = '',
        pch = "",
        cex = .5,
        axes = FALSE)
-  if(3 %in% axes) axis(side = 3, col.axis = "red")
+  if(3 %in% axes) axis(side = 3, col.axis = "red", mgp = tp_mgp)
 
   if (title == T)
     title(main = ("Full Timeseries"), line = 1.2, font = 2)
@@ -217,74 +223,110 @@ rollreg.p <- function(rolldf, ranked.b1, rownums, xlim, rate.rev = TRUE, ...) {
 
 
 #' Plots multiple auto_rate results in a nice way
-#' x = auto_rate or auto_rate_filt object
-#' n = max no. of plots
-#' THIS IS SLOOOOOOOW
-#' Will probably have to revert to base plotting
-#'
-#' @importFrom cowplot plot_grid
-#' @importFrom methods show
-#' @import ggplot2
+#' using base plot
+#' x = auto_rate or auto_rate_subset object
 #'
[…]

### 2022-03-24 R/auto_rate.R
trying auto_rate example again to pass old-R comd check

@@ -194,24 +194,24 @@
 #' \donttest{
 #' # Most linear section of an entire dataset
 #' inspect(sardine.rd, time = 1, oxygen =2) %>%
-#'  auto_rate()
+#'   auto_rate()
 #'
 #' # What is the lowest oxygen consumption rate over a 10 minute (600s) period?
 #' inspect(sardine.rd, time = 1, oxygen =2) %>%
-#'  auto_rate(method = "lowest", width = 600, by = "time") %>%
-#'  summary()
+#'   auto_rate(method = "lowest", width = 600, by = "time") %>%
+#'   summary()
 #'
 #' # What is the highest oxygen consumption rate over a 10 minute (600s) period?
 #' inspect(sardine.rd, time = 1, oxygen =2) %>%
-#'  auto_rate(method = "highest", width = 600, by = "time") %>%
-#'  summary()
+#'   auto_rate(method = "highest", width = 600, by = "time") %>%
+#'   summary()
 #'
 #' # What is the NUMERICAL minimum oxygen consumption rate over a 5 minute (300s)
 #' # period in intermittent-flow respirometry data?
 #' # NOTE: because uptake rates are negative, this would actually be
 #' # the HIGHEST uptake rate.
 #' auto_rate(intermittent.rd, method = "minimum", width = 300, by = "time") %>%
-#'  summary()
+#'   summary()
 #'
 #' # What is the NUMERICAL maximum oxygen consumption rate over a 20 minute
 #' # (1200 rows) period in respirometry data in which oxygen is declining?

### 2022-06-22 R/adjust_rate.ft.R
calc_rate.int complete

@@ -52,15 +52,17 @@
 #' `summary()`, and `mean()`.
 #'
 #' - `print()`: prints a single result, by default the first adjusted rate.
-#' Others can be printed by passing the `pos` input. e.g. `print(x, pos = 2)`
+#' Others can be printed by passing the `pos` input. e.g. `print(x, pos = 2)`.
+#' See `help("print.adjust_rate.ft")`.
 #'
 #' - `summary()`: prints summary table of all results and metadata, or those
 #' specified by the `pos` input. e.g. `summary(x, pos = 1:5)`. The summary can
-#' be exported as a separate dataframe by passing `export = TRUE`.
+#' be exported as a separate dataframe by passing `export = TRUE`. See
+#' `help("summary.adjust_rate.ft")`.
 #'
 #' - `mean()`: calculates the mean of all adjusted rates, or those specified by
 #' the `pos` input. e.g. `mean(x, pos = 1:5)` The mean can be exported as a
-#' separate value by passing `export = TRUE`.
+#' separate value by passing `export = TRUE`. See `help("mean.adjust_rate.ft")`.
 #'
 #' ## More
 #'
@@ -229,6 +231,7 @@ adjust_rate.ft <- function(x, by) {
 #' @param x adjust_rate.ft object
 #' @param pos integer. Which result to print.
 #' @param ... Pass additional inputs
+#' @keywords internal
 #' @return Print to console. No returned value.
 #' @export
 print.adjust_rate.ft <- function(x, pos = 1, ...) {
@@ -254,6 +257,7 @@ print.adjust_rate.ft <- function(x, pos = 1, ...) {
 #' @param pos integer(s). Which summary row(s) to print.
 #' @param export logical. Export summary table as data frame.
 #' @param ... Pass additional inputs
+#' @keywords internal
 #' @return Print to console. No returned value.
 #' @export
 #' @importFrom data.table data.table
@@ -288,6 +292,7 @@ summary.adjust_rate.ft <- function(object, pos = NULL, export = FALSE, ...) {
 #' @param pos integer(s). Which result(s) to average.
 #' @param export logical. Export averaged values as single value.
 #' @param ... Pass additional inputs
+#' @keywords internal
 #' @return Print to console. No returned value.
 #' @export
 mean.adjust_rate.ft <- function(x, pos = NULL, export = FALSE, ...){
[…]

### 2022-06-22 R/calc_rate.bg.R

@@ -194,6 +194,7 @@ calc_rate.bg <- function(x, time = NULL, oxygen = NULL, plot = TRUE, ...) {
 #' Print calc_rate.bg objects
 #' @param x calc_rate.bg object
 #' @param ... Pass additional inputs
+#' @keywords internal
 #' @return Print to console. No returned value.
 #' @export
 print.calc_rate.bg <- function(x, ...) {
@@ -213,6 +214,7 @@ print.calc_rate.bg <- function(x, ...) {
 #' @param pos integer(s). Which summary row(s) to print.
 #' @param export logical. Export summary table as data frame.
 #' @param ... Pass additional inputs
+#' @keywords internal
 #' @return Print to console. No returned value.
 #' @export
 summary.calc_rate.bg <- function(object, pos = NULL, export = FALSE, ...) {
@@ -248,6 +250,7 @@ summary.calc_rate.bg <- function(object, pos = NULL, export = FALSE, ...) {
 #' @param quiet logical. Suppress console output.
 #' @param legend logical. Suppress labels and legends.
 #' @param ... Pass additional plotting parameters
+#' @keywords internal
 #' @return A plot. No returned value.
 #' @export
 plot.calc_rate.bg <- function(x, pos = NULL, quiet = FALSE, legend = TRUE, ...) {
@@ -321,6 +324,7 @@ plot.calc_rate.bg <- function(x, pos = NULL, quiet = FALSE, legend = TRUE, ...) 
 #' @param pos integer(s). Which result(s) to average.
 #' @param export logical. Export averaged values as single value.
 #' @param ... Pass additional inputs
+#' @keywords internal
 #' @return Print to console. No returned value.
 #' @export
 mean.calc_rate.bg <- function(x, pos = NULL, export = FALSE, ...){

### 2022-06-22 R/convert_DO.R

@@ -191,6 +191,7 @@ convert_DO <- function(x, from = NULL, to = NULL, S = NULL, t = NULL,
 #' Print convert_DO objects
 #' @param x convert_DO object
 #' @param ... Pass additional inputs
+#' @keywords internal
 #' @return Print to console. No returned value.
 #' @export
 print.convert_DO <- function(x, ...) {
@@ -216,6 +217,7 @@ print.convert_DO <- function(x, ...) {
 #' Summarise convert_DO objects
 #' @param object convert_DO object
 #' @param ... Pass additional inputs
+#' @keywords internal
 #' @return Print to console. No returned value.
 #' @export
 summary.convert_DO <- function(object, ...) {
@@ -225,6 +227,7 @@ summary.convert_DO <- function(object, ...) {
 #' Plot convert_DO objects
 #' @param x convert_DO object
 #' @param ... Pass additional plotting parameters
+#' @keywords internal
 #' @return A plot. No returned value.
 #' @export
 plot.convert_DO <- function(x, ...) {
@@ -237,6 +240,7 @@ plot.convert_DO <- function(x, ...) {
 #' @param pos integer(s). Which result(s) to average.
 #' @param export logical. Export averaged values as single value.
 #' @param ... Pass additional inputs
+#' @keywords internal
 #' @return Print to console. No returned value.
 #' @export
 mean.convert_DO <- function(x, pos = NULL, export = FALSE, ...){

### 2022-06-22 R/convert_rate.R

@@ -1,16 +1,16 @@
 #' Convert a unitless oxygen rate value to absolute, mass-specific or
 #' area-specific rate
 #'
-#' Converts a unitless rate derived from [`calc_rate()`], [`auto_rate()`],
-#' [`adjust_rate()`], or [`calc_rate.bg()`] into an absolute rate (i.e. whole
-#' chamber or whole specimen), or mass-specific rate (i.e. normalised by
-#' specimen mass), or area-specific rate (i.e. normalised by specimen surface
-#' area) in any common unit.
+#' Converts a unitless rate derived from [`calc_rate()`], [`calc_rate.int()`],
+#' [`auto_rate()`], [`adjust_rate()`], or [`calc_rate.bg()`] into an absolute
+#' rate (i.e. whole chamber or whole specimen), or mass-specific rate (i.e.
+#' normalised by specimen mass), or area-specific rate (i.e. normalised by
+#' specimen surface area) in any common unit.
 #'
 #' By default, `convert_rate` converts the primary `$rate` element from
-#' `calc_rate` and `auto_rate` objects, the `$rate.adjusted` from `adjust_rate`
-#' objects, and the `$rate.bg` from `calc_rate.bg` objects. Additionally, any
-#' numeric value or vector of rates can be input as `x`.
+#' `calc_rate`, `calc_rate.int`, and `auto_rate` objects, the `$rate.adjusted`
+#' from `adjust_rate` objects, and the `$rate.bg` from `calc_rate.bg` objects.
+#' Additionally, any numeric value or vector of rates can be input as `x`.
 #'
 #' ## Respirometer volume
 #'
@@ -73,7 +73,17 @@
 #' input. e.g. `summary(x, pos = 1:5)`. This can be exported as a separate data
 #' frame by passing `export = TRUE`. This will be the *full* summary table, not
 #' the one printed to the console, including all rate regression parameters, and
-#' data locations, adjustments if applied, units, and more.
+#' data locations, adjustments if applied, units, and more. The `$rank` column
+#' requires special notice depending on the type of experiment you have analysed
+#' or the function you used to determine the rates. If `calc_rate` was used, it
+#' is the order of rates as entered using `from` and `to` (if multiple rates
+#' were determined). For `auto_rate` it relates to the `method` input, for
+#' example it indicates kernel density ranking if the `linear` method was used,
[…]

### 2022-06-22 R/convert_rate.ft.R

@@ -371,6 +371,7 @@ convert_rate.ft <- function(x,
 #' @param x convert_rate.ft object
 #' @param pos integer. Which result to print.
 #' @param ... Pass additional inputs
+#' @keywords internal
 #' @return Print to console. No returned value.
 #' @export
 print.convert_rate.ft <- function(x, pos = NULL, ...) {
@@ -399,6 +400,7 @@ print.convert_rate.ft <- function(x, pos = NULL, ...) {
 #' @param pos integer(s). Which summary row(s) to print.
 #' @param export logical. Export summary table as data frame.
 #' @param ... Pass additional inputs
+#' @keywords internal
 #' @return Print to console. No returned value.
 #' @export
 summary.convert_rate.ft <- function(object, pos = NULL, export = FALSE, ...) {
@@ -433,6 +435,7 @@ summary.convert_rate.ft <- function(object, pos = NULL, export = FALSE, ...) {
 #' @param pos integer(s). Which result(s) to average.
 #' @param export logical. Export averaged values as single value.
 #' @param ... Pass additional inputs
+#' @keywords internal
 #' @return Print to console. No returned value.
 #' @export
 mean.convert_rate.ft <- function(x, pos = NULL, export = FALSE, ...){
@@ -467,6 +470,7 @@ mean.convert_rate.ft <- function(x, pos = NULL, export = FALSE, ...){
 #' Plot convert_rate.ft objects
 #' @param x convert_rate.ft object
 #' @param ... Pass additional plotting parameters
+#' @keywords internal
 #' @return A plot. No returned value.
 #' @export
 plot.convert_rate.ft <- function(x, ...) {

### 2022-06-22 R/inspect.ft.R

@@ -557,6 +557,7 @@ inspect.ft <- function(x, time = NULL, out.oxy = NULL, in.oxy = NULL,
 #' Print inspect.ft objects
 #' @param x inspect.ft object
 #' @param ... Pass additional inputs
+#' @keywords internal
 #' @return Print to console. No returned value.
 #' @export
 print.inspect.ft <- function(x, ...) {
@@ -639,6 +640,7 @@ print.inspect.ft <- function(x, ...) {
 #' Summarise inspect.ft objects
 #' @param object inspect.ft object
 #' @param ... Pass additional inputs
+#' @keywords internal
 #' @return Print to console. No returned value.
 #' @export
 summary.inspect.ft <- function(object, ...) {
@@ -648,6 +650,7 @@ summary.inspect.ft <- function(object, ...) {
 #' Average inspect.ft object rates
 #' @param x calc_rate.bg object
 #' @param ... Pass additional inputs
+#' @keywords internal
 #' @return Print to console. No returned value.
 #' @export
 mean.inspect.ft <- function(x, ...){
@@ -664,6 +667,7 @@ mean.inspect.ft <- function(x, ...){
 #' @param legend logical. Suppress labels and legends.
 #' @param rate.rev logical. Control direction of y-axis in delta oxygen plot.
 #' @param ... Pass additional plotting parameters
+#' @keywords internal
 #' @return A plot. No returned value.
 #' @export
 plot.inspect.ft <- function(x, width = NULL, pos = NULL, quiet = FALSE,

### 2022-09-26 R/select_rate.R
select_rate for subset_rate

@@ -1,0 +1,1163 @@
+#' @title Select rate results based on a range of criteria
+#'
+#' @description The functions in `respR` are powerful, but outputs can be large
+#'   and difficult to explore, especially when there are hundreds to thousands
+#'   of results, for example the output of `auto_rate` on large datasets, or the
+#'   outputs of `calc_rate.int` from long intermittent-flow experiments.
+#'
+#'   The `select_rate` function helps explore, reorder, and filter
+#'   `convert_rate` results according to various criteria. For example,
+#'   extracting only positive or negative rates, only the highest or lowest
+#'   rates, only those from certain data regions, and numerous other methods
+#'   that allow advanced filtering of results so the final selection of rates is
+#'   well-defined towards the research question of interest. This also allows
+#'   for highly consistent reporting of results and rate selection criteria.
+#'
+#'   Multiple selection criteria can be applied by saving the output and
+#'   processing it through the function multiple times using different methods,
+#'   or alternatively via piping (`|>` or `%>%`). See Examples.
+#'
+#'   *Note:* when choosing a `method`, keep in mind that to remain
+#'   mathematically consistent, `respR` outputs oxygen consumption (i.e.
+#'   respiration) rates as negative values. This is particularly important in
+#'   the difference between `highest/lowest` and `minimum/maximum` methods. See
+#'   Details.
+#'
+#'   When a rate result is omitted by the selection criteria, it is removed from
+#'   the `$rate.output` element of the `convert_rate` object, and all associated
+#'   data in `$summary` (i.e. the associated row) is removed. Some methods can
+#'   also be used with an `n = NULL` input to reorder the `$rate` and `$summary`
+#'   elements in various ways.
+#'
+#'   ## Rank
+#'
+#'   The summary table `$rank` column is context-specific, and what it
+#'   represents depends on the type of experiment analysed or the function used
+#'   to determine the rates. If numeric values were converted, it is the order
+#'   in which they were entered. Similarly, if `calc_rate` was used, it is the
+#'   order of rates as entered using `from` and `to` (if multiple rates were
+#'   determined). For `auto_rate` it relates to the `method` input. For example
+#'   it indicates the kernel density ranking if the `linear` method was used,
+#'   the ascending or descending ordering by absolute rate value if `lowest` or
+#'   `highest` were used, or the numerical order if `minimum` or `maximum` were
[…]

### 2022-12-06 R/auto_rate.R
intermittent-flow fns dones

@@ -69,17 +69,10 @@
 #' specified `width` are extracted from the rolling regressions, ordered by
 #' time.
 #'
-#' - NOTE: `max`, `min`: These methods were used in previous versions of `respR`
-#' but have been deprecated. They were intended to order oxygen uptake
-#' (negative) rates by magnitude, but this resulted in incorrect ordering of
-#' oxygen production (positive) rates. They have been retained for code
-#' compatibility, but will be removed in a future version of `respR`, and so
-#' *should not be used*.
-#'
 #' ## Further selection and filtering of results
 #'
 #' For further selection or subsetting of `auto_rate` results, see the dedicated
-#' [`subset_rate()`] function, which allows subsetting of rates by various
+#' [`select_rate()`] function, which allows subsetting of rates by various
 #' criteria, including r-squared, data region, percentiles, and more.
 #'
 #' ## Units
@@ -91,13 +84,13 @@
 #' ## The `width` and `by` inputs
 #'
 #' If `by = "time"`, the `width` input represents a time window in the units of
-#' the time data.
+#' the time data in `x`.
 #'
-#' If `by = "row"` and between 0 and 1, `width` represents a proportion of the
-#' total data length, as in the equation `floor(width * number of data rows)`.
-#' For example, 0.2 represents a rolling window of 20% of the data width.
-#' Otherwise, if entered as an integer of 2 or greater, the `width` represents
-#' the number of rows.
+#' If `by = "row"` and `width` is between 0 and 1 it represents a proportion of
+#' the total data length, as in the equation `floor(width * number of data
+#' rows)`. For example, 0.2 represents a rolling window of 20% of the data
+#' width. Otherwise, if entered as an integer of 2 or greater, the `width`
+#' represents the number of rows.
 #'
 #' For both `by` inputs, if left as `width = NULL` it defaults to 0.2 or a
 #' window of 20% of the data length.
@@ -230,8 +223,12 @@
 #'   summary()
 #'  }
 
-auto_rate <- function(x, method = "linear", width = NULL,
-                      by = "row", plot = TRUE, ...) {
+auto_rate <- function(x,
+                      method = "linear",
+                      width = NULL,
+                      by = "row",
+                      plot = TRUE,
+                      ...) {
 
   ## Save function call for output
   call <- match.call()
@@ -245,6 +242,7 @@ auto_rate <- function(x, method = "linear", width = NULL,
   # perform checks
   checks <- validate_auto_rate(x, by, method)
   dt <- checks$df  # extract df from validation check
[…]

### 2022-12-06 R/calc_rate.ft.R

@@ -484,8 +484,11 @@ calc_rate.ft <- function(x = NULL, flowrate = NULL, from = NULL, to = NULL,
               subsets = dfs,
               delta.oxy = delta,
               input_type = xtype,
-              summary = data.table(cbind(rank = 1:nrow(summary),
-                                         summary, flowrate = flowrate, rate = rate)),
+              summary = data.table(cbind(rep = NA,
+                                         rank = 1:nrow(summary),
+                                         summary,
+                                         flowrate = flowrate,
+                                         rate = rate)),
               rate = rate)
 
   class(out) <- "calc_rate.ft"

### 2022-12-06 R/inspect.R

@@ -573,36 +573,19 @@ plot.inspect <- function(x, width = NULL, pos = NULL, quiet = FALSE,
     mtext("Full Timeseries",
           outer = TRUE, cex = 1.2, line = 0.3, font = 2)
 
-    ## Adding this fn here to avoid using static_roll
-    roll_reg_plot <- function(df, width) {
-      roll_width <- floor(width * nrow(df))
-      ## Calc all rates, even there is a min_obs of only 1 datapoint
-      ## This means rate is returned even if there are NA in data
-      ## Also replace Inf with NA of ylim in plot fails
-      df[[1]][which(is.infinite(df[[1]]))] <- NA
-      df[[2]][which(is.infinite(df[[2]]))] <- NA
-      rates <- roll::roll_lm(matrix(df[[1]]), matrix(df[[2]]),
-                             roll_width, min_obs = 1)$coefficients[,2]
-      ## However this means rates are ALSO calculated at the start of the data
-      ## before the width is even reached, so we remove these.
-      rates <- rates[-(1:(roll_width-1))]
-      rates <- na.omit(rates)
-      return(rates)
-    }
-
     ## Rolling reg plot
     ## Width needs to be half on each side
-    rates <- roll_reg_plot(dt, width)
+    rates <- rollreg2.p(dt, width)
     half_width <- width/2
     xdt <- dt[[1]]
     xlim <- range(nainf.omit(xdt))
 
-    ## now to get ylim, if width is higher than 0.1 we run roll_reg_plot with
+    ## now to get ylim, if width is higher than 0.1 we run rollreg2.p with
     ## 0.1 width to get y range of rates, and plot against these. This means if
     ## a higher width is used, and rates have very little variation the plots
     ## don't look weird - i.e. should get flatter with higher widths
     if(width > 0.1) {
-      rates01 <- roll_reg_plot(dt, 0.1)
+      rates01 <- rollreg2.p(dt, 0.1)
       y_lim <- range(nainf.omit(rates01))
       buffer <- diff(y_lim)*0.05
       y_lim <- c(y_lim[1] - buffer, y_lim[2] + buffer) ## add a little more space

### 2022-12-06 R/select_rate.R

@@ -24,12 +24,12 @@
 #'   Details.
 #'
 #'   When a rate result is omitted by the selection criteria, it is removed from
-#'   the `$rate.output` element of the `convert_rate` object, and all associated
-#'   data in `$summary` (i.e. the associated row) is removed. Some methods can
-#'   also be used with an `n = NULL` input to reorder the `$rate` and `$summary`
+#'   the `$rate.output` element of the `convert_rate` object, and the associated
+#'   data in `$summary` (i.e. that row) is removed. Some methods can also be
+#'   used with an `n = NULL` input to reorder the `$rate` and `$summary`
 #'   elements in various ways.
 #'
-#'   ## Rank
+#'   ## Replicate and Rank columns
 #'
 #'   The summary table `$rank` column is context-specific, and what it
 #'   represents depends on the type of experiment analysed or the function used
@@ -40,26 +40,37 @@
 #'   it indicates the kernel density ranking if the `linear` method was used,
 #'   the ascending or descending ordering by absolute rate value if `lowest` or
 #'   `highest` were used, or the numerical order if `minimum` or `maximum` were
-#'   used. For intermittent-flow experiments analysed via `calc_rate.int` it
-#'   indicates the replicate number. The `$rank` column can be used to keep
-#'   track of selection or reordering because the original rank value will be
+#'   used. For intermittent-flow experiments analysed via `calc_rate.int` and
+#'   `auto_rate.int` these will be ranked *within* each replicate as indicated
+#'   in the `$rep` column. The `$rep` and `$rank` columns can be used to keep
+#'   track of selection or reordering because the original values will be
 #'   retained unchanged through selection or reordering operations. The original
-#'   rank order can always be restored by using `method = "rank"` and `n =
-#'   NULL`.
+#'   order can always be restored by using `method = "rep"` or `method = "rank"`
+#'   with `n = NULL`. In both these cases the `$summary` table and
+#'   `$rate.output` will be reordered by `$rep` (if used) then `$rank` to
+#'   restore the original ordering.
+#'
+#'   Note that if you are analysing intermittent-flow data and used
+#'   `auto_rate.int` but changed the `n` input to output more than one rate
+#'   result per replicate, the selection or reordering operations will not take
+#'   any account of this. You should carefully consider if or why you need to
+#'   output multiple rates per replicate in the first place. If you have, you
[…]

### 2022-12-06 R/subset_data.R

@@ -9,24 +9,21 @@
 #' passed to an analytical function such as `calc_rate` or `auto_rate`. See
 #' examples and vignettes.
 #'
-#' The function can subset data based on ranges of `"time"`, `"oxygen"`, `"row"`
-#' , or `"proportion"` of total oxygen used or produced (note, this last option
-#' works poorly with noisy or fluctuating data). For data frames, to subset by
-#' `"time"`, `"oxygen"`, or `"proportion"`, the time data is assumed to be in
-#' the first column, and oxygen data in the second column. For [`inspect()`] and
-#' [`inspect.ft()`] objects, the data will have been coerced to this structure
-#' already. In these cases the `$dataframe` element in the output is replaced by
-#' the subset, and in `inspect.ft` the `$data` element is also subset and
-#' replaced. Note for `inspect.ft` objects, the oxygen data in column 2 will be
-#' either `out.oxy` data or `delta.oxy` data depending on what was inspected.
-#' The function can subset *any* data frame by `row`.
+#' The function can subset data based on ranges of `"time"`, `"oxygen"`, or
+#' `"row"`. For data frames, to subset by `"time"` or `"oxygen"` the time data
+#' is assumed to be in the first column, and oxygen data in the second column.
+#' For [`inspect()`] and [`inspect.ft()`] objects, the data will have been
+#' coerced to this structure already. In these cases the `$dataframe` element in
+#' the output is replaced by the subset, and in `inspect.ft` the `$data` element
+#' is also subset and replaced. Note for `inspect.ft` objects, the oxygen data
+#' in column 2 will be either `out.oxy` data or `delta.oxy` data depending on
+#' what was inspected. The function can subset *any* data frame by `row`.
 #'
 #' When multiple columns are present, for example time in column 1, and multiple
 #' columns of oxygen data, the subset object will include *all* columns. In the
-#' case of subsetting `by = "oxygen"` or `by = "proportion"`, subsetting is
-#' based on the *first* column of oxygen data only (i.e. column 2), and all
-#' subsequent columns are subset between the same rows regardless of oxygen
-#' values.
+#' case of subsetting `by = "oxygen"`, subsetting is based on the *first* column
+#' of oxygen data only (i.e. column 2), and all subsequent columns are subset
+#' between the same rows regardless of oxygen values.
[…]

### 2022-12-06 R/util_plots.R

@@ -220,29 +220,67 @@ rollreg.p <- function(rolldf, ranked.b1, rownums, xlim, rate.rev = TRUE, ...) {
   title(main = ("Rolling Rate"), line = 1.2, font = 2)
 }
 
+# rolling regression 2
+# this is a version especially for inspect roll reg plot
+# Can't quite remember why i didn't use the above, but there was a reason...
+# Maybe just better control over appearance...
+rollreg2.p <- function(df, width) {
+  roll_width <- floor(width * nrow(df))
+  ## Calc all rates, even there is a min_obs of only 1 datapoint
+  ## This means rate is returned even if there are NA in data
+  ## Also replace Inf with NA of ylim in plot fails
+  df[[1]][which(is.infinite(df[[1]]))] <- NA
+  df[[2]][which(is.infinite(df[[2]]))] <- NA
+  rates <- roll::roll_lm(matrix(df[[1]]), matrix(df[[2]]),
+                         roll_width, min_obs = 1)$coefficients[,2]
+  ## However this means rates are ALSO calculated at the start of the data
+  ## before the width is even reached, so we remove these.
+  rates <- rates[-(1:(roll_width-1))]
+  rates <- na.omit(rates)
+  return(rates)
+}
 
-#' Plots multiple auto_rate results in a nice way
-#' using base plot
-#' x = auto_rate or auto_rate_subset object
-#'
-#' @keywords internal
-plot_ar_grid <- function(x, pos = NULL, msg = "subset_rate",
-                         title = "Subset", max_msg = "", ...){
+# Plots multiple auto_rate or convert_rate results in a nice way
+# using base plot
+# x = auto_rate or convert_rate object
+grid.p <- function(x, pos = NULL, msg = "grid.p",
+                   title = "Rank", quiet = FALSE, ...){
 
   parorig <- par(no.readonly = TRUE) # save original par settings
   on.exit(par(parorig)) # revert par settings to original
 
-  totres <- length(x$rate)
+  ## warning if empty - but return to allow piping
+  if(length(x$summary$rate.output) == 0){
+    message(glue::glue("{msg}: Nothing to plot! No rates found in object."))
+    return(invisible(x))
+  }
+
+  # total number of results
+  totres <- nrow(x$summary)
+  # set pos
   if(is.null(pos)) pos <- 1:totres
-  if(any(pos > totres))
[…]

### 2022-12-06 tests/testthat/test-calc_rate.R

@@ -41,14 +41,6 @@ capture.output({  ## stops printing outputs on assigning
     expect_equal(as.numeric(cr$subsets[[1]][nrow(cr$subsets[[1]]),2]),
                  93)
 
-
-    cr <- calc_rate(sard_data, 0.8, 0.5, by = "proportion", plot = F)
-    expect_is(cr,
-              "calc_rate")
-    expect_equal(as.numeric(cr$subsets[[1]][1,2]),
-                 min(cr$dataframe[[2]]) + (diff((c(min(cr$dataframe[[2]]), max(cr$dataframe[[2]])))) * 0.8))
-    expect_equal(as.numeric(cr$subsets[[1]][nrow(cr$subsets[[1]]),2]),
-                 (min(cr$dataframe[[2]]) + max(cr$dataframe[[2]]))/2)
   })
 
   test_that("calc_rate - S3 generics work", {
@@ -104,10 +96,6 @@ capture.output({  ## stops printing outputs on assigning
     expect_error(calc_rate(sard_data, plot = F, by = "oxygen"), regexp = NA)
     expect_error(calc_rate(sard_data, plot = F, by = "Row"), regexp = NA)
     expect_error(calc_rate(sard_data, plot = F, by = "r"), regexp = NA)
-    expect_error(calc_rate(sard_data, plot = F, by = "Proportion"),
-                 regexp = "calc_rate: please enter a proportion 'from' input.")
-    expect_error(calc_rate(sard_data, plot = F, by = "prop"),
-                 regexp = "calc_rate: please enter a proportion 'from' input.")
   })
 
   test_that("calc_rate - stops with wrong 'by' inputs", {
@@ -128,8 +116,8 @@ capture.output({  ## stops printing outputs on assigning
                  "calc_rate: Input must be a 'data.frame' or 'inspect' object.")
   })
 
-  test_that("calc_rate - warns with multi column input", {
-    expect_warning(calc_rate(urchins.rd, plot = F),
+  test_that("calc_rate - message with multi column input", {
+    expect_message(calc_rate(urchins.rd, plot = F),
                    "calc_rate: Multi-column dataset detected in input.")
   })
 
@@ -159,8 +147,6 @@ capture.output({  ## stops printing outputs on assigning
     expect_equal(calc_rate(urch, from = NULL, to = 7, by = "oxygen", plot = FALSE)$summary$oxy,
                  urch[[2]][1])
 
[…]

### 2022-12-06 tests/testthat/test-convert_rate.ft.R

@@ -461,11 +461,11 @@ test_that("convert_rate.ft - objects work with summary().", {
   })
   sapply(objs, function(z) {
     expect_output(summary(eval(parse(text=z))),
-                  regexp = "rank rate.input")
+                  regexp = "rep rank intercept_b0")
   })
 
 })
-
+summary(conv.crft1)
 test_that("convert_rate.ft - objects work with summary() and 'pos' input", {
 
   expect_error(summary(conv.crftmany.ms, pos = 2),

### 2022-12-13 R/select_rate.R
minor docs updates

@@ -167,7 +167,7 @@
 #'   to rows of the raw data, **not** rows of the summary table - see `manual`
 #'   method for this. For all of these methods, if `n = NULL` no results will be
 #'   removed, instead the results will be reordered by that respective column
-#'   (descending for `rsq` and `density`, ascending for `row`, and `time`).
+#'   (descending for `rsq` and `density`, ascending for `row` and `time`).
 #'
 #'   ## `time_omit`, `row_omit`
 #'

### 2023-01-05 R/adjust_rate.ft.R
v2.2. Flowthrough updates etc,

@@ -208,6 +208,11 @@ adjust_rate.ft <- function(x, by) {
     out <- list(call = call,
                 inputs = list(x = x,
                               by = by),
+                dataframe = NULL,
+                data = NULL,
+                subsets = NULL,
+                delta.oxy = NULL,
+                input_type = NULL,
                 summary = data.table::data.table(rank = 1:length(rate.adjusted),
                                                  rate = rate,
                                                  adjustment = adjustment,
@@ -279,7 +284,7 @@ summary.adjust_rate.ft <- function(object, pos = NULL, export = FALSE, ...) {
 
   out <- object$summary[pos,]
 
-  print(out, class = FALSE)
+  print(out, nrows = 50, class = FALSE)
   cat("-----------------------------------------\n")
 
   if(export)

### 2023-01-05 R/calc_rate.int.R

@@ -504,7 +504,7 @@ summary.calc_rate.int <- function(object, pos = NULL, export = FALSE, ...) {
   }
 
   out <- object$summary[pos,]
-  print(out, class = FALSE)
+  print(out, nrows = 50, class = FALSE)
   cat("-----------------------------------------\n")
 
   if(export)

### 2023-01-05 R/import_file.R

@@ -179,9 +179,9 @@ import_file <- function(path, export = FALSE) {
   } else if (suppressWarnings(any(grepl("Fractional error", raw[1:20])))) {
     cat("Loligo AutoResp metadata file detected\n")
     stop("Currently these files are unsupported in respR.
-  This is an AutoResp metadata file, and contains no raw time~O2 data.
+  This is an AutoResp metadata file, and contains no raw time~oxygen data.
   It may contain other experimentally useful values (e.g. mass and volume).
-  Please import the associated file appended with \"_raw\" which contains time~O2 data." )
+  Please import the associated file appended with \"_raw\" which contains time~oxygen data." )
     ## PreSens Datamanager
   } else if (suppressWarnings(any(grepl("PreSens Datamanager", raw[1:20])))) {
     cat("PreSens Datamanager output file detected\n")

### 2023-01-05 tests/testthat/test-auto_rate.int.R

@@ -2,8 +2,8 @@
 # library(testthat)
 # test_file("tests/testthat/test-auto_rate.int.R")
 # covr::file_coverage("R/auto_rate.int.R", "tests/testthat/test-auto_rate.int.R")
-# x <- covr::package_coverage()
-# covr::report(x)
+# cvr <- covr::package_coverage()
+# covr::report(cvr)
 # covr::report(covr::package_coverage())
 
 capture.output({  ## stops printing outputs on assigning

### 2023-01-05 tests/testthat/test-subset_data.R

@@ -2,8 +2,8 @@
 # library(testthat)
 # test_file("tests/testthat/test-subset_data.R")
 # covr::file_coverage("R/subset_data.R", "tests/testthat/test-subset_data.R")
-# x <- covr::package_coverage()
-# covr::report(x)
+# cvr <- covr::package_coverage()
+# covr::report(cvr)
 
 capture.output({  ## stops printing outputs on assigning
 

### 2023-05-04 R/auto_rate_funs.R
convert_MR and much more

@@ -13,16 +13,13 @@ validate_auto_rate <- function(x, by, method) {
 
   # x validation
   if (!(any(class(x) %in% c("inspect", "data.frame"))))
-    stop("auto_rate: Input data must be of class 'data.frame' or 'inspect'")
+    stop("auto_rate: Input data must be of class 'data.frame' or 'inspect'", call. = FALSE)
 
   # validate by
-  by <- verify_by(by,  which = c("t", "r"), msg = "auto_rate:") # validate `by` input
+  by <- by.val(by,  which = c("t", "r"), msg = "auto_rate") # validate `by` input
 
   # validate method
-  if (!(method %in% c("linear", "max", "min", "interval",
-                      "rolling", "highest", "lowest", "maximum", "minimum")))
-    stop("auto_rate: The 'method' input is not recognised: it should be 'linear',
-    'highest', 'lowest', 'maximum', 'minimum', 'rolling', or 'interval'")
+  method.val(method, "auto_rate")
 
   # extract df
   if (any(class(x) %in% "inspect")) df <- x$dataframe else
@@ -30,7 +27,7 @@ validate_auto_rate <- function(x, by, method) {
 
   # select only first two columns by default if dataset is multi-column
   if (length(df) > 2) {
-    warning("auto_rate: Multi-column dataset detected in input. Selecting first two columns by default.\n  If these are not the intended data, inspect() or subset the data frame columns appropriately before running auto_rate()")
+    warning("auto_rate: Multi-column dataset detected in input. Selecting first two columns by default.\n  If these are not the intended data, inspect() or subset the data frame columns appropriately before running auto_rate()", call. = FALSE)
     df <- df[, 1:2]
   }
 
@@ -124,7 +121,7 @@ auto_rate_highest <- function(dt, width, by = 'row') {
 
   # stop if mix of -ve and +ve
   if(any(rollreg$slope_b1 > 0) && any(rollreg$slope_b1 < 0))
-    stop("auto_rate: Analysis produces both negative and positive rates. \n The 'highest' method is intended to order by the lowest *absolute* rate amongst rates all having the same sign.\n Use 'maximum' or 'minimum' method to order rates by *numerical* value.")
[…]

### 2023-05-04 R/calc_rate.bg.R

@@ -125,12 +125,10 @@ calc_rate.bg <- function(x, time = NULL, oxygen = NULL, plot = TRUE, ...) {
                  oxygen = oxygen,
                  plot = plot)
 
-  # Import x from inspect function. We convert to data.frame object here as
+  # Import x from inspect function. We convert to data.frame here as
   # data.table doesn't like subsetting columns by variable names.
-  if(any(class(x) %in% "inspect")) {
-    x <- data.frame(x$dataframe)
-  } else {
-    x <- data.frame(x)}
+  if(any(class(x) %in% "inspect")) df <- data.frame(x$dataframe) else
+    df <- data.frame(x)
 
   ## if NULL use col1 for time, all other cols for oxygen
   if(is.null(time)) {
@@ -138,17 +136,20 @@ calc_rate.bg <- function(x, time = NULL, oxygen = NULL, plot = TRUE, ...) {
     message("calc_rate.bg: Applying column default of 'time = 1'")
   }
   if(is.null(oxygen)) {
-    listcols <- seq.int(1, ncol(x))
+    listcols <- seq.int(1, ncol(df))
     oxygen <- listcols[!listcols %in% time]
     message("calc_rate.bg: Applying column default of all non-time column(s) as 'oxygen'")
   }
-  if (any(time %in% oxygen)) stop("calc_rate.bg: 'time' and 'oxygen' columns conflict.")
+
+  # validate inputs
+  column.val(input=time, req = TRUE, min = 1, max = 1,
+             range = c(1,ncol(df)), conflicts = oxygen, msg = "calc_rate.bg: 'time' input -")
+  column.val(oxygen, req = TRUE, min = 1, max = ncol(df)-1,
+             range = c(1,ncol(df)), conflicts = time, msg = "calc_rate.bg: 'oxygen' input -")
 
   # Extract data:
-  if(any(time > length(x))) stop("calc_rate.bg: Selected 'time' column not present in the input.") else
-    xval <- x[time]
-  if(any(oxygen > length(x))) stop("calc_rate.bg: Selected 'oxygen' column(s) not present in the input.") else
-    yval <- x[oxygen]
+  xval <- df[time]
+  yval <- df[oxygen]
 
   # Ok, convert back to data.table object
   dt <- data.table(xval, yval)
@@ -221,7 +222,7 @@ print.calc_rate.bg <- function(x, ...) {
[…]

### 2023-05-04 R/convert_mr.R

@@ -1,128 +1,420 @@
-# convert_DO(100, "%Air", "cm3/l", S = 35, t = 25, P = 1.013253)
-# convert_DO(100, "%Air", "ml/l", S = 35, t = 25, P = 1.013253)
-# convert_DO(10, "cm^3/l", "%Air", S = 35, t = 25, P = 1.013253)
-#
-#
-#
-#
-# # -------------------------------------------------------------------------
-#
-# # Oxygen
-# # ⁠"ug" "mg" "pmol" "nmol" "umol" "mmol" "mol" "mL"⁠
-# # add cm3 or cc = ml
-# # mm3 = ul?
-# # pressure?
-#
-# # Time
-# # ⁠"sec" "min" "hour" "day"⁠
-# # Mass
-# # ⁠"ug" "mg" "g" "kg"⁠
-# # Area
-# # ⁠"mm2" "cm2" "m2" "km2"⁠
-#
-#
-#
-#
-# rate <- 0.5
-# in.unit <- "mg/s"
-# out.unit <- "mg/day"
-# S = 35
-# t = 15
-# P = 1.01
-#
-# rate <- 0.5
-# in.unit <- "ml/h/g"
-# out.unit <- "mg/day/kg"
-# S = 35
-# t = 15
-# P = 1.01
-#
-# rate <- 0.5
-# in.unit <- "ml/h/mm2"
-# out.unit <- "mg/day/m2"
-# S = 35
-# t = 15
-# P = 1.01
-#
-#
-# in.unit <- "mg/h/g"
-# in.unit <- "ml/h"
-# in.unit <- "ml/h/g"
-#
-#
-#
-# # separate the input unit
-# in.uns <- as.matrix(read.table(text = gsub("(?:-1|[/.[:space:]])+",
-#                                            " ", in.unit), header = FALSE))
-# # input unit types
-# in.uns.type <- sapply(in.uns, function(x) respR:::unit_type(x))
-#
-# # unit type
-# if(length(in.uns.type) == 2) un.type <- "abs" else
-#   if(in.uns.type[3] == "mass") un.type <- "mass.spec" else
-#     if(in.uns.type[3] == "area") un.type <- "area.spec" else
-#       stop("convert_mr: unit error")
-#
-# if(un.type == "abs"){
-#   # parse time unit
-#   tm.unit <- respR:::verify_units(in.uns[2], "time")
-#   # adjust rate to per hour
[…]

### 2023-05-04 R/select_rate.R

@@ -983,15 +983,22 @@ select_rate <- function(x, method = NULL, n = NULL){
       keep <- c(keep1, keep2)
       # otherwise do this possibly very slow loop
     } else {
-      message("select_rate: For 'row_omit' method, selecting multiple 'n' inputs can be VERY SLOW. \nIf possible, specifying a range using lower and upper values is much faster.")
-      remove <- c() # empty obj for loop results
+      message("select_rate: For 'row_omit' method using multiple 'n' inputs can be VERY SLOW. A range of lower and upper values is MUCH faster.")
+        # progress bar object
+        progress <- txtProgressBar(min = 0,               # Minimum value of the progress bar
+                             max = length(n_order), # Maximum value of the progress bar
+                             style = 3,    # Progress bar style (also available style = 1 and style = 2)
+                             width = NA,   # Progress bar width. Defaults to getOption("width")
+                             char = "=")
       # for each n, go through summary table by row
       # if it occurs between row and endrow then that row of summary gets removed
+      remove <- c() # empty obj for loop results
       for(i in n_order){
         remove_i <- which(apply(x$summary, 1, function(z) any(i %in% z[7]:z[8])))
         remove <- sort(unique(c(remove, remove_i)))
-        progress.bar(which(i == n_order), length(n_order), "select_rate: 'row_omit' progress")
+        setTxtProgressBar(progress, which(i == n_order))
       }
+      close(progress)
       cat("\n")
       # if nothing to remove, above outputs integer(0), so this is for that...
       if(length(remove) > 0) keep <- (1:nrow(x$summary))[-remove] else
@@ -1045,14 +1052,20 @@ select_rate <- function(x, method = NULL, n = NULL){
     } else {
       message("select_rate: For 'time_omit' method, selecting multiple 'n' inputs can be VERY SLOW. \nIf possible, specifying a range using lower and upper values is much faster.")
 
+      # progress bar object
+      progress <- txtProgressBar(min = 0,               # Minimum value of the progress bar
+                           max = length(n_order), # Maximum value of the progress bar
+                           style = 3,    # Progress bar style (also available style = 1 and style = 2)
+                           width = NA,   # Progress bar width. Defaults to getOption("width")
+                           char = "=")
       remove <- c() # empty obj for loop results
-
       for(i in n_order){
         remove_i <- which(apply(x$summary, 1, function(z)
[…]

### 2023-05-04 R/util_funs.R

@@ -1,12 +1,123 @@
-# General global default settings ------------------------------------------------
+# General package default settings ------------------------------------------------
 
 # Number of rows of summary tables to print before it prints a condensed version
 # i.e. nrows in print.data.table
 summ_rows <- 50
 
+# Regex -------------------------------------------------------------------
+
+# Regex patterns for units
+# Put here so don't have to update them in multiple places
+# Used in units.val and unit_type and unit_type_o1 etc.
+
+# Allowed unit separators - used in splitting units up
+unit.sep.rgx  <- "(?:-1|[_/.[:space:]]|per)+"
+# time
+min.time.rgx  <- "^(?i)\\b(minute|min|m)(s)?(.time)?\\b$"
+sec.time.rgx  <- "^(?i)\\b(second|sec|s)(s)?(.time)?\\b$"
+hr.time.rgx   <- "^(?i)\\b(hour|hr|h)(s)?(.time)?\\b$"
+day.time.rgx  <- "^(?i)\\b(day|dy|d)(s)?(.time)?\\b$"
+# o2
+# DO NOT require S,t,P
+mgperL.o2.rgx    <- "^(?i)\\b(mg|milligram|milligramme)(.o2|O2)?(/|per|.|_)?(l|liter|litre)(-1)?(.o2)?\\b$"
+ugperL.o2.rgx    <- "^(?i)\\b(ug|microgram|microgramme)(.o2|O2)?(/|per|.|_)?(l|liter|litre)(-1)?(.o2)?\\b$"
+molperL.o2.rgx   <- "^(?i)\\b(mol|mole)(.o2|O2)?(/|per|.|_)?(l|liter|litre)(-1)?(.o2)?\\b$"
+mmolperL.o2.rgx  <- "^(?i)\\b(mmol|mmole|millimol|millimole)(.o2|O2)?(/|per|.|_)?(l|liter|litre)(-1)?(.o2)?\\b$"
[…]

### 2023-05-04 tests/testthat/test-auto_rate.R

@@ -34,7 +34,12 @@ capture.output({  ## stops printing outputs on assigning
 
   test_that("auto_rate stops with malformed method", {
     expect_error(auto_rate(urch_data, method = "wrong"),
-                 regexp = "auto_rate: The 'method' input is not recognised: ")
+                 regexp = "auto_rate: 'method' input not recognised.")
+  })
+
+  test_that("calc_rate - stops when 'x' input has only 1 row of data", {
+    expect_error(auto_rate(sardine.rd[1,1:2], plot = F),
+                 "auto_rate: Input data contains only 1 row. Please check inputs.")
   })
 
   test_that("auto_rate works using default inputs", {
@@ -141,6 +146,13 @@ capture.output({  ## stops printing outputs on assigning
     ar <- auto_rate(urch_data, method = "interval", plot = F, width = 0.1)
     expect_output(print(ar))
     expect_output(summary(ar))
+  })
+
+  test_that("auto_rate summary works with print.kds", {
+    ar <- auto_rate(urch_data, method = "linear", plot = F, width = 0.1)
+    expect_output(summary(ar, print.kds = TRUE))
+    expect_output(summary(ar, print.kds = TRUE),
+                  "=== Kernel Density Summary ===")
   })
 
   test_that("auto_rate summary export = TRUE works", {
@@ -337,7 +349,7 @@ capture.output({  ## stops printing outputs on assigning
     expect_message(suppressWarnings(auto_rate(intermittent.rd, method = "max", plot = FALSE)),
                    regexp = "auto_rate: Note dataset contains both negative and positive rates.")
     expect_warning(auto_rate(intermittent.rd, method = "max", plot = FALSE),
-                   regexp = "auto_rate: the 'min' and 'max' methods have been deprecated")
+                   regexp = "auto_rate: The 'min' and 'max' methods have been deprecated")
     ## INCORRECTLY orders by numerical max rate
     ## this was the original behaviour
     expect_equal(suppressWarnings(auto_rate(intermittent.rd, method = "max", plot = FALSE)$rate),
@@ -357,7 +369,7 @@ capture.output({  ## stops printing outputs on assigning
     expect_message(suppressWarnings(auto_rate(intermittent.rd, method = "min", plot = FALSE)),
                    regexp = "auto_rate: Note dataset contains both negative and positive rates.")
     expect_warning(auto_rate(intermittent.rd, method = "min", plot = FALSE),
[…]

### 2023-05-04 tests/testthat/test-calc_rate.bg.R

@@ -92,4 +92,4 @@ test_that("calc_rate.bg - plot defaults are correctly restored", {
 })
 
 
-}) ## turns printing back on
+}) ## end capture.output

### 2023-05-04 tests/testthat/test-calc_rate.ft.R

@@ -91,7 +91,7 @@ test_that("calc_rate.ft - accepts single numeric 'x' input", {
   expect_error(calc_rate.ft(0.8, flowrate = 1.5, plot = F),
                regexp = NA)
   expect_message(calc_rate.ft(0.8, flowrate = 1.5, plot = F),
-                 regexp = "calc_rate.ft: calculating rate from delta oxygen.")
+                 regexp = "calc_rate.ft: Calculating rate from delta oxygen.")
 })
 
 test_that("calc_rate.ft - expected output with single numeric 'x' input", {
@@ -111,25 +111,25 @@ test_that("calc_rate.ft - delta output is correct with single numeric 'x' input"
 test_that("calc_rate.ft - single numeric 'x' input, message that other inputs ignored.", {
   expect_message(calc_rate.ft(-0.8, flowrate = 1.5, from = 10,
                               plot = F),
-                 regexp = "calc_rate.ft: requires only 'x' and 'flowrate'. Additional inputs ignored.")
+                 regexp = "calc_rate.ft: Requires only 'x' and 'flowrate'. Additional inputs ignored.")
   expect_message(calc_rate.ft(-0.8, flowrate = 1.5, to = 10,
                               plot = F),
-                 regexp = "calc_rate.ft: requires only 'x' and 'flowrate'. Additional inputs ignored.")
+                 regexp = "calc_rate.ft: Requires only 'x' and 'flowrate'. Additional inputs ignored.")
   expect_message(calc_rate.ft(-0.8, flowrate = 1.5, by = "row",
                               plot = F),
-                 regexp = "calc_rate.ft: requires only 'x' and 'flowrate'. Additional inputs ignored.")
+                 regexp = "calc_rate.ft: Requires only 'x' and 'flowrate'. Additional inputs ignored.")
   expect_message(calc_rate.ft(-0.8, flowrate = 1.5, width = 10,
                               plot = F),
-                 regexp = "calc_rate.ft: requires only 'x' and 'flowrate'. Additional inputs ignored.")
+                 regexp = "calc_rate.ft: Requires only 'x' and 'flowrate'. Additional inputs ignored.")
   expect_message(calc_rate.ft(-0.8, flowrate = 1.5, width = 10, by = "row", from = 2,
                               plot = F),
-                 regexp = "calc_rate.ft: requires only 'x' and 'flowrate'. Additional inputs ignored.")
[…]

### 2023-05-04 tests/testthat/test-convert_val.R

@@ -18,7 +18,7 @@ test_that("convert_val - stops if units not recognised, or not of accepted type"
   expect_error(convert_val(-273.15, "kg", "hr"),
                "convert_val: 'to' unit is not one of the accepted unit types.")
   expect_error(convert_val(-273.15, "F", "blah"),
-               "convert_val: \"blah\" unit not recognised")
+               "convert_val: 'blah' unit not recognised")
 })
 
 test_that("convert_val - stops if 'from' input malformed", {
@@ -88,8 +88,8 @@ test_that("convert_val - Mass conversions have expected results", {
 
 
 test_that("convert_val - Area conversions have expected results", {
-  uns_in <- c("mm2", "cm2", "msq", "sqkm")
-  uns_out <- c("mm2", "cm2", "msq", "sqkm")
+  uns_in <- c("mm2", "cm2", "msq", "kmsq")
+  uns_out <- c("mm2", "cm2", "msq", "kmsq")
   vals <- c(2,5, 0.1, 0.0045, 747, 26, 0.00003829, 90000) # random values
   mult <- c(1/1000000, 1/10000, 1, 1000000)
 
@@ -364,4 +364,4 @@ test_that("convert_val - works with vectors", {
   expect_equal(convert_val(10:20, "C", "K"),
                10:20 + 273.15)
 })
-}) ## turns printing back on
+}) ## end capture.output

### 2023-05-04 tests/testthat/test-select_rate.ft.R

@@ -49,6 +49,6 @@ capture.output({  ## stops printing outputs on assigning
   })
 
 
-}) ## turns printing back on
+}) ## end capture.output
 
 

### 2023-05-16 R/select_rate.R
Vig updates

@@ -298,8 +298,8 @@
 #'   The `overlap` method performs two operations:
 #'
 #'   First, regardless of the `n` value, any rate regressions which are
-#'   completely contained within another are removed (this is also the only
-#'   operation if `n = 1`).
+#'   completely contained within another are removed. This is also the only
+#'   operation if `n = 1`.
 #'
 #'   Secondly, for each regression in `$summary` starting from the bottom of the
 #'   summary table (usually the lowest ranked result, but this depends on the

### 2023-11-15 pkgdown/pkgdown_script.R
vig updates

@@ -15,17 +15,20 @@ pkgdown::build_site(lazy = TRUE, devel = TRUE, preview = TRUE)
 # ? not sure what this does
 #pkgdown::preview_site(pkg = ".", path = ".", preview = NA)
 
+# builds top page
 pkgdown::build_home(preview = TRUE)
 
+# build a single article
 pkgdown::build_article("closed")
 
 pkgdown::build_article("flowthrough")
 pkgdown::preview_page('articles/flowthrough.html')
 
+# add a new vignetter
 usethis::use_vignette("intermittent_old", "Intermittent-flow respirometry: Alternative approaches")
 
 
-# sets up GH Actions workflow. Which isn't working.
+# sets up GH Actions workflow. Needs tested.
 #usethis::use_pkgdown_github_pages()
 
 

### 2024-05-01 R/util_funs.R
v2.3.3

@@ -373,7 +373,7 @@ truncate_data <- function(x, from, to, by) {
   if (is.null(by)) by <- "time"
 
   ## verify by just in case
-  by <- by.val(by)
+  by <- by_val(by)
 
   ## replace NULL inputs with defaults
   if(is.null(from)){

### 2024-05-01 R/util_val.R

@@ -34,7 +34,7 @@ column.conflict <- function(inputs, id = FALSE){
 # default - what should be the default if by is NULL?
 # which - which 'by' methods does the function support?
 # msg - start of message text, typically the function name
-by.val <- function(by, req = TRUE, default = NULL,
+by_val <- function(by, req = TRUE, default = NULL,
                    which = c("t", "o", "r"),
                    msg = ""){
 

### 2024-05-01 tests/testthat/test-adjust_rate.R

@@ -4,6 +4,9 @@
 
 capture.output({  ## stops printing outputs on assigning
 
+  if (!identical(Sys.getenv("NOT_CRAN"), "true")) return()
+  skip_on_cran()
+
   test_that("adjust_rate stops with wrong method", {
     expect_error(adjust_rate(100, 10, method = "text"),
                  regexp = "adjust_rate: 'method' input not recognised")

### 2024-05-01 tests/testthat/test-format_time.R

@@ -2,6 +2,9 @@
 ## testthat::test_file("tests/testthat/test-format_time.R")
 ## covr::file_coverage("R/format_time.R", "tests/testthat/test-format_time.R")
 
+if (!identical(Sys.getenv("NOT_CRAN"), "true")) return()
+skip_on_cran()
+
 test_that("format_time works with default inputs", {
 
   x <- c("09-02-03 01:11:11", "09-02-03 02:11:11","09-02-03 02:25:11")

### 2024-05-01 tests/testthat/test-oxy_crit.R

@@ -7,6 +7,9 @@
 
 capture.output({  ## stops printing of outputs on assigning
 
+  if (!identical(Sys.getenv("NOT_CRAN"), "true")) return()
+  skip_on_cran()
+
   # small testing dataset - data.table
   squid <- squid.rd[seq.int(1, nrow(squid.rd), 100), ]
   # pure data.frame

### 2024-05-01 tests/testthat/test-select_rate.R

@@ -7,7 +7,10 @@
 # Sys.setenv(NOT_CRAN = "false")
 
 capture.output({  ## stops printing outputs on assigning
+
+  if (!identical(Sys.getenv("NOT_CRAN"), "true")) return()
   skip_on_cran()
+
   ## objects for testing
   ## convert_rate-from-auto_rate object
   ## contains both negative and positive rates

### 2024-05-01 tests/testthat/test-utils.R

@@ -2,6 +2,8 @@
 # library(testthat)
 # test_file("tests/testthat/test-utils.R")
 
+if (!identical(Sys.getenv("NOT_CRAN"), "true")) return()
+skip_on_cran()
 
 # units.val ------------------------------------------------------------
 

### 2025-02-23 pkgdown/pkgdown_script.R
Website updates

@@ -42,3 +42,8 @@ usethis::use_vignette("intermittent_old", "Intermittent-flow respirometry: Alter
 ## BUT - even if it seems to freeze - check GHub and site - it might have successfully
 ## triggered the 'pages build and deployment' GH action anyway...
 pkgdown::deploy_to_branch()
+
+
+
+pkgdown::build_home_index()
+pkgdown::init_site()
